### 2021-02-19 Analysis/01_data_preparation.Rmd
Analysis INIT

---
title: "Data preparation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Data preparation

 - [ ] Selecting the data from the subclade and their subtrees.
 - [ ] Running the procrustes superimposition