### 2021-02-19 Analysis/02-ancestral_states_estimations.Rmd

---
title: "Ancestral traits estimations"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Ancestral traits estimations

 - [ ] Run the ancestral states estimations on different trees to have a distribution rather than point estimates. Three options here to build the fuzzy trait space:
    * [ ] **Fuzzy average**: run the ace on multiple trees and use their multiple averages as the ace distribution.
    * [ ] **Random fuzzy**: run the ace on multiple trees and randomly sample one value from the estimation
    * [ ] **Fuzzy distribution**: run the ace on multiple trees and sample *n* random values from the estimation
 - [ ] Combine the data together into `dispRity` objects