### 2021-03-22 Analysis/04-elaboration_exploration_analyses.Rmd
evo elavoration.exploration analyses

---
title: "04-elaboration exploration analyses"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Elaboration/exploration analyses

```{r}
## Loading the packages and data
library(dispRity)
load(file = "../Data/Processed/disparity_matrices.rda")
load(file = "../Data/Processed/trees_list.rda")
load(file = "../Data/Processed/spaces_list.rda")
load(file = "../Data/Processed/tmp_simple_matrices.rda")
```

We can add some evo perspective to the basic Endler's elaboration/exploration idea by making the base vector "evolutionary" and by looking at elaboration/exploration through time!

> TODO: I haven't done implementing all the fuzzy space thingies in `dispRity` so for now it's just going to be a naive 1 tree 1 ace 1 answer version of it

## Elaboration/exploration evolutionary metrics

Measure the different elaboration vs exploration metrics with different vectors:

 - [ ] MRCA trajectory: the vector between the root of the tree and the OTU's descendant (`"root-anc"`)
 - [ ] MRCA average trajectory: the vector between the centroid of the group's nodes and the root of the tree (`"cent-anc"`)
 - [ ] Evolutionary living average trajectory: the vector between the root of the tree and the centroid of the living taxa (`"root-tip"`)
 - [ ] Evolutionary overall average trajectory: the vector between the root of the tree and the centroid of the tips and nodes (`"root-cent"`)
 - [ ] Maximum living trajectory: the vector between the two most distant OTUs (`"tip-span"`)
 - [ ] Maximum evolutionary trajectory: the vector between the root and the OTU most nodes away from the root (`"root-nested"`)

```{r}
## Utility functions
project.reject <- function(data, vector, absolute = TRUE) {

    ## Get the projection
    projection <- projections.tree(
          matrix = data$matrix[[1]], 
          tree = data$tree[[1]],
          measure = "position", type = vector)

    ## Get the rejection
    rejection <- projections.tree(
          matrix = data$matrix[[1]], 
          tree = data$tree[[1]],
          measure = "distance", type = vector)

    ## Combine and out
    if(absolute) {
      return(cbind(abs(projection), rejection))
    } else {
      return(cbind(projection, rejection))
    }
}
plot.project.reject <- function(results, data, xlab = "Elaboration", ylab = "Exploration", col = "black", main = "Project/Reject") {

    pch <- c(19, 21)
    col <- c("grey", col)

    ## Find the first node to separate nodes/tips
    root <- which(rownames(data$matrix[[1]]) == "n1")
    ## Split the plot data
    tips_block <- c(1:root-1)
    nodes_block <- c(root:nrow((data$matrix[[1]])))
    ## Split the colours and pch
    colours <- c(rep(col[1], length(nodes_block)),
                 rep(col[2], length(tips_block)))
    pchs <- c(rep(pch[1], length(nodes_block)),
              rep(pch[2], length(tips_block)))
    ## plot the results
    plot(results[c(nodes_block, tips_block), ], xlab = xlab, ylab = ylab, col = colours, pch = pchs, main = main)
}
plot.project.reject.log <- function(results, data, xlab = "Elaboration (log)", ylab = "Exploration (log)", col = "black", main = "Project/Reject") {

    plot.project.reject(log(results), data, xlab, ylab, col, main)
}

```

```{r, fig.height = 12, fig.width = 12}
## MRCA trajectory 
psita_results <- project.reject(tmp_simple_matrices$psita,
                                 vector = c("root", "ancestor"))
chara_results <- project.reject(tmp_simple_matrices$chara,
                                 vector = c("root", "ancestor"))

par(mfrow = c(2,2))
plot.project.reject(psita_results, tmp_simple_matrices$psita, col = "blue", main = "Psittaciformes - root/ancestor")
plot.project.reject(chara_results, tmp_simple_matrices$chara, col = "orange", main = "Charadriiformes - root/ancestor")
plot.project.reject.log(psita_results, tmp_simple_matrices$psita, col = "blue", main = "Psittaciformes - root/ancestor")
plot.project.reject.log(chara_results, tmp_simple_matrices$chara, col = "orange", main = "Charadriiformes - root/ancestor")
```

```{r, fig.height = 6, fig.width = 12}
## Centroid (tips) ancestor trajectory 
psita_results <- project.reject(tmp_simple_matrices$psita,
                                 vector = c("tips", "ancestor"))
chara_results <- project.reject(tmp_simple_matrices$chara,
                                 vector = c("tips", "ancestor"))

par(mfrow = c(1,2))
plot.project.reject.log(psita_results, tmp_simple_matrices$psita, col = "blue", main = "Psittaciformes - tips/ancestor")
plot.project.reject.log(chara_results, tmp_simple_matrices$chara, col = "orange", main = "Charadriiformes - tips/ancestor")
```

```{r, fig.height = 6, fig.width = 12}
## Centroid (nodes) ancestor trajectory 
psita_results <- project.reject(tmp_simple_matrices$psita,
                                 vector = c("nodes", "ancestor"))
chara_results <- project.reject(tmp_simple_matrices$chara,
                                 vector = c("nodes", "ancestor"))

par(mfrow = c(1,2))
plot.project.reject.log(psita_results, tmp_simple_matrices$psita, col = "blue", main = "Psittaciformes - nodes/ancestor")
plot.project.reject.log(chara_results, tmp_simple_matrices$chara, col = "orange", main = "Charadriiformes - nodes/ancestor")
```

```{r, fig.height = 6, fig.width = 12}
## Centroid (root) tips trajectory 
psita_results <- project.reject(tmp_simple_matrices$psita,
                                 vector = c("root", "tips"))
chara_results <- project.reject(tmp_simple_matrices$chara,
                                 vector = c("root", "tips"))

par(mfrow = c(1,2))
plot.project.reject.log(psita_results, tmp_simple_matrices$psita, col = "blue", main = "Psittaciformes - root/tips")
plot.project.reject.log(chara_results, tmp_simple_matrices$chara, col = "orange", main = "Charadriiformes - root/tips")
```

## Elaboration/exploration in the present

Apply these metrics on both clades to get two scores for each taxa (elaboration vs. exploration plots).

## Elaboration/exploration through time

Do some time-slicing and run these metrics through time.

<!-- ## Annex: some potentially interesting ratio thingy

```{r, fig.height = 6, fig.width = 6, eval = FALSE, print = FALSE}
## Loading the whole shape space
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")$PCscores
## Getting the vector from that space
space_vector <- apply(shapespace, 2, quantile, prob = c(0.25, 0.75))
psita_space_proj <- projections(spaces_list$psittaciformes, point1 = space_vector[1,], point2 = space_vector[2, ])
psita_space_reje <- projections(spaces_list$psittaciformes, point1 = space_vector[1,], point2 = space_vector[2, ], measure = "distance")
chara_space_proj <- projections(spaces_list$charadriiformes, point1 = space_vector[1,], point2 = space_vector[2, ])
chara_space_reje <- projections(spaces_list$charadriiformes, point1 = space_vector[1,], point2 = space_vector[2, ], measure = "distance")
## Groups
group <- list("Psittaciformes" = trees_list$psittaciformes[[1]]$tip.label,
              "Charadriiformes" = trees_list$charadriiformes[[1]]$tip.label,
              "Others" = rownames(shapespace)[!(rownames(shapespace) %in% c(trees_list$psittaciformes[[1]]$tip.label, trees_list$charadriiformes[[1]]$tip.label))])
## Full space plot
plot(custom.subsets(shapespace, group = group), col = c("blue", "orange", "grey"), pch = c(19, 19, 21), main = "Bird beak shape space")
arrows(x0 = space_vector[1,1], y0 = space_vector[1,2],
       x1 = space_vector[2,1], y1 = space_vector[2,2], col = "black", lwd = 4)
```

```{r, fig.height = 8, fig.width = 8, eval = FALSE, print = FALSE}
par(mfrow = c(2, 2))
## The elaboration/exploration profiles
plot(psita_space_proj, psita_space_reje, pch = 19, col = "blue",
    xlab = "Elaboration", ylab = "Exploration", main = "Psitta explo/elaboration profile (total)")
plot(chara_space_proj, chara_space_reje, pch = 19, col = "orange",
    xlab = "Elaboration", ylab = "Exploration", main = "Chara explo/elaboration profile (total)")

## The elaboration/exploration profiles
plot(abs(psita_space_proj), psita_space_reje, pch = 19, col = "blue",
    xlab = "Elaboration (absolute)", ylab = "Exploration", main = "Psitta explo/elaboration profile (total)")
plot(abs(chara_space_proj), chara_space_reje, pch = 19, col = "orange",
    xlab = "Elaboration (absolute)", ylab = "Exploration", main = "Chara explo/elaboration profile (total)")
```

Meh... But we can also do ratios: the one between the within group elaboration/exploration and the total elaboration/exploration.
This can help us find the big time explorers/elaborators? (here: the one in the top 99% CI)

```{r, fig.height = 8, fig.width = 8, eval = FALSE, print = FALSE}
## Get the ratio
reje_ratio <- psita_reje/psita_space_reje
proj_ratio <- abs(psita_proj)/abs(psita_space_proj)

## Find the 99% top explorers/elaborators
top_explo <- which(reje_ratio >= quantile(reje_ratio, prob = 0.99))
top_elab <- which(proj_ratio >= quantile(proj_ratio, prob = 0.99))
top_explo_names <- rownames(spaces_list$psittaciformes)[top_explo]
top_elab_names <- rownames(spaces_list$psittaciformes)[top_elab]

plot(proj_ratio, reje_ratio, pch = 19, col = "blue",
    xlab = "Group/space elaboration", ylab = "Group/space exploration", main = "Psittaciformes")
text(cbind(proj_ratio, reje_ratio)[top_explo, ], labels = top_explo_names, pos = 4)
text(cbind(proj_ratio, reje_ratio)[top_elab, ], labels = top_elab_names, pos = 1)
```

```{r, fig.height = 8, fig.width = 8, eval = FALSE, print = FALSE}
## Get the ratio
reje_ratio <- chara_reje/chara_space_reje
proj_ratio <- abs(chara_proj)/abs(chara_space_proj)

## Find the 99% top explorers/elaborators
top_explo <- which(reje_ratio >= quantile(reje_ratio, prob = 0.99))
top_elab <- which(proj_ratio >= quantile(proj_ratio, prob = 0.99))
top_explo_names <- rownames(spaces_list$charadriiformes)[top_explo]
top_elab_names <- rownames(spaces_list$charadriiformes)[top_elab]

plot(proj_ratio, reje_ratio, pch = 19, col = "orange",
    xlab = "Group/space elaboration", ylab = "Group/space exploration", main = "Charadriiformes")
text(cbind(proj_ratio, reje_ratio)[top_explo, ], labels = top_explo_names, pos = 4)
text(cbind(proj_ratio, reje_ratio)[top_elab, ], labels = top_elab_names, pos = 1)
```

Maybe it's picking up some weird birds...
 -->