### 2021-03-25 Analysis/05-mcmcglmm_analyses.Rmd
Update 05-mcmcglmm_analyses.Rmd

---
title: "05 - MCMCglmm analyses"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Elaboration/exploration analyses

```{r}
## Loading the packages and data
library(dispRity)
load(file = "../Data/Processed/disparity_matrices.rda")
load(file = "../Data/Processed/trees_list.rda")
load(file = "../Data/Processed/spaces_list.rda")
load(file = "../Data/Processed/tmp_simple_matrices.rda")
```

The whole discussion from last meeting was on how to determine the axis of elaboration to measure the elaboration/exploration metrics.

Here I am going to explore the major axis from the var-covar matrix from (`MCMCglmm::...()$VCV)` and do this TODO list:

 - [ ] rerun part 1 to 4 using the beak PCA *with* centroid size (`"2020_08_07_MMB_MORPHO_FORMSPACE_FULL.rds"`)
 - [ ] Use the 95% axis from the mcmcglmm vcv matrix (check [here](https://github.com/andbeck/mcmc-plus-tensor))
 - [ ] Use the 95% axis from a vcv matrix from a MANOVA (h-matrix/e-matrix)
 - [ ] Use the x axis from the Phylo-PCA or a Phylo aligned PCA (`geomorph::gm.prcomp()`; see section 3 and 3(4) [here](https://cran.r-project.org/web/packages/geomorph/vignettes/geomorph.PCA.html))