### 2021-04-07 Analysis/06-post_mcmcglmm_analyses.Rmd
post mcmc analyses init

---
title: "Post MCMCglmm analyses"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

In this script I'll be looking at how to analyse the outputs from the MCMCglmm models to get the something about the elaboration/exploration stories.

I'm going to look at it using two approaches:
 * the **blob-wise** one where I'll look at the differences between whole blobs (ellipses) in terms of elaboration/exploration (Robinson & Beckerman style);
 * and the **tip-wise** one where I'll look at the elaboration/exploration score of the tips on the tree (cf. their blobs) relative to the whole phylogeny and to their respective blobs (clades).

## Data

I'm going to focus on the data from Gavin and especially the models 5 and 6 that are:

 * model 6: `model_phylo1_clade3` with one phylogenetic random effect (across the whole tree) and three clade residual effects.
 * model 7: `model_phylo3_clade3` with three phylogenetic random effect (one for each clade) and three clade residual effects.

```{r}
## Loading the correct models
load("../Data/Processed/model_list.rda")
model_phylo1_clade3 <- model_list[[5]]
model_phylo3_clade3 <- model_list[[6]]
```

## Blob-wise approach

Here we're basically comparing multidimensional ellipses (here 3D ones but I think we should push it to more dimensions!).

```{r}

```

## Clade-wise approach

```{r}

```

```{r, eval = FALSE, message = FALSE, fig.width = 12, fig.height = 12, echo = FALSE}
```