### 2021-04-07 Analysis/06-post_mcmcglmm_analyses.Rmd
basic analyses

---
title: "Post MCMCglmm analyses"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

In this script I'll be looking at how to analyse the outputs from the MCMCglmm models to get the something about the elaboration/exploration stories.

I'm going to look at it using two approaches:
 * the **blob-wise** one where I'll look at the differences between whole blobs (ellipses) in terms of elaboration/exploration (Robinson & Beckerman style);
 * and the **tip-wise** one where I'll look at the elaboration/exploration score of the tips on the tree (cf. their blobs) relative to the whole phylogeny and to their respective blobs (clades).

## Data

I'm going to focus on the data from Gavin and especially the models 5 and 6 that are:

 * model 6: `model_phylo1_clade3` with one phylogenetic random effect (across the whole tree) and three clade residual effects.
 * model 7: `model_phylo3_clade3` with three phylogenetic random effect (one for each clade) and three clade residual effects.

```{r}
## Loading the correct models
load("../Data/Processed/model_list.rda")
model_phylo1_clade3 <- model_list[[4]]
model_phylo3_clade3 <- model_list[[6]]
```

And here's what the trait space looks like:

```{r}
## Loading the data
load("../Data/Processed/morphdat.rda")
## Plotting it
colour_vector <- c("orange", "blue", "darkgreen")
plot(morphdat[, c(1,2)], pch = 19,
     col = colour_vector[morphdat$clade],
     xlab = "PC1 (90.5%)",
     ylab = "PC1 (6.86%)")
legend("topleft", legend = levels(morphdat$clade), col = colour_vector, pch = 19)
```

Note that the PC% are from the whole PC in Gavin's example.

## Blob-wise approach

Here we're basically comparing multidimensional ellipses (here 3D ones but I think we should push it to more dimensions!).

### Phylo + clade model

First we can visualise some of these ellipses (100, randomly drawn from the posterior):

```{r}
source("../Functions/plot.ellipses.R")
source("../Functions/get.covar.R")
plot(morphdat[, c(1,2)], pch = 19, cex = 0.5,
     col = colour_vector[morphdat$clade],
     xlab = "PC1 (90.5%)",
     ylab = "PC1 (6.86%)")
plot.ellipses(model_phylo1_clade3, n = 100,
              centre = "none", add = TRUE, col = c("grey",colour_vector))
```

Because this is really messy, we can centre these ellipses on the groups ellipses average centres:

```{r}
plot(morphdat[, c(1,2)], pch = 19, cex = 0.5,
     col = colour_vector[morphdat$clade],
     xlab = "PC1 (90.5%)",
     ylab = "PC1 (6.86%)")
plot.ellipses(model_phylo1_clade3, n = 100,
              centre = "level", add = TRUE, col = c("grey",colour_vector))
```
<!-- ## Get the ellipses centres
centres <- list(
    "phylo" = colMeans(morphdat[, c(1,2)]),
    "gulls" = colMeans(morphdat[which(morphdat$clade == "gulls"), c(1,2)]),
    "plovers" = colMeans(morphdat[which(morphdat$clade == "plovers"), c(1,2)]),
    "sandpipers" = colMeans(morphdat[which(morphdat$clade == "sandpipers"), c(1,2)]))

 -->

### Phylo clade + clade model

```{r, warn = FALSE}
plot(morphdat[, c(1,2)], pch = 19, cex = 0.5,
     col = colour_vector[morphdat$clade],
     xlab = "PC1 (90.5%)",
     ylab = "PC1 (6.86%)")
plot.ellipses(model_phylo3_clade3, n = 100,
              centre = "none", add = TRUE, col = c("black", colour_vector, colour_vector))
```

This looks much more messier (and also probably not scaled correctly?)...

```{r}
plot(morphdat[, c(1,2)], pch = 19, cex = 0.5,
     col = colour_vector[morphdat$clade],
     xlab = "PC1 (90.5%)",
     ylab = "PC1 (6.86%)")
plot.ellipses(model_phylo3_clade3, n = 100,
              centre = "level", add = TRUE, col = c("black", colour_vector, colour_vector))
```

## Tip-wise approach