### 2021-05-21 Analysis/Older_ideas/01_data_preparation.Rmd

---
title: "Data preparation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r}
library(ape)
library(dispRity)
```

# Data preparation

Here I'm going to focus on the Psittaciformes (parrots and stuff) and Charadriiformes (sea gulls and stuff).

 - [x] Selecting the data from the subclade and their subtrees.
 - [x] Selecting the shape data

## Tree data

For this pilote analysis, I will only select 10 random trees between both Ericson and Hackett backbone trees from Jetz et al big trees (5 each).
To find the mrca to get the subtrees, we will use the pairs `"Micropsitta"` and `"  Nestor"` for the parrots and `"Ichthyaetus"` and `"Burhinus"` for the sea gulls

> Note

```{r, eval = FALSE}
## Get the pairs of birds for both clades
psita_clade <- c("Micropsitta_meeki", "Nestor_notabilis")
chara_clade <- c("Burhinus_oedicnemus", "Larus_saundersi")

## Number of trees to select
n_trees <- 10

## Reading the trees
ericson_tree_1_1000 <- read.tree("../Data/Raw/Tree_data/AllBirdsEricson1.tre")
hackett_tree_1_1000 <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")

## Selecting n/2 random trees for both tree
selected_trees <- replicate(2, sample(1:1000, n_trees/2))

## Sampling the trees
ericson_tree_5 <- ericson_tree_1_1000[selected_trees[,1]]
hackett_tree_5 <- hackett_tree_1_1000[selected_trees[,2]]
## Combine the trees
all_birds_10 <- c(ericson_tree_5, hackett_tree_5)

## Extracting the different clades from the different trees
wrap.extract.clade <- function(tree, tips) {
    ## Extract the clade
    return(extract.clade(tree, node = getMRCA(tree, tip = tips)))
}

## Get the two trees
chara_trees_10 <- lapply(all_birds_10, wrap.extract.clade, tip = chara_clade)
psita_trees_10 <- lapply(all_birds_10, wrap.extract.clade, tip = psita_clade)
class(chara_trees_10) <- class(psita_trees_10) <- "multiPhylo"

## Save as a list of trees
trees_list <- list("charadriiformes" = chara_trees_10,
                   "psittaciformes"  = psita_trees_10)
save(trees_list, file = "../Data/Processed/trees_list.rda")
```

## Beak data

```{r}
## Loading the tree data (to have the species list)
load("../Data/Processed/trees_list.rda")
```

Note that here we are extracting the sets of birds (orders) from the global (full) shape space.
This means that for the following analyses we're only looking at sub-parts in that space, i.e. the PCA does not contain all the data.
If this is undersirable, I guess we can always do the whole pipeline again (select species > procrustes per order > pca per order).

```{r}
## Loading the whole shapespace
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")

## Select the subspecies per shapespace
psita_space <- shapespace$PCscores[rownames(shapespace$PCscores) %in% trees_list$psittaciformes[[1]]$tip.label, ]
chara_space <- shapespace$PCscores[rownames(shapespace$PCscores) %in% trees_list$charadriiformes[[1]]$tip.label, ]

## Clean the tree lists and the dataset
cleaned_psita <- clean.data(psita_space, trees_list$psittaciformes)
cleaned_chara <- clean.data(chara_space, trees_list$charadriiformes)

## Re-saving the trees and saving the spaces
spaces_list <- list(psittaciformes  = cleaned_psita$data,
                    charadriiformes = cleaned_chara$data)

## Adding the node labels for each selected trees
psita_trees <- lapply(cleaned_psita$tree, function(tree) {tree$node.label <- paste0("n", 1:Nnode(tree)); return(tree)})
chara_trees <- lapply(cleaned_chara$tree, function(tree) {tree$node.label <- paste0("n", 1:Nnode(tree)); return(tree)})
class(psita_trees) <- class(chara_trees) <- "multiPhylo"

trees_list  <- list(psittaciformes  = psita_trees,
                    charadriiformes = chara_trees)

save(trees_list,  file = "../Data/Processed/trees_list.rda")
save(spaces_list, file = "../Data/Processed/spaces_list.rda")
```

## Space visualisation

The full birdy space:

```{r, fig.height = 12, fig.width = 12}
## Loading the full space
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")$PCscores
## Groups
group <- list("Psittaciformes" = trees_list$psittaciformes[[1]]$tip.label,
              "Charadriiformes" = trees_list$charadriiformes[[1]]$tip.label,
              "Others" = rownames(shapespace)[!(rownames(shapespace) %in% c(trees_list$psittaciformes[[1]]$tip.label, trees_list$charadriiformes[[1]]$tip.label))])
## Full space plot
plot(custom.subsets(shapespace, group = group), col = c("blue", "orange", "grey"), pch = c(19, 19, 21), main = "Bird beak shape space")
```