### 2021-05-27 Analysis/major.axis.tex
major axis latex formulas

\documentclass[a4paper,11pt]{article}

\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{lastpage}
\usepackage{url}
\usepackage{hyperref}
\pagenumbering{arabic}
\linespread{1.66}

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}
\begin{center}

%Title
\noindent{\Large{\bf{Major axes estimation}}}\
\bigskip
%Author
Thomas Guillerme\\href{mailto:t.guillerme@imperial.ac.uk}{t.guillerme@imperial.ac.uk}

For \bold{O_{n}}, the unit hypersphere matrix of \textit{n} dimensions and of radius being composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation*}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation*}

(i.e $O_{n}$ is the matrix representing each edge point of the unit hyper sphere of \textit{n} dimensions and of radius 1);

And \bodel{O'_{n}} the scaled matrix hyper sphere to the right confidence interval size using the $\chi^2$ distribution:

\begin{equation}
    \bold{O'_{n}} = \bold{O_{n}} \sqrt{\chi^2(0.95)}
\end{equation}

For the variance-covariance matrix \textit{VCV_{n}} of \textit{n} dimensions

\begin{equation*}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation*}

and the eigenvectors \bold{v} and the eigenvalues \lambda satisfying the following eigen decomposition:

\begin{equation}
    VCV_{n}\bold{v} = \lambda \bold{v}
\end{equation}

We can get $M_{n}$ matrix containing all the edges' coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors \bold{v} and the product of the scale 0.95 CI unit sphere \bold{O'_{n}} and the eigenvalues \lambda:

\begin{equation}
    M_{n} = [(\bold{O'_{n}}\sqrt{\lambda}) \times \bold{v}]^{\text{T}}
\end{equation}

Where M_{1,n} is the major axes of the 0.95 CI hyper ellipse fitting the variance-covariance matrix.

\end{center}

\end{document}