### 2021-05-28 Analysis/equations.tex
equation draft board

\documentclass[a4paper,11pt]{article}

\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{lastpage}
\usepackage{url}
\usepackage{hyperref}
\pagenumbering{arabic}
\linespread{1.66}

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}
\begin{center}

%Title
\noindent{\Large{\bf{Major axes estimation}}}\
\bigskip
%Author
Thomas Guillerme\\href{mailto:t.guillerme@imperial.ac.uk}{t.guillerme@imperial.ac.uk}

For \bold{O_{n}}, the unit hypersphere matrix of \textit{n} dimensions and of radius being composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation*}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation*}

(i.e $O_{n}$ is the matrix representing each edge point of the unit hyper sphere of \textit{n} dimensions and of radius 1);

And \bodel{O'_{n}} the scaled matrix hyper sphere to the right confidence interval size using the $\chi^2$ distribution:

\begin{equation}
    \bold{O'_{n}} = \bold{O_{n}} \sqrt{\chi^2(0.95)}
\end{equation}

For the variance-covariance matrix \textit{VCV_{n}} of \textit{n} dimensions

\begin{equation*}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation*}

and the eigenvectors \bold{v} and the eigenvalues \lambda satisfying the following eigen decomposition:

\begin{equation}
    VCV_{n}\bold{v} = \lambda \bold{v}
\end{equation}

We can get $M_{n}$ matrix containing all the edges' coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors \bold{v} and the product of the scale 0.95 CI unit sphere \bold{O'_{n}} and the eigenvalues \lambda:

\begin{equation}
    M_{n} = [(\bold{O'_{n}}\sqrt{\lambda}) \times \bold{v}]^{\text{T}}
\end{equation}

Where M_{1,n} is the major axes of the 0.95 CI hyper ellipse fitting the variance-covariance matrix.

\end{center}

For any vectors $\vec{a}$ and $\vec{b}$, defined either as one set or a pair of sets of coordinates in $n$ dimensions:

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x \\
    y \\
    ... \\
    n \\
    \end{bmatrix}
\end{equation}

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x_{1} & x_{2} \\
    y_{1} & y_{2} \\
    ... & ... \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

We can calculate $\vec{a_{1}}$, the orthogonal projection of $\vec{a}$ onto $\vec{b}$ using:

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\vec{b}\|}
\end{equation}

With $\|\vec{b}\| = \sqrt{\vec{b} \cdot \vec{b}}$ being the norm of $\vec{b}$.

 And $\vec{a_{2}}$, the rejection of $\vec{a}$ onto $\vec{b}$:

\begin{equation}
    \vec{a_{2}} = \vec{a} - \vec{a_{1}}
\end{equation}

And $\theta$, the angle formed by $\vec{a}$ and $\vec{b}$:

\begin{equation}
    \theta = acos \left(\frac{\vec{a} \cdot \vec{b}}{\|\vec{a}\| \cdot \|\vec{b}\|} \right)
\end{equation}

Given a $n$ dimensional trait space $TS_{m,n}$ with $m$ elements defined as:

\begin{equation*}
TS_{m,n} = 
    \begin{bmatrix}
        x_{1} & x_{2} & \cdots & x_{m} \\
        y_{1} & y_{2} & \cdots & y_{m} \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        n_{1} & n_{2} & \cdots & n_{m} \\
    \end{bmatrix}
\end{equation*}

And a base vector $B_{n}$ defined as:

\begin{equation*}
B_{n} = 
    \begin{bmatrix}
        Bx_{1} & Bx_{2}\\
        By_{1} & By_{2}\\
        \vdots  & \vdots \\
        Bn_{1} & Bn_{2}\\
    \end{bmatrix}
\end{equation*}

\begin{equation}
        cTS = TS - [Bx_{1}, By_{1}, \cdots, Bn_{1}]^{\text{T}}
\end{equation}

\begin{equation}
I_{\vec{B}} - \vec{B}\vec{B}^\text{T} - \vec{\hat{\imath}}\vec{\hat{\imath}}^\text{T} + [\vec{B} \vec{\hat{\imath}}]     \begin{bmatrix}
        cos(\theta) & -sin(\theta)\\
        sin(\theta) & cos(\theta)\\
    \end{bmatrix} [\vec{B} \vec{\hat{\imath}}]^\text{T}
\end{equation}

Where $\theta$ is:

\begin{equation}
    \theta = acos \left(\frac{\vec{B} \cdot \vec{\hat{\imath}}}{\|\vec{B}\| \cdot \|\vec{\hat{\imath}}\|} \right)
\end{equation}

Or $\theta = acos (B_x)$ since both \|\vec{B}\| and \|\vec{\hat{\imath}}\| are equal to 1 and \|\vec{\hat{\imath}}\| is the unit vector on the first axis.

\end{document}