### 2021-06-07 Analysis/00-methods_overview.Rmd
Methods overview INIT

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

All the custom code and custom functions called in this document can be downloaded and installed in `R` as the standalone working package `beer` (Beak Elaboration and Exploration in R).
Although `beer` is intentionally designed to be portable and shareable, we advise workers to use it with moderation and tailor its consumption to their specific research needs.
Some analyses in `beer` do take some time to run so we advise users not to operate heavy machinery while using it.
Don't use it while driving.
The following code snippets are used to illustrate the implementation of the functions.
See <!-- §§§run_replicates_script --> for running the actual analyses.

The following contains an overview of the methods used in the paper.

# Data

The detailed and repeatable version of this section can be found [here](01-Data_preparation.Rmd).

We used data.

# MCMCglmm

The detailed and repeatable version of this section can be found [here](02-MCMCglmm_mini_chains.Rmd).

# Elaboration and exploration

The detailed and repeatable version of this section can be found [here](03-elaboration_exploration_analyses.Rmd).

## References