### 2021-07-20 Analysis/08-Number_of_dimensions.Rmd
Create 08-Number_of_dimensions.Rmd

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

Checking the ideal number of dimensions for running the model.

```{r, load pkg and data}
library(beer)
library(MCMCglmm)
data(chara_data)
data(chara_tree)
```

```{r}
## Get the groups
beaks <- readRDS("MCMCglmm_covar/Data/2020_06_19_MMB_LANDMARKS_ALL.rds")
order_groups <- lapply(as.list(unique(beaks$Order)), function(X) unique(beaks$TipLabel[beaks$Order %in% X]))
names(order_groups) <- unique(beaks$Order)
family_groups <- lapply(as.list(unique(beaks$Family)), function(X) unique(beaks$TipLabel[beaks$Family %in% X]))
names(family_groups) <- unique(beaks$Family)

## Get the PCA
load("MCMCglmm_covar/full_PCAs.rda")

## Selecting axis function
# @params strict whether to have less than the threshold (strict) or at least the threshold (false)
select.axis <- function(ordination, threshold = 0.99, group, details = FALSE, strict = FALSE) {
    ## Set the group
    if(missing(group)) {
        group <- list(rownames(ordination))
    }

    ## Cum sum scaled variance
    get.cum.var <- function(one_group, matrix) {
        if(length(one_group) != 1) {
            vars <- apply(matrix[one_group, ], 2, var)
            return(cumsum(vars/sum(vars)))
        } else {
            return(c(0, rep(1, ncol(matrix)-1)))
        }
    }

    ## Get the variances for each group
    group_vars <- lapply(group, get.cum.var, matrix = ordination)

    select.threshold <- function(x, threshold, strict) {
        if(strict) {
            return(which(x < threshold))
        } else {
            return(c(0, which(x < threshold)) + 1)
        }
    }

    ## Get the axis per threshold
    selected_dims <- lapply(group_vars, select.threshold, threshold = threshold, strict = strict)

    ## Return the results
    if(!details) {
        return(unique(unlist(selected_dims)))
    } else {
        return(selected_dims)
    }
}

## Preparing the result table
thresholds <- seq(from = 0.9, to = 0.99, by = 0.01)
results <- matrix(0, 6, length(thresholds))
rownames(results) <- c("sizefree-all", "sizefree-order", "sizefree-family", "sizeform-all", "sizeform-order", "sizeform-family")
colnames(results) <- thresholds

for(i in 1:length(thresholds)) {
  results[1,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i]))
  results[2,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i], group = order_groups))
  results[3,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i], group = family_groups))
  results[4,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i]))
  results[5,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i], group = order_groups))
  results[6,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i], group = family_groups))
}

## Colour vector
colours <- c("purple", "red", "orange",
             "blue", "cyan", "green")

## Plotting the results
plot(NULL, xaxt = "n", xlim = c(1,10), ylim = range(results), ylab = "Dimensions", xlab = "Thresholds")
axis(1, at = c(1:10), labels = thresholds)
abline(h = 7, col = "lightgrey")
for(i in 1:6) {
  lines(results[i,], col = colours[i])
}
legend("topleft", legend = rownames(results), lty = 1, col = colours)
```