### 2021-07-21 Analysis/08-Number_of_dimensions.Rmd
Updated PC axis analyses

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

```{r, echo = FALSE, print = FALSE}
library(beer)
library(MCMCglmm)
data(chara_data)
data(chara_tree)
```

## Number of dimensions required to capture at least X% of variance per group

```{r, echo = FALSE}
## Get the groups
beaks <- readRDS("MCMCglmm_covar/Data/2020_06_19_MMB_LANDMARKS_ALL.rds")
order_groups <- lapply(as.list(unique(beaks$Order)), function(X) unique(beaks$TipLabel[beaks$Order %in% X]))
names(order_groups) <- unique(beaks$Order)
family_groups <- lapply(as.list(unique(beaks$Family)), function(X) unique(beaks$TipLabel[beaks$Family %in% X]))
names(family_groups) <- unique(beaks$Family)

## Get the PCA
load("MCMCglmm_covar/full_PCAs.rda")

## Selecting axis function
# @params strict whether to have less than the threshold (strict) or at least the threshold (false)
select.axis <- function(ordination, threshold = 0.99, group, details = FALSE, strict = FALSE) {
    ## Set the group
    if(missing(group)) {
        group <- list(rownames(ordination))
    }

    ## Cum sum scaled variance
    get.cum.var <- function(one_group, matrix) {
        if(length(one_group) != 1) {
            vars <- apply(matrix[one_group, ], 2, var)
            return(cumsum(vars/sum(vars)))
        } else {
            return(c(0, rep(1, ncol(matrix)-1)))
        }
    }

    ## Get the variances for each group
    group_vars <- lapply(group, get.cum.var, matrix = ordination)

    select.threshold <- function(x, threshold, strict) {
        # if(strict) {
            return(which(x < threshold))
        # } else {
        #     return(c(0, which(x < threshold)) + 1)
        # }
    }

    ## Get the axis per threshold
    selected_dims <- lapply(group_vars, select.threshold, threshold = threshold, strict = strict)

    ## Return the results
    if(!details) {
        return(unique(unlist(selected_dims)))
    } else {
        return(selected_dims)
    }
}

## Preparing the result table
thresholds <- seq(from = 0.9, to = 0.99, by = 0.01)
results <- matrix(0, 6, length(thresholds))
rownames(results) <- c("sizefree-all", "sizefree-order", "sizefree-family", "sizeform-all", "sizeform-order", "sizeform-family")
colnames(results) <- thresholds

for(i in 1:length(thresholds)) {
  results[1,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i]))
  results[2,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i], group = order_groups))
  results[3,i] <- length(select.axis(full_PCAs$sizefree, threshold = thresholds[i], group = family_groups))
  results[4,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i]))
  results[5,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i], group = order_groups))
  results[6,i] <- length(select.axis(full_PCAs$sizeform, threshold = thresholds[i], group = family_groups))
}

## Colour vector
colours <- c("orange", "yellow", "pink",
             "lightblue", "lightgreen", "cyan")

## Plotting the results
plot(NULL, xaxt = "n", xlim = c(1,10), ylim = range(results), ylab = "Dimensions", xlab = "Thresholds")
axis(1, at = c(1:10), labels = thresholds)
abline(h = 7, col = "lightgrey")
for(i in 1:6) {
  lines(results[i,], col = colours[i])
}
legend("topleft", legend = rownames(results), lty = 1, col = colours)
```

This is the number of dimensions required per PCA contain X% of variance per group. The groups being either the full PCA (`all` - the classic way to measure the variance threshold) or each order (`order`) or family (`family`).
The number of dimensions per group is measured either for the PCA with the size component removed (`sizefree`) and for the PCA with the size component (`sizeform`).

## Variance per specific axis for the 37 first dimensions

```{r, echo = FALSE, print = FALSE}
## Calculating the variance for each dimension
get.dim.vars <- function(ordination, group, n_dims) {
    ## Set the group
    if(missing(group)) {
        group <- list(rownames(ordination))
    }
    if(missing(n_dims)) {
        n_dims <- ncol(ordination)
    }

    ## Cum sum scaled variance
    get.scaled.var <- function(one_group, matrix, n_dims) {
        if(length(one_group) != 1) {
            vars <- apply(matrix[one_group, ], 2, var)
            return(c(vars/sum(vars))[1:n_dims])
        } else {
            return(c(1, rep(0, ncol(matrix)-1))[1:n_dims])
        }
    }
    ## Get the variances for each group
    return(lapply(group, get.scaled.var, matrix = ordination, n_dims = n_dims))
}
## Plotting results
barplot.results <- function(result, main, threshold = c(0.95, 0.99), col = "orange") {
  barplot(result, col = col, main = main, ylab = "Scaled variance")
  abline(v = length(which(cumsum(result) < threshold[1])), lwd = 2)
  abline(v = length(which(cumsum(result) < threshold[2])), lwd = 2, col = "grey")
}

## Get all the detailed results
## Full PCA
sizefree_fullpca <- get.dim.vars(full_PCAs$sizefree,
                                 n_dims = 37)
names(sizefree_fullpca) <- c("Sizefree: all species")
## Orders
sizefree_order   <- get.dim.vars(full_PCAs$sizefree,
                                 group = order_groups,
                                 n_dims = 37)
names(sizefree_order) <- paste("Sizefree:", names(order_groups))
sizefree_order <- sizefree_order[match(sort(names(sizefree_order)), names(sizefree_order))]
## Family
sizefree_family  <- get.dim.vars(full_PCAs$sizefree,
                                 group = family_groups,
                                 n_dims = 37)
names(sizefree_family) <- paste("Sizefree:", names(family_groups))
sizefree_family <- sizefree_family[match(sort(names(sizefree_family)), names(sizefree_family))]

## Full PCA
sizeform_fullpca <- get.dim.vars(full_PCAs$sizeform,
                                 n_dims = 37)
names(sizeform_fullpca) <- c("Sizeform: all species")
## Orders
sizeform_order   <- get.dim.vars(full_PCAs$sizeform,
                                 group = order_groups,
                                 n_dims = 37)
names(sizeform_order) <- paste("Sizeform:", names(order_groups))
sizeform_order <- sizeform_order[match(sort(names(sizeform_order)), names(sizeform_order))]
## Family
sizeform_family  <- get.dim.vars(full_PCAs$sizeform,
                                 group = family_groups,
                                 n_dims = 37)
names(sizeform_family) <- paste("Sizeform:", names(family_groups))
sizeform_family <- sizeform_family[match(sort(names(sizeform_family)), names(sizeform_family))]

```

Here I will display the scaled variance per axis for each group (as mentioned above).
For example, for the full PCA, the results are pretty common and easy to grasp:

```{r, fig.height = 4, fig.width = 4, echo = FALSE, print = FALSE, result = FALSE}
## Plotting everything
silent <- mapply(barplot.results, sizefree_fullpca, as.list(names(sizefree_fullpca)))
## Plotting everything
silent <- mapply(barplot.results, sizeform_fullpca, as.list(names(sizeform_fullpca)), MoreArgs = list(col = colours[4]))
```

When considering all the data, the first three axis contain 95% of the variance (black vertical line) and PC1 to 6 for 99% of the variance (grey vertical line).
For the PCA with size, the first axes contains 95% of the variance and the PC1 to 3 contains 99%.
The rest of the variance is distributed pretty log-normally as expected.

We can then measure the same be measuring things per family or per order (i.e. within the PCA, how many axis are required to represent 95% of the variance of order/family X).
The results vary much more depending on the order/family! For example, in Strigiformes, the first PC does not contain much variance at all compared to the other axes!
Note that the order with just one bar at 100% just have one species so they don't count.

```{r, fig.height = 4, fig.width = 4, echo = FALSE, print = FALSE, result = FALSE}
## Plotting everything
silent <- mapply(barplot.results, sizefree_order, as.list(names(sizefree_order)), MoreArgs = list(col = colours[2]))
```

```{r, fig.height = 4, fig.width = 4, echo = FALSE, print = FALSE, result = FALSE}
## Plotting everything
silent <- mapply(barplot.results, sizefree_family, as.list(names(sizefree_family)), MoreArgs = list(col = colours[3]))
```

```{r, fig.height = 4, fig.width = 4, echo = FALSE, print = FALSE, result = FALSE}
## Plotting everything
silent <- mapply(barplot.results, sizeform_order, as.list(names(sizeform_order)), MoreArgs = list(col = colours[5]))
```

```{r, fig.height = 4, fig.width = 4, echo = FALSE, print = FALSE, result = FALSE}
## Plotting everything
silent <- mapply(barplot.results, sizeform_family, as.list(names(sizeform_family)), MoreArgs = list(col = colours[6]))
```