### 2021-08-10 Analysis/07-benchmarking.Rmd
Update 07-benchmarking.Rmd

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

Benchmarking `MCMCglmm` running times

```{r, load pkg and data}
library(beer)
library(MCMCglmm)
data(chara_data)
data(chara_tree)
```

Measure:
 * `size` the object size
 * `d_size` the object size difference compared to the first test
 * `r_size` the object size increase rate (estimated as `d_size/iteration_differences`)
 * `time` the elapsed time
 * `d_time` the elapsed time difference compared to the first test
 * `r_time` the elapsed time increase rate (estimated as `d_time/iteration_differences`)

Variables:
 * number of dimensions
 * thinning rate
 * saving parameters from `MCMCglmm`

> Here I'm not going to test the saving parameters since it's pretty much a constant RAM saving. By removing all the un-necessary stuff to save, we have 4.2 times smaller output!

Protocol:
 1. Run one chain with the set number of parameters for 10 iterations (thinning = 1)
 2. Run it again for 20 iterations (thinning = 1)
 3. Run it again for 30 iterations (thinning = 1)
 4. Run it again for 40 iterations (thinning = 1)
 5. Run it again for 50 iterations (thinning = 1)

```{r}
## Function for setting up the parameters
set.params <- function(dimensions, thin, nitt) {
  ## Set up the parametrising chains models
  param1 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*1,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param2 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*2,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param3 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*3,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param4 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*4,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param5 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*5,
                                 thin   = thin,
                                 burnin = 0))
  return(list(param1,
              param2,
              param3,
              param4,
              param5))
}
## Function for running the test
test.params <- function(dimensions, thin, nitt) {
  ## Setting up the parameters
  param_list <- set.params(dimensions = dimensions, thin = thin, nitt = nitt)

  ## First iteration
  timer_base <- system.time(param_chains <- run.mini.chains(param_list[[1]], replicates = 1))["elapsed"]
  chain_size_base <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt, dimensions, sep = ", ")
  data <- paste(timer_base, 0, timer_base/nitt, chain_size_base, 0, chain_size_base/nitt, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Second run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[2]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*2, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*1, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*1, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Third run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[3]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*3, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*2, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*2, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Fourth run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[4]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*4, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*3, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*3, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Fifth run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[5]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*5, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*4, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*4, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  return(NULL)
}
```

## Running the tests for 1k generations (sampling every 100)

```{r, eval = FALSE}
## Initialising the results result table
system(paste("echo \"thin, iterations, dimensions, time, d_time, r_time, size, d_size, r_size\" > recording.csv"))

## Running the test
for(dim in 2:17) {
  test.params(dimensions = dim, thin = 100, nitt = 1000)
}
```

```{r}
## Upload the table
rable <- read.csv("recording.csv")

## Log values
run_time <- log(rable[, "time"])
run_size <- log(rable[, "size"])
cols <- rainbow(16)

par(mfrow = c(2,1))
plot(NULL, xlim = c(1, 5), ylim = range(run_time), main = "Run time", xlab = "10s of samples (per 1k iterations)", ylab = "log elapsed seconds")
for(i in 1:(nrow(rable)/5)) {
  lines(run_time[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_size), main = "Object size", xlab = "10s of samples (per 1k iterations)", ylab = "log bytes")
for(i in 1:(nrow(rable)/5)) {
  lines(run_size[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}
```

## Running the tests for 10k generations (sampling every 1000)

```{r, eval = FALSE}
## Running the test
for(dim in 2:17) {
  test.params(dimensions = dim, thin = 1000, nitt = 10000)
}
```

```{r}
## Upload the table
rable <- read.csv("recording.csv")
rable_1k <- rable[1:80, ]
rable_10k <- rable[-c(1:80), ]

## Log values
run_time_1k <- rable_1k[, "time"]
run_size_1k <- rable_1k[, "size"]
run_time_10k <- rable_10k[, "time"]
run_size_10k <- rable_10k[, "size"]

cols <- rainbow(16)

par(mfrow = c(2,2))
plot(NULL, xlim = c(1, 5), ylim = range(run_time_1k), main = "Run time", xlab = "iterations (thin = 100)", ylab = "elapsed seconds", xaxt = "n")
axis(1, at = c(1:5), labels = seq(from = 1000, to = 5000, by = 1000))
for(i in 1:(nrow(rable_1k)/5)) {
  lines(run_time_1k[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_time_10k), main = "Run time", xlab = "iterations (thin = 1000)", ylab = "elapsed seconds", xaxt = "n")
axis(1, at = c(1:5), labels = seq(from = 10000, to = 50000, by = 10000))
for(i in 1:(nrow(rable_10k)/5)) {
  lines(run_time_10k[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_size_1k), main = "Object size", xlab = "iterations (thin = 100)", ylab = "bytes", xaxt = "n")
axis(1, at = c(1:5), labels = seq(from = 1000, to = 5000, by = 1000))
for(i in 1:(nrow(rable_1k)/5)) {
  lines(run_size_1k[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_size_10k), main = "Object size", xlab = "iterations (thin = 1000)", ylab = "bytes", xaxt = "n")
axis(1, at = c(1:5), labels = seq(from = 10000, to = 50000, by = 10000))
for(i in 1:(nrow(rable_10k)/5)) {
  lines(run_size_10k[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}
```