### 2021-09-24 Analysis/03-elaboration_exploration_analyses.Rmd
Update 03-elaboration_exploration_analyses.Rmd

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

<!-- TG: The analyses will be eventually run in dispRity so we will miss out the following notices: 
Some analyses in `beer` do take some time to run so we advise users not to operate heavy machinery while using it.
Don't use it while driving. -->
The following code snippets are used to illustrate the implementation of the functions.

# Elaboration and exploration analyses

```{r, eval = TRUE, echo = FALSE, print = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen", "grey")
```

## Extracting the variance-covariance matrices from the MCMCglmm mini-chains

To analyse the exploration and elaboration aspects of both clades and tips, we first selected 1000 random covariance matrices from the combined MCMCglmm posterior variance-covariance matrices represented in this space.

```{r, eval = TRUE, echo = FALSE}
covar_matrices <- MCMCglmm.dispRity(
                  data          = demo_data,
                  posteriors    = combined_results,
                  group         = MCMCglmm.levels(combined_results)[1:4],
                  rename.groups = c("gulls", "plovers", "sandpipers", "phylogeny"))
```

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, echo = FALSE}
## Plotting just the space
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "demo space",
            legend = TRUE)
```

From this space and the mini chains run before, we can extract the posterior variance covariance matrices to calculate the major axes of variation in the `r dim(length(covar_matrices$call$dimensions))` dimensions trait space at different levels:

 * The whole bird phylogeny level ("phylogeny level");
 * And the named node levels ("clade levels").

These major axes were calculated as the longest distance within the 95% confidence interval hyper-ellipse from the variance-covariance matrix (i.e. the major axes of the 95% CI ellipse of matrix).

```{r, echp = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, echo = FALSE}
get.one.axis <- function(data, axis = 1, level = 0.95, dimensions) {

    # The magic: https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331

    ## Select the right dimensions
    data$VCV <- data$VCV[dimensions, dimensions, drop = FALSE]

    ## Get the data dimensionality
    dims <- dim(data$VCV)[1]

    ## Create the unit hypersphere (a hypersphere of radius 1) for the scaling
    unit_hypersphere1 <- unit_hypersphere2 <- matrix(0, ncol = dims, nrow = dims)
    ## The "front" (e.g. "top", "right") units
    diag(unit_hypersphere1) <- 1
    ## The "back" (e.g. "bottom", "left") units
    diag(unit_hypersphere2) <- -1
    unit_hypersphere <- rbind(unit_hypersphere1, unit_hypersphere2)
    ## Scale the hypersphere (where level is the confidence interval)
    unit_hypersphere <- unit_hypersphere * sqrt(qchisq(level, 2))

    ## Do the eigen decomposition (symmetric - faster)
    eigen_decomp <- eigen(data$VCV, symmetric = TRUE)

    ## Re-scaling the unit hypersphere
    scaled_edges <- unit_hypersphere * rep(sqrt(eigen_decomp$values), each = dims*2)
    ## Rotating the edges coordinates
    edges <- tcrossprod(scaled_edges, eigen_decomp$vectors)

    ## Move the matrix around
    edges <- edges + rep(data$Sol[dimensions, drop = FALSE], each = dims*2)

    ## Get the edges coordinates
    return(edges[c(axis, axis+dims), , drop = FALSE])
}

set.seed(42)
## Space plot
cor_matrix <- matrix(cbind(1, 0.8, 0.8, 1), nrow = 2)
space <- space.maker(50, 2, rnorm, cor.matrix = cor_matrix)
lim <- c(floor(range(space)[1]), ceiling(range(space)[2]))
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim)

## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix), col = "blue", lwd = 3)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0))), lwd = 3, col = "orange")
```

Figure: the main axis of variation (in orange) of the 95% CI ellipse of the variance-covariance matrix (in blue) of 20 elements in 2 dimensions (in black).
Note that when estimated using a `mcmcmcglmmm` approach there are as many 95% CI ellipses and corresponding main axes as they are posterior samples.

In practice, we calculated the coordinates of these axes end points  directly from the cross product between the 95% unit hypersphere scaled by the variance-covariance matrix's squared eigenvalues and the variance-covariance matrix's eigenvectors (see details 1).

One can visualise these ellipses and major axes in 2D using the `sauron.plot`functions with the `major.axis = mean` and `ellipses = mean` options (based on @ellipse) to display the mean major.axis and the mean ellipses:

```{r, echo = FALSE, fig.height = 6, fig.width = 6}
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "Average ellipses and\nmajor axes",
            major.axes = mean,
            ellipses = mean,
            legend = TRUE,
            cex = 0.5)
```

We can also visualise 100 random major axes and 100 random ellipses:

```{r, echo = FALSE, fig.height = 12, fig.width = 6}
par(mfrow = c(2,1))
## Plotting 100 axes
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random major axes",
            major.axes = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
## Plotting 100 ellipses
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random ellipses",
            ellipses = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
```

We can then compare these axes either with each other or relative to the tip positions in the trait space using elaboration and exploration scores as described below.

## Elaboration and exploration scores calculations

From the major axes calculated above, we can use linear algebra in $n$ dimensions to estimate exploration and elaboration scores for clades or tips of the tree.
These scores are a direct interpretation of @endler2005's elaboration and innovation (here exploration) patterns (Fig 1. in @endler2005):
Given a major axis of evolutionary "trajectory" (in the sense of evolutionary trend, not an evolutionary direction), we can calculate the elaboration score as the projection value of any vector defined by a clade or a tip on the major axis and the exploration (or "innovation" in @endler2005) as that vector's rejection.

```{r, echo = FALSE, eval = TRUE}
source("../Functions/test.rotation.R")
```

Figure: consider a space with five elements named "a" to "e" (in grey) and one focal axes in this space defined as the vector from point1: "d" to point2: "e" (the thin solid grey line linking the grey "d" and "e"). We can then define all points from "d" (point1) as vectors (the dashed grey lines). We can then rotate the whole space so that the vector "d" to "e" becomes the reference coordinate vector (0,1) (the thick grey line) and rotate and translate all the other vectors accordingly (the solid grey lines linking the coloured letters "a" to "c"). We can then project each of these vectors on the reference vector and measure their distance *on* the reference vector (the projection; the coloured solid lines) and their distance *from* the reference vector (the rejection; the coloured dotted lines). We can then also measure the angles between each vector and their respective projections (the angles between a grey solid line to a coloured letter and the corresponding coloured vector). Note that if this angle is equal to 90 degrees, the projection is equal to 0 (point "c") or if that angle is > 90 degrees, the projection is negative (point "b").

We measured the exploration and elaboration using two main approaches:
 * By clades (i.e exploration/elaboration **between** clades): where we calculated the projections of the major axes of each clade onto the global phylogenetic major axes.
 * By species (i.e exploration/elaboration **within** clades): where we calculated the projections of each species onto 1) the global phylogenetic major axes and 2) the phylogenetic major axes of their respective clade.

## Elaboration and exploration by clade

To run the analyses for each clade we compared the major axes for each clade to the major axes for the overall phylogenetic effect (e.g. plovers' VCV vs. all birds' VCV, etc.).

For each posterior of the `mcmcmcglmmm` model, we projected each group's major axes onto the global phylogenetic major axis of the specific posterior sample.
In practice, for the projection to work, we translated and centred each group's major axes onto the origin of the global phylogenetic major axis (so that both origins have the coordinates 0) and then measure the projection and rejection as specified in Details 2.

```{r, fig.height = 6, fig.width = 18, echo = FALSE}
## Run the analyses between groups
results_between <- covar.projections.wrapper(covar_matrices,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Plot the results
wrap.plot <- function(results, col, main) {
    ## How many sub plots?
    n <- length(results)
    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    par(mfrow = c(1, n))
    for(i in 1:n) {
        plot(results[[i]], col = col, main = ifelse(i == centre, main, ""))
    }
}
wrap.plot(results_between,
          col = colour_vector[-4],
          main = "Groups difference")
```
Figure: these results show the projections ("position"), rejection ("distance") and angles ("degree") of the posteriors of each group's major axes onto the corresponding global phylogenetic major axes. For each panel, the highest the score the more the clade is elaborating, exploring, exploring perpendicular to the main axes for the position, distance and degree respectively.

## Elaboration and exploration by species

Similarly to the analyses described above, we ran the analyses for each species.
Here, instead of projecting the major axes of the groups onto the global phylogenetic major axis we projected each species onto the phylogenetic major axis (for each posterior distribution).

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses of groups on the phylogeny
results_phylo <- covar.projections.wrapper(covar_matrices,
                                            type = "within",
                                            base = "phylogeny",
                                            dispRity.out = TRUE)
## Plot the results
wrap.plot(results_phylo,
          col = colour_vector[-4],
          main = "Projection on the\nglobal phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of global major phylogenetic axis.

Finally, we ran the same analyses but instead of the projection of the species onto the global phylogenetic major axes, we projected them onto their respective group's phylogenetic major axes.

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses groups on themselves
results_within <- covar.projections.wrapper(covar_matrices,
                                            type = "within",
                                            dispRity.out = TRUE)
## Plot the results
wrap.plot(results_within,
          col = colour_vector,
          main = "Projection on the\ngroup's phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of each clade's major phylogenetic axis (the grey distributions corresponds to the projections of all the species, irregardless of their group, onto the global phylogenetic major axes).

```{r, echo = FALSE, eval = TRUE}
library(phytools)
source("~/Packaging/phytools/R/plotTree.wBars.R")

get.disparity.matrix <- function(data, col) {
    ## Get the matrices
    matrices <- lapply(data$disparity, `[[`, 1)
    ## Add the rownames
    list_of_names <- list()
    for(one_group in 1:length(matrices)) {
        rownames(matrices[[one_group]]) <- list_of_names[[one_group]] <- rownames(data$matrix[[1]])[c(data$subsets[[one_group]]$elements)]
    }
    ## Name the groups
    names(list_of_names) <- names(data$subsets)

    ## Matrix
    matrix <- do.call(rbind, matrices)

    ## Colours
    if(!missing(col)) {
        name.cols <- function(names, col) {
            out <- rep(col, length(names))
            names(out) <- names
            return(out)
        }
        col <- mapply(name.cols, list_of_names, as.list(col))
        names(col) <- NULL
        col <- unlist(col)
    }

    ## Output
    if(missing(col)) {
        return(list(data = matrix, names = list_of_names))
    } else {
        return(list(data = matrix, names = list_of_names, col = col))
    }
}

## Get the different elements
phylo_elaboration  <- get.disparity.matrix(results_phylo$position,
                                          col = colour_vector[-4])
within_elaboration <- get.disparity.matrix(
    get.subsets(results_within$position, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
phylo_exploration  <- get.disparity.matrix(results_phylo$distance,
                                          col = colour_vector[-4])
within_exploration <- get.disparity.matrix(
    get.subsets(results_within$distance, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
phylo_angle        <- get.disparity.matrix(results_phylo$degree,
                                          col = colour_vector[-4])
within_angle       <- get.disparity.matrix(
    get.subsets(results_within$degree, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
```

## Projections (on the main phylogeny)

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, phylo_elaboration$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Elaboration", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, phylo_exploration$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Exploration", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, phylo_angle$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Angle", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

## Projections (onto each clade's phylogeny)

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, within_elaboration$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Elaboration (within)", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, within_exploration$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Exploration (within)", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, within_angle$data, type = "phylogram", col = phylo_elaboration$col, tip.label = TRUE, main = "Angle (within)", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.scale = TRUE))
```

These graphs can also be plotted as circular for the big trees later.

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot per species
plotTree.wBars_modif(demo_tree, within_angle$data, type = "fan", col = phylo_elaboration$col, tip.label = FALSE, main = "Angle (within)", mar = c(3, 0.1, 0.1, 0.1), args.distribution = list(show.grid.scale = TRUE))
```

## Color PCA according to mean proj/rej?

And there's also this way of looking at it. But bof. Kinda ugly and probably too much info. (also it seems something is not working correctly).

```{r,echo = FALSE, eval = TRUE}
## Plotting fun
plot.gradient.space <- function(data, values, cent.tend = median, col, gradient.fun = heat.colors, ...) {

    get.colour.gradient <- function(values, col.fun, col.range, get.range = FALSE) {

        ## Sort the data by range
        histo <- hist(c(col.range, values), plot = FALSE)
        n_col <- length(histo$counts)

        ## Get the gradient
        avail_cols <- rev(col.fun(n_col))

        ## Attribute the colours
        col_out <- rep(avail_cols[n_col], length(values))
        range <- avail_cols[c(n_col, 1)]

        ## Attribute the colour by range
        for(colour in rev(1:n_col)) {
          col_out[(values <= histo$breaks[colour])] <- avail_cols[colour]
        }

        if(get.range) {
            return(list(range = range, col = col_out))
        } else {
            return(col_out)
        }
    }

    ## Get the colour gradient
    gradient_values <- apply(values, 1, cent.tend)
    gradient_cols <- get.colour.gradient(gradient_values, col.fun = gradient.fun, col.range = range(gradient_values), get.range = TRUE)
    colour_gradient <- gradient_cols$col
    names(colour_gradient) <- rownames(values)

    ## Plot the points
    sauron.plot(data,
                col = col,
                major.axes = mean,
                ellipses = mean,
                legend = TRUE,
                points = TRUE,
                pch = 19,
                cex = 1.5
                , ...)
    points(data$matrix[[1]], pch = 19, col = colour_gradient)
    legend("bottomright", pch = 19, col = gradient_cols$range, legend = round(range(gradient_values), 2))
}
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = phylo_elaboration$data, col = colour_vector, main = "Elaboration gradient")
```
```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = phylo_exploration$data, col = colour_vector, main = "Exploration gradient")
```
```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = phylo_angle$data, col = colour_vector, main = "Angle gradient")
```

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = within_elaboration$data, col = colour_vector, main = "Elaboration gradient (within)")
```
```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = within_exploration$data, col = colour_vector, main = "Exploration gradient (within)")
```
```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
plot.gradient.space(data = covar_matrices, values = within_angle$data, col = colour_vector, main = "Angle gradient (within)")
```

# Supplementary details

## Details 1: Major axes calculation

For $O_{n}$, the unit hypersphere matrix of _n_ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation}

In other words $O_{n}$ is the matrix representing each edge point of the unit hypersphere of $n$ dimensions and of radius $1$;

And $O'_{n}$ is the scaled matrix hypersphere to the 95\% confidence interval size using the $\chi^2$ distribution:

\begin{equation}
    O'_{n} = O_{n} \sqrt{\chi^2(0.95)}
\end{equation}

For the variance-covariance matrix $VCV_{n}$ of $n$ dimensions

\begin{equation}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation}

and the eigenvectors *v* and the eigenvalues $\lambda$ satisfying the following eigen decomposition:

\begin{equation}
    VCV_{n} v = \lambda v
\end{equation}

We can get $M_{n}$ matrix containing all the edge coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors *v* and the product of the scaled 0.95 CI unit sphere $O'_{n}$ and the eigenvalues $\lambda$:

\begin{equation}
    M_{n} = [(O'_{n}\sqrt{\lambda}) \times v]^{\text{T}}
\end{equation}

Where $M_{1,n}$ is the major axis of the 0.95 CI hyper-ellipse fitting the variance-covariance matrix, $M_{2,n}$ the second axis (minor axis in 2D), etc. etc.

Finally, we centred the matrix $M_{1,m}$ on the centroid of each clade.

> The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `beer::get.axes` ([specific procedure on lines 40-73](https://github.com/TGuillerme/elaboration_exploration_bird_beaks/blob/master/beer/R/get.axes.R)).

## Details 2: Measuring projection and rejection

For any vectors $\vec{a}$ and $\vec{b}$, defined either as one set or a pair of sets of coordinates in $n$ dimensions:

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x \\
    y \\
    \cdots \\
    n \\
    \end{bmatrix}
\end{equation}

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x_{1} & x_{2} \\
    y_{1} & y_{2} \\
    \cdots & \cdots \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

We can calculate $\vec{a_{1}}$, the orthogonal projection of $\vec{a}$ onto $\vec{b}$ using:

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\vec{b}\|}
\end{equation}

With $\|\vec{b}\| = \sqrt{\vec{b} \cdot \vec{b}}$ being the norm of $\vec{b}$.
And $\vec{a_{2}}$, the rejection of $\vec{a}$ onto $\vec{b}$:

\begin{equation}
    \vec{a_{2}} = \vec{a} - \vec{a_{1}}
\end{equation}

### Generalisation of projection onto any vector in a set space

Using this, we can calculate the projection and rejection for any element within a trait space $TS_{m,n}$:

\begin{equation}
    TS_{m,n} = 
    \begin{bmatrix}
    x_{1} & x_{2} & \cdots & x_{m} \\
    y_{1} & y_{2} & \cdots & y_{m} \\
    \vdots  & \vdots  & \ddots & \vdots \\
    n_{1} & n_{2} & \cdots & n_{m} \\
    \end{bmatrix}
\end{equation}

And any base vector $\vec{B}$ defined as:

\begin{equation}
    B = 
    \begin{bmatrix}
    x_{1} & x_{2}\\
    y_{1} & y_{2}\\
    \vdots  & \vdots  \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

By using the linear transformation $f_{\vec{B}}$ of the trait space $TS$ moving $\vec{B}$ onto $TS$'s first axes unit vector $\vec{\hat{\imath}}$:

\begin{equation}
    f_{\vec{B}}(TS) = \left( \frac{TS - [Bx_{1}, By_{1}, \cdots, Bn_{1}]^{\text{T}}}{\|\vec{B}\|} \right) \cdot R_{\vec{B}}
\end{equation}

With $R_{\vec{B}}$ being the rotation matrix of the vector $\vec{B}$ onto $\vec{\hat{\imath}}$:

\begin{equation}
R_{\vec{B}} = I_{\vec{B}} - \vec{B}\vec{B}^\text{T} - \vec{\hat{\imath}}\vec{\hat{\imath}}^\text{T} + [\vec{B} \vec{\hat{\imath}}]     \begin{bmatrix}
        cos(\theta) & -sin(\theta)\\
        sin(\theta) & cos(\theta)\\
    \end{bmatrix} [\vec{B} \vec{\hat{\imath}}]^\text{T}
\end{equation}

Where $\theta$ is:

\begin{equation}
    \theta = acos \left(\frac{\vec{B} \cdot \vec{\hat{\imath}}}{\|\vec{B}\| \cdot \|\vec{\hat{\imath}}\|} \right)
\end{equation}

Or $\theta = acos (B_x)$ since both $\|\vec{B}\|$ and $\|\vec{\hat{\imath}}\|$ are equal to 1 and $\|\vec{\hat{\imath}}\|$ is the unit vector on the first axis.

#### Algorithm for calculating the projection/rejection of any element in a defined space

In practice we followed [this procedure](https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions) and applied a modification of [this implementation](https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385) (see @aguilera2004 for the formal generalisation of this algorithm) using the following algorithm implemented in `dispRity::projections` (@disprity):

 1. In the trait space, define $\vec{B}$ as the base vector (typically $\vec{B}$ is defined as the pair of coordinates set from the major axis described above).
 2. Centre the trait space on the origin of $\vec{B}$ so that the first set of coordinates of $\vec{B}$ are 0.
 3. Scale the trait space to the norm of $\vec{B}$ so that the norm of $\vec{B}$ is now 1.
 4. Rotate the trait space using the rotation matrix $R_{\vec{B}}$ to satisfy the linear transformation $\vec{B} \arrow \vec{\hat{\imath}}$ (with $\vec{\hat{\imath}}$ being the first unit vector of the trait space - typically the x axes unit vector). 
 5. Project/reject every element in the trait on $\vec{B}$ (that is now $\vec{\hat{\imath}}$). In practice, the first coordinate (x) of each element is now its projection onto $\vec{B}$.

## References