### 2021-11-01 Analysis/02.1-shapespace_allbirds_lvl_superorder_order.Rmd
removed unused data

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script is a working script to run the chains on the SHARC Sheffield cluster.
Refer to `02-MCMCglmm_mini_chains.Rmd` for more details and reproducibility

```{r load_packages, eval = TRUE, message = FALSE}
set.seed(42)
## Loading beer
library(mcmcmcglmmm)
library(MCMCglmm)
library(dispRity)

## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")

## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
selected_dim <- shapespace_allbirds_lvl_superorder_order[[1]]$dimensions
consensus_tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
trees_list <- shapespace_allbirds_lvl_superorder_order[[1]]$trees_list
```

# All birds shapespace analyses

## Mini-chains parametrization

```{r, parametrise_the_model, eval = FALSE}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = shapespace,
                                   dimensions   = selected_dim,
                                   tree         = consensus_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "level1", "level2"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

## Running the three MCMCglmm models
100k_params_shapespace_8D_31levels <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(param_chains, file = "../Data/Processed/MCMCglmm/100k_params_shapespace_8D_31levels_chain1.rda")
```

From these three chains, we first extracted the three burnin periods (defined as the generation when the chain reaches the median likelihood value times 1.25).
We then extracted the posterior estimates of the R-Structure, G-Structure and fixed effects (mean and covariance) (ignoring the runs from the previously estimated burnin periods) to serve as our priors for our mini-chains. 
For those values, we took the highest burnin as the overall burnin and the median of the three posteriors as the overall priors.

```{r, get_model_parameters, eval = FALSE}
## Load the computed models
param_chains <- list()
for(i in 1:3) {
    load(paste0("../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/shapespace_allbirds_lvl_superorder_order_param", i, ".rda"))
    param_chains[[i]] <- shapespace_allbirds_lvl_superorder_order_param[[1]]
}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

### Running the mini-chains

```{r, set_the_mini_chains, eval = TRUE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Set up the mini-chains models
shapespace_allbirds_lvl_superorder_order <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1", "level2"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order.mini.chains")
```

The chains are then run on the cluster
TODO: details

### Combining the model data

```{r}
## Preparing the model files
model_list <- list()
class(model_list) <- c("beer", "mini.chains")
models_path <- "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/cluster/"
## Getting all the model files
model_files <- list.files(models_path, pattern = ".rda")
## Get the first model
load(paste0(models_path, model_files[1], collapse = "/"))
print(model_files[1])
## Clean the bits that are not going to be used (but are huge)
model[[1]][c("error.term", "family", "y.additional", "ginverse", "ZR")] <- NULL
## Update the model list
model_list[[length(model_list)+1]] <- model[[1]]
model <- NULL
model_files <- model_files[-1]

## Get the subsequent models and merge them as we go
while(length(model_files) != 0) {
    load(paste0(models_path, model_files[1], collapse = "/"))
    print(model_files[1])
    ## Clean the bits that are not going to be used (but are huge)
    model[[1]][c("error.term", "family", "y.additional", "ginverse", "ZR")] <- NULL
    ## Update the model list
    model_list[[length(model_list)+1]] <- model[[1]]
    ## Merge the models (overwrite the model list)
    model_list <- list(combine.mini.chains(model_list))
    ## DEBUG
    print("model list size:")
    print(format(object.size(model_list), units = "MB"))
    model <- NULL
    model_files <- model_files[-1]
}

## Combining the model files
shapespace_allbirds_lvl_superorder_order <- model_list[[1]]

## Checking the diagnosis

## Saving the results
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")

```