### 2022-02-24 Analysis/03-projection_analysis.Rmd

---
title: "Projection analysis to measure biological elaboration and innovation."
author: "Thomas Guillerme, Natalie Cooper, Andrew P Beckerman,  Gavin H Thomas"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 8
    fig_height: 8
    keep_tex: true
  self_contained: true
---

# Projection analyses

With the variance-covariance matrices obtained from the `mcmcmcglmmm` analyses, we can use linear algebra to make a mathematical interpretation of @endler2005's elaboration and innovation (here exploration) patterns (Fig 1. in @endler2005).
In fact, if we can interpret the variance-covariance matrix as estimated evolutionary space in the trait space that contains all the observed traits of a group and takes evolutionary history into account, we can then use the first eigen vector of that variance-covariance matrix as the major evolutionary axis, i.e. the axis in multidimensional space along which most trait values co-evolve.
In other words, if the variance-covariance matrix is a multidimensional ellipse, it's major axis is its major evolutionary axis.
Then, if species evolve parallel to this axis, they are elaborating on traits combinations; conversely, if species evolve orthogonaly to this axis, they are exploring new traits combinations (innovating).

Mathematically we can use linear algebra to project and reject each species in the trait-space along this major axis of evolution (Figure 4).
The detailed operations are described in the [supplementary materials](#projections-supplementary).

```{r, fig.height = 6, fig.width = 6, echo = FALSE, eval = TRUE}
source("../Functions/test.rotation.R")
```

##### Figure 3: {#fig3}
Consider a space with five elements named \"a\" to \"e\" (in grey) and one focal axes in this space defined as the vector from point1: \"d\" to point2: \"e\" (the thin solid grey line linking the grey \"d\" and \"e\"). We can then define all points from \"d\" (point1) as vectors (the dashed grey lines). We can then rotate the whole space so that the vector \"d\" to \"e\" becomes the reference coordinate vector (0,1) (the thick grey line) and rotate and translate all the other vectors accordingly (the solid grey lines linking the coloured letters \"a\" to \"c\"). We can then project each of these vectors on the reference vector and measure their distance *on* the reference vector (the projection; the coloured solid lines) and their distance *from* the reference vector (the rejection; the coloured dotted lines). We can then also measure the angles between each vector and their respective projections (the angles between a grey solid line to a coloured letter and the corresponding coloured vector). Note that if this angle is equal to 90 degrees, the projection is equal to 0 (point \"c\") or if that angle is > 90 degrees, the projection is negative (point \"b\").

In practice, we can use these projections and rejections of vectors or points onto a specific other vector as a way to measure the elaboration and exploration species within a group or of entire groups.
The first instance, measuring the elaboration and exploration of species within a group, is rather straightforward from [Figure 3](#fig3): the idea is to project each points (defined as the dashed vectors in figure 4) onto a vector of interest (e.g. either the group's own main axis of variance-covariance or the entire phylogeny's one).

We measured the exploration and elaboration using two main approaches (see [Figure 4](#fig4)):

 * By clades (i.e exploration/elaboration **between** clades), where we calculated the projections of the major axes of each clade onto the global phylogenetic major axes.
 * By species (i.e exploration/elaboration **within** clades), where we calculated the projections of each species onto 1) the global phylogenetic major axes and 2) the phylogenetic major axes of their respective clade.

```{r, echo = FALSE, eval = TRUE}
## Get projection lines (from a centred matrix)
get.proj.line <- function(point, base) {
  point <- t(point)
  base <- t(base)
  proj_x <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[1];
  proj_y <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[2];
  return(rbind(cbind(proj_x, proj_y), point))
}

set.seed(42)
## Space plot
cor_matrix <- matrix(cbind(1, 0.8, 0.8, 1), nrow = 2)
space <- space.maker(10, 2, rnorm, cor.matrix = cor_matrix)
lim <- c(floor(range(space)[1]), ceiling(range(space)[2]))

par(mfrow = c(2, 2))

## Plot original space
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim, col = "black", main = "Original space")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "blue", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "blue", lty = 1)
## Plotting a phylogenetic major axis
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "orange")
# legend("topright", pch = c(19, NA, NA, NA), lty = c(NA, 1,1,1), lwd = c(NA, 1, 3, 3), legend = c("Group of species", "95% CI ellipse for the group", "Major axis for the group", "Major axis for the phylogeny"), col = c("black", "blue", "blue", "orange"))
legend("bottomright", legend = "A")

## Projections of the groups
## Translate the major axis
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim, col = "grey", main = "Groups' projection\non the phylogeny")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "grey", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "grey", lty = 1)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "orange")
legend("bottomright", legend = "B")

## Do the translation
major_axis <- get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75)
point_to_translate <- c(-1.2, 0)
# lines(rbind(major_axis[2,], point_to_translate), lty = 3, col = "grey")
arrows(x0 = major_axis[2,1], y0 = major_axis[2,2], x1 = point_to_translate[1], y1 = point_to_translate[2], col = "black", lty = 3, length = 0.1)

move_x <- point_to_translate[1] - major_axis[2, 1]
move_y <- point_to_translate[2] - major_axis[2, 2]
axis_trans <- major_axis
axis_trans[,1] <- axis_trans[,1] + move_x
axis_trans[,2] <- axis_trans[,2] + move_y
lines(axis_trans, lwd = 3, col = "blue", lty = 1)

## Project the translation
matrix <- axis_trans
point1 <- c(-1.2, 0)
point2 <- c(2, 0.75)
base_vector <- rbind(point1, point2)
## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]
## Projected translation
cent_proj <- get.proj.line(matrix_cent[1,], base_vector_cent[2, ])
cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
lines(cent_proj, col = "darkgreen")
## Extend the phylogeny line
extend_base <- rbind(base_vector[1, ], cent_proj[1, ])
lines(extend_base, lwd = 1, col = "orange", lty = 2)

## Projections per species
## Translate the major axis
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim, col = "grey", main = "Species' projection\non the phylogeny")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "grey", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "grey", lty = 1)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "orange")
legend("bottomright", legend = "C")

## Do the projection
matrix <- space
point1 <- c(-1.2, 0)
point2 <- c(2, 0.75)
base_vector <- rbind(point1, point2)

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]

## Plot the projections
for(i in 1:nrow(matrix)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  lines(cent_proj, col = "darkgreen")
}

## Species projection onto their own axis
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim, col = "grey", main = "Species' projection\non their own axis")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "grey")
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "blue", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "blue", lty = 1)
legend("bottomright", legend = "D")

## Do the projection
matrix <- space
base_vector <- get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75)
point1 <- base_vector[1,]
point2 <- base_vector[2,]

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]
## Extend the base vector a little bit
extend_base <- base_vector
extend_base[1, ] <- extend_base[1,1] + 0.05
lines(extend_base, lwd = 1, col = "blue", lty = 2)

## Plot the projections
for(i in 1:nrow(matrix_cent)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  cent_proj[, 2] <- (cent_proj[, 2] + point1[2])
  lines(cent_proj, col = "darkgreen")
}

```

##### Figure 4: {#fig4}
Different types of elaboration exploration analyses for a group of species. A) Given a group of 10 species (in black) with two traits, we can calculate the 95% confidence interval variance-covariance ellipse and its major axis (in blue). Say we also calculated another vector (in orange) that could be for example the major axis of another variance-covariance matrix (e.g. from another group or the whole phylogeny). B) we can calculate the projection of the group's major axis on the other major axis by translating it (black doted arrow) and calculate it's elaboration relative to the orange vector (orange dashed line) or its relative exploration (green line). C) We can also calculate the elaboration and exploration scores for each individual species relative to the orange major axis (e.g. the phylognetic major axis); D) or relative to their own major axis.

<!-- Scaling and interpretation -->
Note that the examples depicted in [Figure 4](#fig4) are not scaled for clarity.
In the algorithm implemented in `dispRity`, the results are always effectively scaled as described in [Figure 3](#fig3) and in the [detailed algorithm](#projections-supplementary).
In practice, this makes the vector onto which elements are projected of length one.
Furthermore, because the projections happen based on the first point of the major axis of interest (which can be arbitrary), we advice to centre the elaboration results on 0.5 and rescaled (effectively removing 0.5 from the scaled projection value and rescaling it to 1) and to make them absolute.
This makes the elaboration and exploration results more intuitive to interpret: an elaboration score comprised in $[0, 1]$ is distributed onto the major axis, a score of $[1, \infty)$ is beyond the major axis (in any _n_ direction).
If the major axis is for example the 95% confidence interval from the phylogenetic variance-covariance matrix, this denotes a species that is especially elaborative.
Similarly for exploration, any exploration score comprised in $[0, 1]$ indicates an exploration that is within the major axis' variance-covariance 95% confidence interval; a score of $[1, \infty)$ denotes a particularly explorative species.

# Reproducible example

Here we illustrate the BEER pipeline by applying it to a data set from @cooney2017 focused on the birds charadriiformes order.

## Data

This order contains 359 species divided into three clades: gulls, sandpipers and plovers and sandpipers that have respectively 159, 102 and 98 species each.
For each of these species, we used the 3D beak dataset from @cooney2017 that is an ordination of the shapes of the beaks of 8@@@ birds (the resulting shape space used here does not contain any information about the beak size - i.e. centroid size).
We extracted the 359 charadriiformes species from this space where the first three dimensions account for more than 99% of the variance in the dataset.
For the rest of the example analysis here, we used as a traitspace of these 359 species with 3 dimensions (hereafter the shapespace).

```{r, echo = FALSE, eval = TRUE}
## Loading the charadriiformes data
data(charadriiformes) 
## Extracting the tree
tree <- charadriiformes$tree
## Extracting the data column that contains the clade assignments
data <- charadriiformes$data[, "clade"]
## Changing the levels names (the clade names) to colours
levels(data) <- c("orange", "blue", "darkgreen")
data <- as.character(data)
## Matching the data rownames to the tip order in the tree
data <- data[match(ladderize(tree)$tip.label, rownames(charadriiformes$data))]

## Matching the tip colours (labels) to their descending edges in the tree
## (and making the non-match edges grey)
clade_edges <- match.tip.edge(data, tree, replace.na = "grey")

## Plotting the results
plot(ladderize(tree), show.tip.label = FALSE, edge.color = clade_edges, main = "Charadriiformes")
legend("bottomleft", lty = c(1,1,1), col = c("blue", "darkgreen", "orange"), legend = c("plovers", "sandpipers", "gulls"))
axisPhylo()
```

##### Figure 5: {#fig5}
The Charadriiformes data used in this pipeline illustration. The axis units are millions of years ago.

## Projections analysis

We then used these results to understand measure which of the three charadriiformes is the most elaborative and the most exploring.
We tested that both at 1) the entire clade level, by projecting the 1000 posterior major axis of each clade's phylogenetic residuals onto the global phylogeny's residuals; and 2) the species level to see which clade presented the most elaborative and exploring species.

```{r, echo = FALSE, eval = TRUE}
covar_data <- MCMCglmm.subsets(data = charadriiformes$data,
                               posteriors = posteriors,
                               group = MCMCglmm.levels(posteriors)[1:4],
                               rename.groups = c("gulls", "plovers", "sandpipers", "phylogeny"))
par(mfrow = c(2,2))
covar.plot(covar_data, points = TRUE, points.cex = 0.5, col = c("orange", "blue", "darkgreen"))
legend("topleft", legend = "A")
covar.plot(covar_data, points = TRUE, points.cex = 0.5, ellipses = mean, major.axes = mean, legend = TRUE, col = c("orange", "blue", "darkgreen", "grey"), apply.to.VCV = TRUE, legend.x = "bottomright")
legend("topleft", legend = "B")
covar.plot(covar_data, points = FALSE, ellipses = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"))
legend("topleft", legend = "C")
covar.plot(covar_data, points = FALSE, major.axes = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"))
legend("topleft", legend = "D")
```

##### Figure 6: {#fig.6}
Visualisation of the posteriors of the model: A) the first two dimensions of the inpu trait space; B) the same space with the average ellipses and average major.axis plotted; C) the distribution of 100 random ellipses; D) the distribution of 100 random major axes.")

```{r, echo = FALSE, eval = FALSE}
## Measure the group differences
group_differences <- dispRity.covar.projections(covar_data,
                                                type = "groups",
                                                base = "phylogeny",
                                                output = c("position", "distance"))
## Measure the individual differences
species_differences <- dispRity.covar.projections(covar_data,
                                                  type = "elements",
                                                  base = "phylogeny",
                                                  output = c("position", "distance"))

```

# References

# Supplementary details {#projections-supplementary}

## Detailed projection and rejection operations

In details, we can define the major axes from the variance-covariance matrices and then project and reject each element of interest in the space onto this axis.
The following steps are generalised to $n$ dimensions and detailed below (as well as the algorithm used in `dispRity` to perform the transformations:

### Major axes definition

For $O_{n}$, the unit hypersphere matrix of $n$ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation}

In other words $O_{n}$ is the matrix representing the hypersphere of $n$ dimensions and of radius $1$ that fits in the centre of the trait space;

And $O'_{n}$ is the scaled matrix hypersphere to the 95\% confidence interval size using the $\chi^2$ distribution:

$$O'_{n} = O_{n} \sqrt{\chi^2(0.95)}$$

Then, for the variance-covariance matrix $VCV_{n}$ of $n$ dimensions obtained from the posterior distribution of the `mcmcmcglmmmm`:

\begin{equation}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation}

and the eigenvectors *v* and the eigenvalues $\lambda$ satisfying the following eigen decomposition:

$$VCV_{n} v = \lambda v$$

We can get $M_{n}$, the matrix containing all the edge coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors *v* and the product of the scaled 0.95 CI unit sphere $O'_{n}$ and the eigenvalues $\lambda$:

$$M_{n} = [(O'_{n}\sqrt{\lambda}) \times v]^{\text{T}}$$

Finally, we can centre the matrix $M_{n}$ on the estimated solution of each generalised linear mixed model (`model$Sol`, in `MCMCglmm`) corresponding to the estimation of the position of the variance-covariance matrix in the trait space.
$M_{n}$ then contains all the major axis of the 0.95 hyper-ellipse fitting the variance-covariance matrix.
We can then define the first row of the matrix, $M_{1,n}$, as the major axis, the second row, $M_{2,n}$, as the second major axis (the minor axis in a 2D ellipse), etc.

The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `dispRity::axis.covar`.

### Measuring projection and rejection

Once we have defined a major axis, we can project any elements in the trait space onto that axis.
Specifically, for any elements in the trait space, we can define it as the vector $\vec{a}$ with one set of coordinates in $n$ dimensions:

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x \\
    y \\
    \cdots \\
    n \\
    \end{bmatrix}
\end{equation}

And for any major axis that we can define as a vector $\vec{b}$ as a set of pairs of coordinates in $n$ dimensions:

\begin{equation}
    \vec{b} = 
    \begin{bmatrix}
    x_{1} & x_{2} \\
    y_{1} & y_{2} \\
    \cdots & \cdots \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

We can then calculate $\vec{a_{1}}$, the orthogonal projection of $\vec{a}$ onto $\vec{b}$ using:

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\vec{b}\|}
\end{equation}

With $\|\vec{b}\| = \sqrt{\vec{b} \cdot \vec{b}}$ being the norm of $\vec{b}$.
And $\vec{a_{2}}$, the rejection of $\vec{a}$ onto $\vec{b}$:

\begin{equation}
    \vec{a_{2}} = \vec{a} - \vec{a_{1}}
\end{equation}

#### Generalisation of projection onto any vector in a set space

Using this, we can generalise the procedure so as to calculate the projection and rejection for any element within a trait space $TS_{m,n}$:

\begin{equation}
    TS_{m,n} = 
    \begin{bmatrix}
    x_{1} & x_{2} & \cdots & x_{m} \\
    y_{1} & y_{2} & \cdots & y_{m} \\
    \vdots  & \vdots  & \ddots & \vdots \\
    n_{1} & n_{2} & \cdots & n_{m} \\
    \end{bmatrix}
\end{equation}

And any major axis defined as a vector $\vec{B}$:

\begin{equation}
    B = 
    \begin{bmatrix}
    x_{1} & x_{2}\\
    y_{1} & y_{2}\\
    \vdots  & \vdots  \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

By using the linear transformation $f_{\vec{B}}$ of the trait space $TS$ moving $\vec{B}$ onto $TS$'s first axes unit vector $\vec{\hat{\imath}}$:

$$f_{\vec{B}}(TS) = \left( \frac{TS - [Bx_{1}, By_{1}, \cdots, Bn_{1}]^{\text{T}}}{\|\vec{B}\|} \right) \cdot R_{\vec{B}}$$

With $R_{\vec{B}}$ being the rotation matrix of the vector $\vec{B}$ onto $\vec{\hat{\imath}}$:

\begin{equation}
R_{\vec{B}} = I_{\vec{B}} - \vec{B}\vec{B}^\text{T} - \vec{\hat{\imath}}\vec{\hat{\imath}}^\text{T} + [\vec{B} \vec{\hat{\imath}}]
    \begin{bmatrix}
        cos(\theta) & -sin(\theta)\\
        sin(\theta) & cos(\theta)\\
    \end{bmatrix} [\vec{B} \vec{\hat{\imath}}]^\text{T}
\end{equation}

Where $\theta$ is:

\begin{equation}
    \theta = acos \left(\frac{\vec{B} \cdot \vec{\hat{\imath}}}{\|\vec{B}\| \cdot \|\vec{\hat{\imath}}\|} \right)
\end{equation}

Or $\theta = acos (B_x)$ since both $\|\vec{B}\|$ and $\|\vec{\hat{\imath}}\|$ are equal to 1 and $\|\vec{\hat{\imath}}\|$ is the unit vector on the first axis.

### Algorithm for calculating the projection/rejection of any element in a defined space

In practice we followed [this procedure](https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions) and applied a modification of [this implementation](https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385) (see @aguilera2004 for the formal generalisation of this algorithm in $n$ dimensions) using the following algorithm implemented in `dispRity::projections` (@disprity):

 1. In the trait space, define $\vec{B}$ as the base vector (typically $\vec{B}$ is defined as the pair of coordinates set from the major axis described above).
 2. Centre the trait space on the origin of $\vec{B}$ so that the first set of coordinates of $\vec{B}$ are 0.
 3. Scale the trait space to the norm of $\vec{B}$ so that the norm of $\vec{B}$ is now 1.
 4. Rotate the trait space using the rotation matrix $R_{\vec{B}}$ to satisfy the linear transformation $\vec{B} \rightarrow \vec{\hat{\imath}}$ (with $\vec{\hat{\imath}}$ being the first unit vector of the trait space - typically the x axes unit vector). 
 5. Project/reject every element in the trait on $\vec{B}$ (that is now $\vec{\hat{\imath}}$). In practice, the first coordinate (x) of each element is now its projection onto $\vec{B}$.