### 2022-03-10 Analysis/Older_ideas/04-full_analyses_run.Rmd

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script can be used to repeat the full analyses.
You can read about the methods overview [here](00-methods_overview.Rmd) or find details about the full analyses in the respective detailed descriptions of the [data](01-Data_preparation.Rmd), the [MCMCglmm mini chains](02-MCMCglmm_mini_chains.Rmd) and the [elaboration exploration analyses](03-elaboration_exploration_analyses.Rmd).
You can reproduce the data preparation, models estimations and calculations of the elaboration and exploration from the [publication](§§§pub) with the [running script](04-full_analyses_run.Rmd).

```{r, load pkg and data}
install.packages(c("devtools", "dispRity", "MCMCglmm"))
devtools::install_github("TGuillerme/elaboration_exploration_bird_beaks/beer")
library(beer)
library(MCMCglmm)
data(chara_data)
data(chara_tree)
```

## Check the following options:

 * saving parameters
 * number of dimensions
 * numbers of iterations (especially rate)
 * check options in brms

```{r, run param chains}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:17),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 5000000,
                                       thin   = 5000,
                                       burnin = 0))

## Running the three MCMCglmm models
param_chains <- run.mini.chains(param_MCMCglmm, replicates = 3)
save(param_chains, file = "param_chains.rda")
```

```{r, get params}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

```{r, set_the_mini_chains, eval = TRUE}
## The thinning parameter
thin <- 1000
## The number of samples per chains
samples <- 100

## Set up the mini-chains models
mini_chains <- make.mini.chains(data         = chara_data[, 1:17],
                                dimensions   = c(1:17),
                                tree         = chara_tree,
                                trait.family = "gaussian",
                                randoms      = c("global", "clade"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
```

```{r, run the mini_chains}
## The number of replicates
n_reps <- 1000

## Run the required chains
results <- run.mini.chains(mini_chains, replicates = n_reps, path = "../Data/Processed/MCMC/")

## Combining the results as one big MCMC
combined_results <- combine.mini.chains(results)

## Saving the results
save(combined_results, file = "../Data/processed/combined_results.rda")
```