### 2022-03-10 Analysis/05-Figure2.Rmd
Figure name files UP

---
title: "Bird beak's elaboration and innovation"
author: "Gavin Thomas"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

TODO!