### 2022-05-16 Analysis/06-Figure3_with_correlations.Rmd
sync

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This section contains the code for reproducing the analyses. To get the details, refer to 03-elaboration_innovation_analyses.Rmd.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
```

# All birds shapespace analyses

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Creating the subsets
superorder_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

order_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)
```

```{r, message = FALSE, warning = FALSE, fig.height = 8, fig.width = 8, echo = FALSE}
## The ggplot color panel
gg.color.hue <- function(n) {
    grDevices::hcl(h = seq(15, 375, length = n + 1), l = 65, c = 100)[1:n]
}

# colour.palette <- rainbow
colour.palette <- gg.color.hue
```

Loading the elaboration exploration results (from 03_results)

```{r, echo = FALSE}
## Quick load
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
results <- shapespace_allbirds_lvl_superorder_order_results_list
rm(shapespace_allbirds_lvl_superorder_order_results_list)
```

## Figure 3

Density plot of the correlation of the median elaboration and median exploration for all species.

```{r, echo = FALSE}
## Calculating all the correlations
get.correlations <- function(res, cent.tend = median, exclude.phylo = FALSE) {

    ## Remove the phylo one (if needed)
    if(exclude.phylo) {
        select_pos <- names(size.subsets(res$position))
        select_dis <- names(size.subsets(res$distance))
        select_pos <- select_pos[-which(select_pos %in% "phylogeny")]
        select_dis <- select_dis[-which(select_dis %in% "phylogeny")]
        res$position <- get.subsets(res$position, subsets = select_pos)
        res$distance <- get.subsets(res$distance, subsets = select_dis)
    }

    ## Extracting all the results
    disparity_pos <- lapply(res$position$disparity, function(X, type) return(X$elements))
    disparity_dis <- lapply(res$distance$disparity, function(X, type) return(X$elements))

    ## Measure the central tendencies
    centrals_pos <- lapply(disparity_pos, function(X, fun) apply(X, 1, fun), fun = cent.tend)
    centrals_dis <- lapply(disparity_dis, function(X, fun) apply(X, 1, fun), fun = cent.tend)

    ## Number of groups
    n_groups <- length(centrals_pos)

    ## Prepare the results
    prep.cor.results <- function(pos, dis, name) {
        ## Run the correlation test
        cor_test <- cor.test(pos, dis)
        ## Prepare the plot title
        plot_main <- paste0(name, ": ", unname(round(cor_test$estimate, 3)), " (p=", unname(round(cor_test$p.value, 3)), ")")
        ## Return the plotting parameters
        return(list(x = pos, y = dis, main = plot_main))
    }

    return(mapply(prep.cor.results, centrals_pos, centrals_dis, as.list(names(centrals_pos)), SIMPLIFY = FALSE))
}

## Plotting one correlation 
plot.correlations <- function(cor.results, col.sub, col.fun = colour.palette, legend.pos = "topleft", pt.cex = 0.5, legend.cex = 0.5, tree, full.tree = FALSE, ...) {
    dots <- list(...)

    if(missing(col.sub)) {
        col.sub <- NULL
    }

    ## Subbing the colours
    if(!is.null(col.sub)) {
        ## Reset the levels
        col_sub <- as.factor(as.character(col.sub))
        ## Make the empty level grey
        if(any(levels(col_sub) == "")) {
            col_avail <- c("grey", col.fun(length(levels(col_sub))-1))
        } else {
            col_avail <- col.fun(length(levels(col_sub)))
        }
        ## Create the colour vector
        col_vector <- col_avail[as.numeric(col_sub)]

    } else {
        if(is.null(dots$col)) {
            col_vector <- "black"
        } else {
            col_vector <- dots$col
        }
    }

    if(!missing(tree)) {
        op <- par(mfrow = c(1,2))
        #c(bottom, left, top, right); c(5, 4, 4, 2) + 0.1
        par(mar = c(5, 4, 4, 0)+0.1)   
    }

    ## Plotting the results
    plot(cor.results$x, cor.results$y, pch = 19, xlab = "elaboration", ylab = "innovation", main = cor.results$main, cex = pt.cex, col = col_vector)

    if(!is.null(col.sub)) {
        ## Adding the legend
        legend(x = legend.pos, legend = levels(col_sub), pch = 19, cex = legend.cex, col = col_avail)
    }

    if(!missing(tree)) {
        par(mar = c(5, 0, 4, 2) + 0.1)

        ## Recycle the colour vector
        named_col_vector <- col_vector
        names(named_col_vector) <- names(cor.results$x)

        ## Reduce the tree
        if(!full.tree) {
            tree <- drop.tip(tree, tip = tree$tip.label[which(!(tree$tip.label %in% names(named_col_vector)))])
        }

        ## Matching edges and colours
        match.edge.colours <- function(tree, named_col_vector, background = "grey") {
            ## Get the background colour
            edge_cols <- rep(background, Nedge(tree))
            ## Get the clades
            clades <- unique(named_col_vector)
            ## Ignore the greys
            if(any(which_grey <- clades == "grey")) {
                clades <- clades[!which_grey]
            }
            ## Find the edges for each clade
            for(clade in 1:length(clades)) {
                ## Get the tips for the clade
                tips <- which(named_col_vector == clades[clade])
                ## Get the MRCA node
                mrca_node <- getMRCA(tree, tips)
                # stop_edge <- which(tree$edge[, 2] %in% mrca_node)

                ## Handle non-monophyletic groups
                clade_tree <- extract.clade(tree, node = mrca_node)
                if(Ntip(clade_tree) > length(tips)) {
                    clade_tree <- drop.tip(clade_tree, tip = clade_tree$tip.label[!(clade_tree$tip.label %in% names(tips))])
                }
                n_edges_target <- Nedge(clade_tree)

                ## Get the tips ID
                tip_ids <- which(tree$tip.label %in% names(tips))
                ## Get the vector of edges
                edges <- integer()
                ## Get the first nodes (and edges)
                new_edges <- which(tree$edge[, 2] %in% tip_ids)
                edges <- unique(c(edges, new_edges))
                nodes <- tree$edge[new_edges, 1]

                # plot(tree, show.tip.label = FALSE)
                # test.col <- c(replicate(10, c("red", "blue", "orange", "green", "pink", "yellow")))
                # edgelabels(".", cex = 0.5, new_edges, bg = test.col[1]) ; test.col <- test.col[-1]

                ## Recursively loop through each node up until the last edge
                # while(!(stop_edge %in% edges)) {
                while(length(edges) <= n_edges_target) {
                    ## Get the node edges
                    new_edges <- which(tree$edge[, 2] %in% nodes)

                    # edgelabels(".", cex = 0.5, new_edges, bg = test.col[1]) ; test.col <- test.col[-1]

                    edges <- unique(c(edges, new_edges))
                    ## Update the nodes
                    nodes <- tree$edge[new_edges, 1]
                }

                # ## Do the last recursion
                # new_edges <- which(tree$edge[, 2] %in% nodes)
                # edges <- unique(c(edges, new_edges))

                # edgelabels(".", cex = 0.5, new_edges, bg = test.col[1]) ; test.col <- test.col[-1]

                ## Colour these edges
                edge_cols[edges] <- clades[clade]
            }
            return(edge_cols)
        }
        ## Match the colour to the correct edge
        edge_cols <- match.edge.colours(tree, named_col_vector)

        ## Add the tree
        plot(tree, show.tip.label = FALSE, edge.color = edge_cols)
    }

    return(invisible())
}
```

Correlations per super order (species on phylogeny):

```{r, eval = TRUE, echo = FALSE}
## Calculate the correlations for the projections onto the phylo axis
super_phylo_cor <- get.correlations(results$super_results_phylo)

## Calculate the correlations for the projections onto the group' axes
super_group_cor <- get.correlations(results$super_results_within, exclude.phylo = TRUE)
```

### Figure 3a - gg_version

Density plot of the correlation of the median elaboration and median exploration for all species.

```{r, echo = FALSE, eval = FALSE}
library(ggplot2)
## Getting the data to plot
global_data <- super_phylo_cor$phylogeny

gg.density.plot <- function(data, palette = "viridis") {
    ggplot(data.frame("elaboration" = data$x, "exploration" = data$y), aes(x=elaboration, y=exploration) ) +
      geom_hex() +
      scale_fill_continuous(type = palette) +
      theme_bw()
}

pdf("../Manuscript/Figures/Figure3_b.pdf", height = 4, width = 4)
gg.density.plot(global_data)
dev.off()
```

Can't seem to integrate the density plot to the base R layout so I'll just do it in Inkscape for now

```{r, echo = FALSE, eval = FALSE}
pdf("../Manuscript/Figures/Figure3_a.pdf", height = 6, width = 3)
## Plotting the projection cartoon
library(mcmcmcglmmm)
## Get projection lines (from a centred matrix)
get.proj.line <- function(point, base) {
  point <- t(point)
  base <- t(base)
  proj_x <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[1];
  proj_y <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[2];
  return(rbind(cbind(proj_x, proj_y), point))
}

set.seed(42)
## Space plot
cor_matrix <- matrix(cbind(1, 0.8, 0.8, 1), nrow = 2)
space <- space.maker(10, 2, rnorm, cor.matrix = cor_matrix)
lim <- c(floor(range(space)[1]), ceiling(range(space)[2]))

par(mfrow = c(2, 1))
## Projections per species
## Translate the major axis
par(bty = "n", mar = c(1,1,1,1)+0.1)
plot(space, pch = 19, xlab = "", ylab = "", xlim = lim, ylim = lim, col = "grey", main = "Panels B and C: species'\nprojection on the phylogeny", xaxt = "n", yaxt = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "grey", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "grey", lty = 1)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "orange")
# legend("bottomright", legend = "C")

## Do the projection
matrix <- space
point1 <- c(-1.2, 0)
point2 <- c(2, 0.75)
base_vector <- rbind(point1, point2)

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]

## Plot the projections
for(i in 1:nrow(matrix)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  lines(cent_proj, col = "darkgreen")
}

## Species projection onto their own axis
par(bty = "n", mar = c(1,1,1,1)+0.1)
plot(space, pch = 19, xlab = "", ylab = "", xlim = lim, ylim = lim, col = "grey", main = "Panels D: species'\nprojection on their own group", xaxt = "n", yaxt = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "grey")
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "blue", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "blue", lty = 1)
# legend("bottomright", legend = "D")

## Do the projection
matrix <- space
base_vector <- get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75)
point1 <- base_vector[1,]
point2 <- base_vector[2,]

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]
## Extend the base vector a little bit
extend_base <- base_vector
extend_base[1, ] <- extend_base[1,1] + 0.05
lines(extend_base, lwd = 1, col = "blue", lty = 2)

## Plot the projections
for(i in 1:nrow(matrix_cent)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  cent_proj[, 2] <- (cent_proj[, 2] + point1[2])
  lines(cent_proj, col = "darkgreen")
}
dev.off()
```

## Figure 3b

```{r, echo = FALSE, eval = FALSE}
## Loading the distances from centre for each species
load("../Data/Processed/ploting_data.rda")
centroids <- ploting_data$distances_to_centre[[1]]

pdf("../Manuscript/Figures/Figure3_cd.pdf", height = 12, width = 8)
## Great the plot layout
layout_matrix <- cbind(matrix(0,2,2), matrix(0,2,2))
layout_matrix <- rbind(rbind(layout_matrix, matrix(1:8, 2, 4, byrow = TRUE)), matrix(9:16, 2, 4,  byrow = TRUE))
plot_layout <- layout(layout_matrix)
# layout.show(plot_layout)
one.plot.correlations <- function(cor.results, col.sub, col.fun = colour.palette, legend.pos = "topleft", pt.cex = 0.5, legend.cex = 0.5, ID, legend_counter, contour = NULL) {

    if(missing(col.sub)) {
        col.sub <- NULL
    }

    ## Subbing the colours
    if(!is.null(col.sub)) {
        ## Reset the levels
        col_sub <- as.factor(as.character(col.sub))
        ## Make the empty level grey
        if(any(levels(col_sub) == "")) {
            col_avail <- c("grey", col.fun(length(levels(col_sub))-1))
        } else {
            col_avail <- col.fun(length(levels(col_sub)))
        }
        ## Create the colour vector
        col_vector <- col_avail[as.numeric(col_sub)]
    } else {
        if(is.null(dots$col)) {
            col_vector <- "black"
        } else {
            col_vector <- dots$col
        }
    }

    ## Sort the data by group size
    all_data <- cbind(x = cor.results$x, y = cor.results$y)
    ## Classify the groups per numbers of species
    all_data <- data.frame(all_data, col = col_vector)

    ## Sort the data per group size
    group_order <- sort(table(all_data$col), decreasing = TRUE)

    ## Set the plotting data
    plot_data <- data.frame()
    # col_adjust <- seq(from = 1/length(group_order), to = 1, by = 0.1)*3
    for(i in 1:length(group_order)) {
        ## Get the data from the biggest group
        selected_data <- all_data[which(all_data$col == names(group_order[i])), ]
        ## Transparantise the colours (for the background)
        # selected_data$col <- adjustcolor(selected_data$col, alpha = col_adjust[i])
        ## Combine the plotting data
        plot_data <- rbind(plot_data, selected_data)
    }

    ## Empty plot
    plot(NULL, xlim = range(plot_data$x), ylim = range(plot_data$y), xaxt = "n", yaxt = "n", xlab = "", ylab = "", main = strsplit(cor.results$main, split = ":")[[1]][1])
    axis(1, at = pretty(plot_data$x, n = 4))
    axis(2, at = pretty(plot_data$y, n = 4), las = 2)

    ## Adding contours
    if(!is.null(contour)) {
        ## Select the right contour data
        contour_data <- contour[match(rownames(plot_data), names(contour))]        
        ## Calculate the contour
        contour_z <- MASS::kde2d(contour_data/max(contour_data)*max(plot_data$x), contour_data/max(contour_data)*max(plot_data$y), n = 50)

        contour_z <- matrix(0, nrow = length(contour_data), ncol = length(contour_data))
        diag(contour_z) <-  contour_data
        # contour_z <- outer(contour_data/max(contour_data)*max(plot_data$x), contour_data/max(contour_data)*max(plot_data$y))

        ## Adding the contour in the background
        contour(contour_z, lwd = 1, add = TRUE, col = hcl.colors(12, palette = "viridis"), labels = pretty(contour_data))
    }

    # ## col test for contour
    # color.gradient <- function(x, colors=c("red","yellow","green"), colsteps=100) {
    #     return( colorRampPalette(colors) (colsteps) [ findInterval(x, seq(min(x),max(x), length.out=colsteps)) ] )
    # }

    ## Plot the points
    points(plot_data$x, plot_data$y, pch = 19, cex = pt.cex, col = plot_data$col)
    # points(plot_data$x, plot_data$y, pch = 19, cex = pt.cex, col = color.gradient(contour_data))

    if(!is.null(col.sub)) {
        ## See if there is any others
        levels <- levels(col_sub)
        if(length(other <- which(levels == "")) > 0) {
            legend_text <- c(1, 1:(length(levels(col_sub))-1)+legend_counter)
        } else {
            legend_text <- c(1:(length(levels(col_sub)))+legend_counter)
        }
        ## Update the legend counter
        legend_counter <<- max(legend_text)
        ## Adding the legend
        legend(x = legend.pos, legend = legend_text, pch = 19, cex = legend.cex, col = col_avail, bg = "white")
    }

    return(invisible())
}

# par(mfrow = c(2, 4))
## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))

## Set the vector of legend positions
legends_pos <- c("topleft", "topright", "topright", "topright",
                 "topleft", "topright", "topleft", "topright")
legend_counter <- 1

for(i in 1:8) {
    par(mar = c(2,2.5,2,1.5))
    one.plot.correlations(super_phylo_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i], ID = i, pt.cex = 0.5, legend.cex = 2/3, legend_counter = legend_counter, contour = centroids)
}

## Fig 3c

# par(mfrow = c(2, 4))
## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))

## Set the vector of legend positions
legends_pos <- c("bottomright", "topright", "topleft", "topleft",
                 "topleft", "topleft", "topright", "bottomright")
legend_counter <- 1

for(i in 1:8) {
    par(mar = c(2,2.5,2,1.5))
    one.plot.correlations(super_group_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i], ID = i, pt.cex = 0.5, legend.cex = 2/3, legend_counter = legend_counter)
}
dev.off()
```

### Correlations within super orders (species on group)

Regardless, here are the results when projecting the species on their own group's VCV

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12, eval = TRUE}
## Getting the subsets for the colouring
col_subs <- split(shapespace$level2, f = shapespace$level1)[-1]

## Calculate the correlations
super_within_cor <- get.correlations(results$super_results_within, exclude.phylo = TRUE)
## Set the vector of legend positions
legends_pos <- c("topleft", "topright", "topright", "topright", "topleft", "topright", "topleft", "topright", "topright")
for(i in 1:length(super_within_cor)) {
    plot.correlations(super_within_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i], tree = tree)
}
```

<!-- THIS BIT IS IGNORED FOR NOW -->

### Correlations per order (species on phylogeny)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = FALSE}
## Loading the sub-order levels
load(file = "../Data/Processed/shapespace_superorder_lvl_order_suborder.rda")
## Select the data
shapespaces <- lapply(shapespace_superorder_lvl_order_suborder, function(X) X$space)
## Create a level 3
order_space <- do.call(rbind, shapespaces)

## Calculate the correlations
order_phylo_cor <- get.correlations(results$order_results_phylo, exclude.phylo = TRUE)

## Getting the subsets for the colouring
split_tmp <- split(order_space$level2, f = order_space$level1)[-1] ## "Charadriiformes" "Galliformes" not included?
col_subs <- replicate(length(order_phylo_cor), NULL)
col_subs[-c(6,12)] <- split_tmp[match(sort(names(split_tmp)), names(split_tmp))]

## Set the vector of legend positions
for(i in 1:length(order_phylo_cor)) {
    plot.correlations(order_phylo_cor[[i]], col.sub = col_subs[[i]])
}
```

### Correlations within orders (species on group)

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 6, eval = FALSE}
## Calculate the correlations
order_within_cor <- get.correlations(results$order_results_within, exclude.phylo = TRUE)

## Getting the subsets for the colouring
split_tmp <- split(order_space$level2, f = order_space$level1)[-1] ## "Charadriiformes" "Galliformes" not included?
col_subs <- replicate(length(order_within_cor), NULL)
col_subs[-c(6,12)] <- split_tmp[match(sort(names(split_tmp)), names(split_tmp))]

## Set the vector of legend positions
for(i in 1:length(order_within_cor)) {
    plot.correlations(order_within_cor[[i]], col.sub = col_subs[[i]])
}
```