### 2022-05-25 Analysis/04-Figure1_with_ellipse_scalling_and_8D_histograms.Rmd
Updated passeriformes (with TODOS for functionalisation)

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This section contains the code for reproducing the analyses. To get the details, refer to 03-elaboration_innovation_analyses.Rmd.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
```

# All birds shapespace analyses

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Creating the subsets
superorder_shapespace <- MCMCglmm.subsets(n = 4000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

order_shapespace <- MCMCglmm.subsets(n = 4000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)
```

First we can visualise the global bird shape space (or at least the 2 first of the 8 dimensions):

```{r, message = FALSE, warning = FALSE, fig.height = 8, fig.width = 8, echo = FALSE}
## The ggplot color panel
gg.color.hue <- function(n) {
    grDevices::hcl(h = seq(15, 375, length = n + 1), l = 65, c = 100)[1:n]
}

# colour.palette <- rainbow
colour.palette <- gg.color.hue
```

```{r, echo = FALSE}
## Quick load
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list_4k.rda")
results <- shapespace_allbirds_lvl_superorder_order_results_list
rm(shapespace_allbirds_lvl_superorder_order_results_list)
```

## Phylogeny window (detailed)

```{r}
source("../Functions/covar.utilities.R")
## Plotting the ellipses
plot.one.ellipse <- function(data, name.col, x = TRUE, y = TRUE, main = "", legend = FALSE, with.phylo = FALSE) {

    ## Get the PC %ages
    if(x || y) {
        vars <- dispRity::variances(data$matrix[[1]])
        pc_var <- round((vars/sum(vars)*100), 2)
    }

    ## Do the empty plot
    lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
    plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = ifelse(x, paste0("PC1 (", pc_var[1], "%)"), ""), ylab = ifelse(y, paste0("PC2 (", pc_var[2], "%)"), ""), main = main)
    if(x) axis(1)
    if(y) axis(2)
    abline(v = 0, col = "grey", lwd = 0.5)
    abline(h = 0, col = "grey", lwd = 0.5)

    if(!with.phylo) {
        subsets <- names(name.col)
        cols <- name.col
    } else {
        subsets <- c("phylogeny", names(name.col))
        cols <- c("grey", name.col)
    }
    if(!with.phylo) {
        covar.plot(get.subsets(data, subsets = subsets),
               col = cols,
               ellipses = mean,
               legend = FALSE,
               points = FALSE,
               cex = 0.5, apply.to.VCV = TRUE,
               lwd = 4,
               add = TRUE,
               legend = legend)
    } else {
        covar.plot(get.subsets(data, subsets = subsets),
               col = cols,
               ellipses = mean,
               legend = FALSE,
               points = FALSE,
               cex = 0.5, apply.to.VCV = TRUE,
               lwd = 4,
               add = TRUE,
               legend = legend,
               scale = "phylogeny")        
    }
}
## Plot the projection/rejection
plot.one.proj.rej <- function(data, name.col, x = TRUE, y = TRUE, main = "") {

    ## Get the proj/rej
    elaboration <- get.disparity(data$position, concatenate = FALSE)
    exploration <- get.disparity(data$distance, concatenate = FALSE)

    ## Get the correct group
    if(length(ID <- which(names(elaboration) %in% names(name.col))) < 1) {
        ID <- which(names(elaboration) %in% paste0(names(name.col), ":phylogeny"))
    }
    elaboration <- c(elaboration[[ID]])
    exploration <- c(exploration[[ID]])

    ## Change the colours
    col <- c(name.col, adjustcolor(name.col, alpha.f = 0.5))

    ## Set the limits
    quant_95 <- quantile(c(elaboration, exploration), prob = 0.975)
    lim_up <- c(ceiling(quant_95/5)*5)

    ## Get the quantiles
    elaboration_quantiles <- quantile(elaboration, prob = c(0.025, 0.25, 0.50, 0.975))
    exploration_quantiles <- quantile(exploration, prob = c(0.025, 0.25, 0.50, 0.975))

    ## Empty plot
    par(bty = "o")
    plot(NULL, xlim = c(-0.5,1.5), ylim = c(0,lim_up), xaxt = "n", yaxt = "n", xlab = "", ylab = "")

    ## Add the lines
    lines(lty = 2, x = c(0,0), y = elaboration_quantiles[c(1,4)], col = col[1], lwd = 1)
    lines(lty = 1, x = c(0,0), y = elaboration_quantiles[c(2,3)], col = col[1], lwd = 3)
    lines(lty = 2, x = c(1,1), y = exploration_quantiles[c(1,4)], col = col[2], lwd = 1)
    lines(lty = 1, x = c(1,1), y = exploration_quantiles[c(2,3)], col = col[2], lwd = 3)
    points(0, median(elaboration), pch = 19, col = col[1])
    points(1, median(exploration), pch = 19, col = col[2])

    ## Add the y axis
    axis(2, at = seq(from = 0, to = lim_up, by = 5), labels = seq(from = 0, to = lim_up, by = 5), las = 2)

    ## Plot the results
    # boxplot(cbind(elaboration, exploration), horizontal = TRUE, ylim = c(0,lim_up), col = col, las = 1, xaxt = "n", yaxt = "n")
    # if(x) {
    #     axis(1, at = seq(from = 0, to = lim_up, by = 5), labels = seq(from = 0, to = lim_up, by = 5))
    # }
    # if(y) {
    #     axis(2, at = c(1, 2), labels = c("exploration", "innovation"), las = 2)
    # }
}
## Plot the variation per dimension
add.dims <- function(data, name.col, n = 1000) {
    covars_cent_tend <- lapply(sample.n(get.subsets(data, subset = names(name.col))$covar, n), VCV.cent.tend, mean)
    dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
    dim_var <- dim_var/max(dim_var)/10

    ## Select the quadrant
    if(colMeans(data$matrix[[1]][c(data$subsets[[names(name.col)]]$elements), c(1,2)])[1] > 0.1) {
        x_start <- -0.2
    } else {
        x_start <- 0.2
    }

    y_start <- -0.1

    buffer <- 0
    for(one_dim in 1:length(dim_var)) {
        lines(x = c(x_start, x_start+dim_var[one_dim]), y = rep(y_start+buffer, 2), lwd = 3, col = name.col)
        buffer <- buffer - 0.02
    }
}

# pdf(file = "Figure1a.pdf", height = 11, width = 5.5)
# ## Plot the phylo ellipse
# grey_palette <- c(colour.palette(8), "grey")
# par(mfrow = c(2,1), bty = "n")
# plot.one.ellipse(superorder_shapespace, ID = c(1,9), colours = grey_palette, x = TRUE, y = TRUE, main = "A")
# legend("topright", lwd = 5, col = grey_palette[c(1,9)], lty = 1, legend = c("Aequornithes' median variance-covariance", "Overall phylogenetic's median variance-covariance"), bg = "white")
# plot.one.proj.rej(results$super_results_between, ID = 1, colours = grey_palette)
```

```{r}
box_size <- 1.5
# pdf(file = "../Manuscript/Figures/Figure1_phylo_Dplot.pdf", height = box_size*2, width = box_size*2)

# lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
# plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "", ylab = "", main = "")
# abline(v = 0, col = "grey", lwd = 0.5)
# abline(h = 0, col = "grey", lwd = 0.5)

# covars_cent_tend <- lapply(sample.n(get.subsets(data, subset = "phylogeny")$covar, n), VCV.cent.tend, mean)
# dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
# dim_var <- dim_var/max(dim_var)/10
# x_start <- 0.2
# y_start <- -0.1
# buffer <- 0
# for(one_dim in 1:length(dim_var)) {
#     lines(x = c(x_start, x_start+dim_var[one_dim]), y = rep(y_start+buffer, 2), lwd = 3, col = "grey")
#     buffer <- buffer - 0.02
# }

# dev.off()

pdf(file = "../Manuscript/Figures/Figure1.pdf", height = box_size*8, width = box_size*10)

## Global window
lay_mat_order <- rbind(
        matrix(c(1,1, 2:9), ncol = 10, nrow = 1), # line 1
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1), # line 2
        matrix(c(1,1, 18:25), ncol = 10, nrow = 1), # line 3
        matrix(c(26:35), ncol = 10, nrow = 1), # line 4
        matrix(c(36:45), ncol = 10, nrow = 1), # line 5
        matrix(c(46:55), ncol = 10, nrow = 1) # line 6
    )
lay_mat_super <- rbind(
        matrix(c(1,1, 2:9), ncol = 10, nrow = 1), # line 7
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1) # line 8
    )

heights_vals <- rep(1, 9)
widths_vals <- rep(c(1, 1/3), 5)
layout_matrix <- layout(rbind(lay_mat_super, (lay_mat_order+17)), widths = widths_vals, heights = heights_vals, respect = TRUE)

# layout.show(layout_matrix)

# layout_matrix <- layout(cbind(matrix(1, ncol = 1, nrow = 27),
#                               matrix(2:((27*2)+1), ncol = 2, nrow = 27, byrow = TRUE)))
# layout.show(layout_matrix)

## PLOTTING THE SUPER ORDERS
tree <- ladderize(tree)
## Get the tip_values
orders_lvl <- shapespace$level2
superorders_lvl <- shapespace$level1

## Select the level
selected_level <- coloured_levels <- superorders_lvl

## Replace the order levels as colours
colour_palette <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
levels(coloured_levels) <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
# par(mar = c(1, 1, 1, 1)+0.1)
# plot(tree, show.tip.label = FALSE, edge.color =  match.tip.edge(as.character(coloured_levels), tree, replace.na = "grey"))

## Shorten the tree
selected_tips <- character()
for(one_level in 2:length(levels(selected_level))) {
    selected_tips[one_level-1] <- rownames(shapespace)[which(selected_level == levels(selected_level)[one_level])[1]]
    names(selected_tips)[one_level-1] <- levels(selected_level)[one_level]
}

## Shorten the tree
tree_short <- drop.tip(tree, tip = tree$tip.label[!(tree$tip.label %in% selected_tips)])
tree_short$tip.label <- names(selected_tips)[match(tree_short$tip.label, selected_tips)]

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))
## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)
# tiplabels()

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Paleognathae" = 1,
    "Galloanserae" = 2,
    "Otidimorphae" = 5,
    "Aequornithes" = 6,
    "Mirandornithes" = 3,
    "Columbimorphae" = 4,
    "Telluraves" = 8,
    "Strisores" = 7)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
tip_colours_super_orders <- tip_colours
save(tip_colours_super_orders, file = "../Data/Processed/tip_colours_super_orders.rda")

# par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
# plot(rep(1,27), 1:27, col = "white")
# labs <- names(tip_colours)
# cols <- tip_colours
# text(x = rep(1,27), y = 1:27, label = rev(labs), col = rev(cols))

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    # name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(superorder_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)

    ## Calculate which dimension contains the most variance
    add.dims(superorder_shapespace, name.col)
    n <- 1000
    covars_cent_tend <- lapply(sample.n(get.subsets(superorder_shapespace, subset = names(name.col))$covar, n), VCV.cent.tend, mean)
    dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
    # text(0,0.3, paste0(names(name.col), " (", which(dim_var == max(dim_var)), ")"), cex = 1, col = name.col)
    text_main <- paste0(names(name.col))#, "\n", "λe = ", round(lambdas_superorder[paste0("superorders.", names(name.col)),][1], 2), "; λi = ", round(lambdas_superorder[paste0("superorders.", names(name.col)),][2], 2))
    text(0,0.3, text_main, cex = 1.2, col = name.col)

    ## Plot the elaboration/exploration
    par(mar = c(2, 1, 1, 1)+0.1)
    plot.one.proj.rej(results$super_results_between, name.col, x = TRUE, y = FALSE)
}

## PLOTTING THE ORDERS
tree <- ladderize(tree)
## Get the tip_values
orders_lvl <- shapespace$level2
superorders_lvl <- shapespace$level1

## Select the level
selected_level <- coloured_levels <- orders_lvl

## Replace the order levels as colours
colour_palette <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
levels(coloured_levels) <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))

## Shorten the tree
selected_tips <- character()
for(one_level in 2:length(levels(selected_level))) {
    selected_tips[one_level-1] <- rownames(shapespace)[which(selected_level == levels(selected_level)[one_level])[1]]
    names(selected_tips)[one_level-1] <- levels(selected_level)[one_level]
}

## Shorten the tree
tree_short <- drop.tip(tree, tip = tree$tip.label[!(tree$tip.label %in% selected_tips)])
tree_short$tip.label <- names(selected_tips)[match(tree_short$tip.label, selected_tips)]

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))

## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

# tip_order <- sort(tip_order)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
tip_colours_orders <- tip_colours
save(tip_colours_orders, file = "../Data/Processed/tip_colours_orders.rda")

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    # name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(order_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)

    ## Calculate which dimension contains the most variance
    add.dims(order_shapespace, name.col)
    n <- 1000
    covars_cent_tend <- lapply(sample.n(get.subsets(order_shapespace, subset = names(name.col))$covar, n), VCV.cent.tend, mean)
    dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
    # text(0,0.3, paste0(names(name.col), " (", which(dim_var == max(dim_var)), ")"), cex = 1, col = name.col)
    text_main <- paste0(names(name.col))#, "\n", "λe = ", round(lambdas_order[paste0("orders.", names(name.col)),][1], 2), "; λi = ", round(lambdas_order[paste0("orders.", names(name.col)),][2], 2))
    text(0,0.3, text_main, cex = 1.2, col = name.col)

    ## Plot the elaboration/exploration
    par(bty = "n", mar = c(2, 1, 1, 1)+0.1)
    plot.one.proj.rej(results$order_results_between, name.col, x = TRUE, y = FALSE)
}
dev.off()

```

# Passeriformes bit

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapsepace_passeriformes_suborder_family.rda")
## Select the data
shapespace <- shapsepace_passeriformes_suborder_family[[1]]$space
tree <- shapsepace_passeriformes_suborder_family[[1]]$consensus_tree
## Clean garbage
rm(shapsepace_passeriformes_suborder_family)

## Load the covar
load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
shapespace_passeriformes <- shapespace_passeriformes[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_passeriformes) == "animal")
subor_lvl <- grep("level1_", MCMCglmm.levels(shapespace_passeriformes))
famil_lvl <- grep("level2_", MCMCglmm.levels(shapespace_passeriformes))

## Creating the subsets
suborder_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_passeriformes,
      group         = MCMCglmm.levels(shapespace_passeriformes)[c(subor_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

family_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_passeriformes,
      group         = MCMCglmm.levels(shapespace_passeriformes)[c(famil_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_passeriformes)
```

First we can visualise the global bird shape space (or at least the 2 first of the 8 dimensions):

```{r, echo = FALSE}
## Quick load
load(file = "../Data/Processed/shapespace_passeriformes_suborder_family_results_list.rda")
results <- shapespace_passeriformes_suborder_family_results_list
rm(shapespace_passeriformes_suborder_family_results_list)
```

## Phylogeny window (detailed)

```{r}

source("../Functions/covar.utilities.R")
source("../Functions/ploting.utilities.R")

box_size <- 1.5
pdf(file = "../Manuscript/Figures/Figure1_passeriformes.pdf", height = box_size*7, width = box_size*10)

## Global window
lay_mat_family <- rbind(
        matrix(c(1,1, 2:9), ncol = 10, nrow = 1), # line 1
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1), # line 2
        matrix(c(1,1, 18:25), ncol = 10, nrow = 1), # line 3
        matrix(c(26:35), ncol = 10, nrow = 1), # line 4
        matrix(c(36:45), ncol = 10, nrow = 1), # line 5
        matrix(c(46:55), ncol = 10, nrow = 1) # line 6
    )
lay_mat_sub <- rbind(
        matrix(c(1,1, 2:5, rep(0, 4)), ncol = 10, nrow = 1) # line 7
    )

heights_vals <- rep(1, 8)
widths_vals <- rep(c(1, 1/3), 5)
layout_matrix <- layout(rbind(lay_mat_sub, (lay_mat_family+5)), widths = widths_vals, heights = heights_vals, respect = TRUE)

# layout.show(layout_matrix)

## TODO: FUNCTIONALISE THE TREE PLOTTING: START

## PLOTTING THE SUBORDERS
tree <- ladderize(tree)
## Get the tip_values
families_lvl <- shapespace$level2
suborders_lvl <- shapespace$level1

## Select the level
selected_level <- coloured_levels <- suborders_lvl

## Replace the order levels as colours
colour_palette <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
levels(coloured_levels) <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))

## Shorten the tree
selected_tips <- character()
for(one_level in 2:length(levels(selected_level))) {
    selected_tips[one_level-1] <- rownames(shapespace)[which(selected_level == levels(selected_level)[one_level])[1]]
    names(selected_tips)[one_level-1] <- levels(selected_level)[one_level]
}

## Shorten the tree
tree_short <- drop.tip(tree, tip = tree$tip.label[!(tree$tip.label %in% selected_tips)])
tree_short$tip.label <- names(selected_tips)[match(tree_short$tip.label, selected_tips)]

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))
## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)

## TODO: FUNCTIONALISE THE TREE PLOTTING: END

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tyranni" = 1,
    "Passeri" = 2)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
tip_colours_suborders <- tip_colours
save(tip_colours_suborders, file = "../Data/Processed/tip_colours_suborders.rda")

## Plotting the ellipses for each suborder
for(i in 1:length(tip_order)) {

    ## TODO: FUNCTIONALISE THE ELLIPSE PLOTTING: START

    ## Select the colour and name
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(suborder_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)

    ## Calculate which dimension contains the most variance
    add.dims(suborder_shapespace, name.col)
    n <- 1000
    covars_cent_tend <- lapply(sample.n(get.subsets(suborder_shapespace, subset = names(name.col))$covar, n), VCV.cent.tend, mean)
    dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
    text_main <- paste0(names(name.col))
    text(0,0.3, text_main, cex = 1.2, col = name.col)

    ## Plot the elaboration/exploration
    par(mar = c(2, 1, 1, 1)+0.1)
    plot.one.proj.rej(results$suborder_results_between, name.col, x = TRUE, y = FALSE)

        ## TODO: FUNCTIONALISE THE ELLIPSE PLOTTING: END
}

## PLOTTING THE FAMILIES
tree <- ladderize(tree)
## Select the level
selected_level <- coloured_levels <- families_lvl

## Replace the order levels as colours
colour_palette <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
levels(coloured_levels) <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))

## Shorten the tree
selected_tips <- character()
for(one_level in 2:length(levels(selected_level))) {
    selected_tips[one_level-1] <- rownames(shapespace)[which(selected_level == levels(selected_level)[one_level])[1]]
    names(selected_tips)[one_level-1] <- levels(selected_level)[one_level]
}

## Shorten the tree
tree_short <- drop.tip(tree, tip = tree$tip.label[!(tree$tip.label %in% selected_tips)])
tree_short$tip.label <- names(selected_tips)[match(tree_short$tip.label, selected_tips)]

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))

## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Pittidae" = 2,
    "Eurylaimidae" = 1,
    "Tyrannidae" = 5,
    "Pipridae" = 3,
    "Cotingidae" = 4,
    "Thamnophilidae" = 6,
    "Furnaridae" = 9,
    "Rhinocryptidae" = 8,
    "Grallariidae" = 7,
    "Ptilonorhynchidae" = 10,
    "Maluridae" = 11,
    "Meliphagidae" = 13,
    "Acanthizidae" = 12,
    "Vireonidae" = 20,
    "Pachycephalidae" = 18,
    "Oriolidae" = 19,
    "Laniidae" = 25,
    "Corvidae" = 24,
    "Paradisaeidae" = 23,
    "Monarchidae" = 22,
    "Dicruridae" = 21,
    "Campephagidae" = 14,
    "Malaconotidae" = 16,
    "Vangidae" = 15,
    "Platysteiridae" = 17,
    "Troglodytidae" = 27,
    "Sittidae" = 26)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
tip_colours_families <- tip_colours
save(tip_colours_families, file = "../Data/Processed/tip_colours_families.rda")

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(family_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)

    ## Calculate which dimension contains the most variance
    add.dims(family_shapespace, name.col)
    n <- 1000
    covars_cent_tend <- lapply(sample.n(get.subsets(family_shapespace, subset = names(name.col))$covar, n), VCV.cent.tend, mean)
    dim_var <- apply((get.one.axis(covars_cent_tend[[1]], axis = 1, level = 0.95)), 2, dist)
    text_main <- paste0(names(name.col))
    text(0,0.3, text_main, cex = 1.2, col = name.col)

    ## Plot the elaboration/exploration
    par(bty = "n", mar = c(2, 1, 1, 1)+0.1)
    plot.one.proj.rej(results$family_results_between, name.col, x = TRUE, y = FALSE)
}
dev.off()

```