### 2022-06-10 Analysis/09-novelty_predictions.Rmd
fast piggles done

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This markdown contains the code to reproduce the figures containing the elaboration and innovation per species for the all birds and the passeriformes analyses.
For the details on how to get the data objects called here, refer to `03-elaboration_innovation_analyses.Rmd`.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
library(phylolm)
source("../Functions/ploting.utilities.R")
source("../Functions/correlations.utilities.R")
```

# All birds analysis

## PGLS tables

This table contains the results of the model `centroids ~ elaboration*innovation` where `centroids` is a proxy for novelty.
Results at different levels are for that level (i.e. or the Anseriformes results, it's their distance from their centroid and their elaboration/projection on their VCV major axis).

```{r, eval = TRUE, echo = FALSE}
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_superorder.rda")
load("../Data/Processed/ploting_data_order.rda")
load("../Data/Processed/clade_info.rda")

## Running one pgls
run.one.pgls <- function(group, data, clade_info) {
      cat(".")
      if(group != "all_birds") {
            ## Extract the specific group
            if(group %in% names(clade_info[[1]])) {
                  ## Is super order
                  clade_info <- clade_info[[1]]
            } else {
                  ## Is order
                  clade_info <- clade_info[[2]]
            }

            ## Get the species in the clade
            sp_list <- clade_info[[which(names(clade_info) == group)]]$species

            ## Clean the tree
            short_tree <- drop.tip(data$consensus_tree,
                                   tip = data$consensus_tree$tip.label[!(data$consensus_tree$tip.label %in% sp_list)])
      } else {
            ## Doing it on all birds
            sp_list <- names(data$median_elaborations)
            short_tree <- data$consensus_tree
      }

      ## Set up the data
      short_data <- data.frame("centroids" = data$distances_to_centre[sp_list],
                               "elaboration" = data$median_elaborations[sp_list],
                               "innovation" = data$median_innovation[sp_list])

      ## Run the piggles
      return(phylolm::phylolm(centroids ~ elaboration*innovation, data = short_data, phy = short_tree, model = "lambda"))
}

## Get the mega level pgls
all_birds_pgls <- run.one.pgls("all_birds", ploting_data, clade_info)

## Get the macro level pgls
superorder_pgls <- lapply(as.list(names(clade_info[[1]])), run.one.pgls, ploting_data_superorder, clade_info)
order_pgls <- lapply(as.list(names(clade_info[[2]])), run.one.pgls, ploting_data_order, clade_info)

## Summarise that?
extract.summary <- function(model) {
      sum_mod <- summary(model)
      coefficients <- c(t(sum_mod$coefficients[-1, c(1, 4)]))
      names(coefficients) <- c("elaboration", "p_elab", "innovation", "p_innov", "interaction", "p_interact")
      c("lambda" = sum_mod$optpar, coefficients)
}

## Combine these results
all_results <- as.data.frame(rbind(extract.summary(all_birds_pgls),
                     do.call(rbind, lapply(superorder_pgls, extract.summary)),
                     do.call(rbind, lapply(order_pgls, extract.summary))))
## Add the labels
labels <- data.frame("name"  = c("All_birds", names(clade_info[[1]]), names(clade_info[[2]])),
                     "level" = c("all_birds", rep("super_order", length(clade_info[[1]])),  rep("order", length(clade_info[[2]]))))

knitr::kable(cbind(labels, round(all_results, 3)))
```