### 2022-06-13 Analysis/10-sensitivity_analysis.Rmd
Create 10-sensitivity_analysis.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This markdown contains the script for runing the sensitivty analyses.
Due to computational time, the sensitivity analysis are only run on a subset of the entire trait space, the charadriiformes.

```{r}
library(dispRity)
library(mcmcmcglmmm)
```

## Observed results for charadriiformes

For this section we are just running the normal analysis (as in the main text) but on the charadriiformes species only.
The charadriiformes dataset is a present as an example dataset in the dispRity package

```{r, echo = FALSE, eval = TRUE}
## Loading the Charadriiformes data
data(charadriiformes)
## Extracting the tree
tree <- charadriiformes$tree
## Extracting the data column that contains the clade assignments
data <- charadriiformes$data[, "clade"]
## Changing the levels names (the clade names) to colours
levels(data) <- c("orange", "blue", "darkgreen")
data <- as.character(data)
## Matching the data rownames to the tip order in the tree
data <- data[match(ladderize(tree)$tip.label, rownames(charadriiformes$data))]

## Matching the tip colours (labels) to their descending edges in the tree
## (and making the non-match edges grey)
clade_edges <- match.tip.edge(data, tree, replace.na = "grey")

## Plotting the results
plot(ladderize(tree), show.tip.label = FALSE, edge.color = clade_edges, main = "Charadriiformes")
legend("bottomleft", lty = c(1,1,1), col = c("blue", "darkgreen", "orange"), legend = c("plovers", "sandpipers", "gulls"))
axisPhylo()
```
Fgure 1: the charadriiformes consensus tree

```{r, echo = FALSE, eval = TRUE}
## Creating a dispRity object from the charadriiformes model
plot(shapespace[, c(1,2)], xlab = "PC1", ylab = "PC2", xlim = c(-0.4, 0.4), ylim = c(-0.4, 0.4), main = "shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
legend("bottomrigh", pch = 19, legend = c("gulls", "plovers", "sandpipers"), col= c("orange", "blue", "darkgreen"))
```
Figure 2: the charadriiformes shapespace

## Ball shapesapce

In this sensitivity analysis we transform the shapespace to make each dimension the same size (i.e. each dimension explains the same amount of variance).
We do that by simply scaling the values on each dimensions.

## No phylogenetic signal

In this sensitivity we are aiming to "remove" the phylogenetic signal by randomising the attribution of each species traits.
We keep the phylogenetic structure as presented above (i.e. the charadriiformes tree) but randomly attribute a beak shape value to each species to dilute or "remove" the phylogenetic signal in each clade.

### Preparing the data

```{r, eval = FALSE, echo = FALSE}
source("../Functions/prep.data.R")
## Reading the different trait spaces
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")
shapespace <- shapespace$PCscores
## Reading the consensus tree
consensus_tree <- read.nexus("../Data/Raw/Tree_data/9993taxa_1_tree_Global_MCC_CA.tre")
## Reading the distribution of trees
trees_list <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")

## Making a charadriiformes only list
data(charadriiformes)
clade_list <- charadriiformes$data[,18, drop = FALSE]
colnames(clade_list) <- "level1"

## Shapespace passeriformes
shapespace_charadriiformes <- shapespace[match(rownames(clade_list),rownames(shapespace)), ]

## Do the prep data from here (with all the trees)
shapespace_charadriiformes_family <- prep.data(
                                          level = 0, lvl.inc = c(1),
                                          clades    = clade_list,
                                          space     = shapespace_charadriiformes,
                                          dim       = 1:8,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_charadriiformes_family, file = "../Data/Processed/shapespace_charadriiformes_family.rda")
```

### Setting the mini-chains

```{r, eval = TRUE, message = FALSE}
## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_charadriiformes_family.rda")

## Select the data
shapespace <- shapespace_charadriiformes_family[[1]]$space
selected_dim <- shapespace_charadriiformes_family[[1]]$dimensions
consensus_tree <- shapespace_charadriiformes_family[[1]]$consensus_tree
trees_list <- shapespace_charadriiformes_family[[1]]$trees_list
```

## Mini-chains parametrization

```{r, eval = TRUE}
## Set up the parametrising chains models
param_chains <- make.mini.chains(data         = shapespace,
                                 dimensions   = selected_dim,
                                 tree         = consensus_tree,
                                 trait.family = "gaussian",
                                 randoms      = c("global", "level1"),
                                 residuals    = "global",
                                 priors       = 0.02,
                                 verbose      = TRUE,
                                 parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

# Running the three MCMCglmm models
save(param_chains, file = "../Data/Processed/params_shapespace_charadriiformes.rda")
```

```{r, eval = TRUE}
## Run the parametrisation chains (local)
params <- run.mini.chains(param_chains, replicates = 3)

## Get the estimated parameters
est_params <- extract.parameters(params)
```

## Set up the mini-chain

```{r, eval = TRUE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10
## Set up the mini-chains models
shapespace_charadriiformes <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 400)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                      nitt   = est_params$burnin + thin * samples,
                                      thin   = thin,
                                      burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_charadriiformes, file = "../Data/Cluster/shapespace_charadriiformes/shapespace_charadriiformes.mini.chains")
```

# Randomspace

```{r}
## Get the scaling the variance of each dimension
randospace <- shapespace
randomise.space <- function(space, dims) {
      rando <- space
      rando[, dims] <- space[sample(1:nrow(space), nrow(space)), dims]
      return(rando)
}

## Comparing D1 and D3 for the contrast
par(mfrow = c(2,1))
plot(shapespace[, c(1, 2)], xlab = "PC1", ylab = "PC2", xlim = c(-0.4, 0.4), ylim = c(-0.4, 0.4),
      main = "shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
plot(randomise.space(shapespace, selected_dim)[, c(1, 2)], xlab = "PC1", ylab = "PC2", xlim = c(-0.4, 0.4), ylim = c(-0.4, 0.4),
      main = "randomised shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
```

## Mini-chains (using the parameters from the normal chain above)

```{r, eval = TRUE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Setting up 20 random replicates
for(i in 1:25) {
      ## Set up the mini-chains models
      randospace_charadriiformes_ <- make.mini.chains(
                                      data         = randomise.space(shapespace, selected_dim),
                                      dimensions   = selected_dim,
                                      tree         = trees_list[sample(1:1000, 400)],
                                      trait.family = "gaussian",
                                      randoms      = c("global", "level1"),
                                      residuals    = "global",
                                      priors       = est_params$priors,
                                      verbose      = TRUE,
                                      parameters   = list(
                                            nitt   = est_params$burnin + thin * samples,
                                            thin   = thin,
                                            burnin = est_params$burnin))
      ## Save the mini.chains
      save(randospace_charadriiformes_, file = paste0("../Data/Cluster/shapespace_charadriiformes/randospace_charadriiformes_", i, ".mini.chains"))
}
```

# Ballspace

```{r}
## Get the scaling the variance of each dimension
ballspace <- shapespace
ballspace[, selected_dim] <- apply(shapespace[, selected_dim], 2, scale)

## Comparing D1 and D3 for the contrast
par(mfrow = c(2,1))
plot(shapespace[, c(1,3)], xlab = "PC1", ylab = "PC3", xlim = c(-0.4, 0.4), ylim = c(-0.4, 0.4),
      main = "shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
plot(ballspace[, c(1,3)], xlab = "PC1", ylab = "PC3", xlim = c(-3, 3), ylim = c(-3, 3), ,
      main = "balled shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
```

## Mini-chains parametrization

```{r, eval = TRUE}
## Set up the parametrising chains models
param_chains <- make.mini.chains(data         = ballspace,
                                 dimensions   = selected_dim,
                                 tree         = consensus_tree,
                                 trait.family = "gaussian",
                                 randoms      = c("global", "level1"),
                                 residuals    = "global",
                                 priors       = 0.02,
                                 verbose      = TRUE,
                                 parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

# Running the three MCMCglmm models
save(param_chains, file = "../Data/Processed/params_ballspace_charadriiformes.rda")
```

```{r, eval = TRUE}
## Run the parametrisation chains (local)
params <- run.mini.chains(param_chains, replicates = 3)

## Get the estimated parameters
est_params <- extract.parameters(params)
```

## Set up the mini-chain

```{r, eval = TRUE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10
## Set up the mini-chains models
ballspace_charadriiformes <- make.mini.chains(
                                data         = ballspace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 400)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                      nitt   = est_params$burnin + thin * samples,
                                      thin   = thin,
                                      burnin = est_params$burnin))
## Save the mini.chains
save(ballspace_charadriiformes, file = "../Data/Cluster/shapespace_charadriiformes/ballspace_charadriiformes.mini.chains")
```