### 2022-06-16 Analysis/10-sensitivity_analysis.Rmd
Update 10-sensitivity_analysis.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This markdown contains the script for runing the sensitivty analyses.
Due to computational time, the sensitivity analysis are only run on a subset of the entire trait space, the charadriiformes.

```{r}
set.seed(42)
library(dispRity)
library(mcmcmcglmmm)
```

# Null models

From this space we simulated three null model scenarios to test the impact of specific parameters in our approach:

 * the `randomspace` model to test the effect of phylogenetic signal on our results: we can test the effect of phylogenetic correlation and phylogenetic signal on our results by randomly assigning beaks to different species to "dilute" the phylogenetic signal: i.e. so that there is no assumption that two closely related species have similar beaks. We do that be reasigning beaks shape in the shapespace randomly for each species in each step of the `mcmcmcglmmm` analysis.

 * the `ballspace` model to test the effect of decreasing variance on each dimension of the shapespace: in fact, our main findings (elaboration is more common than innovation at the species level) might be due to the nature of the shapespace where the first dimension (PC1) has more variance than the other ones. This can pull the major axis of the global phylogenetic VCV to be aligned with PC1 where there is more opportunity for each individiual species to have higher values of elaboration (i.e. more changes aligned with PC1) that innovation (i.e. less changes on the other PCs). We can test this by rescaling the variance of each PC axes to be equal.

 * the `BMspace` model to test the deviation from a base Brownian Motion trait evolution model: for this model we simulate a series of shapespaces with the same number of species and the same phylogeny using a Brownian motion process for each 8 dimensions with a decreasing variance on each dimensions to match the observed variance in per dimension in the shapespace.

For each of these null models, we resimulated a shapespace with the modifications highlighted above and rerun the whole `mcmcmcglmmm` and projection pipeline as for in the main study.
For calculation speed reasons we only ran these models on a subset of the global dataset, the charadriiformes containing 359 sp divided into three groups (159 gulls, 98 plovers and 102 sandpipers).

# Simulating the different spaces

## Original data

```{r, eval = FALSE, echo = FALSE}
source("../Functions/prep.data.R")
## Reading the different trait spaces
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")
shapespace <- shapespace$PCscores
## Reading the consensus tree
consensus_tree <- read.nexus("../Data/Raw/Tree_data/9993taxa_1_tree_Global_MCC_CA.tre")
## Reading the distribution of trees
trees_list <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")

## Making a charadriiformes only list
data(charadriiformes)
clade_list <- charadriiformes$data[,18, drop = FALSE]
colnames(clade_list) <- "level1"

## Shapespace passeriformes
shapespace_charadriiformes <- shapespace[match(rownames(clade_list),rownames(shapespace)), ]

## Do the prep data from here (with all the trees)
shapespace_charadriiformes_family <- prep.data(
                                          level = 0, lvl.inc = c(1),
                                          clades    = clade_list,
                                          space     = shapespace_charadriiformes,
                                          dim       = 1:8,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_charadriiformes_family, file = "../Data/Processed/shapespace_charadriiformes_family.rda")
```

```{r, eval = TRUE, echo = FALSE}
## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_charadriiformes_family.rda")

## Select the data
shapespace <- shapespace_charadriiformes_family[[1]]$space
selected_dim <- shapespace_charadriiformes_family[[1]]$dimensions
consensus_tree <- shapespace_charadriiformes_family[[1]]$consensus_tree
trees_list <- shapespace_charadriiformes_family[[1]]$trees_list
```

```{r, echo = FALSE, eval = TRUE}
## Loading the Charadriiformes data
data(charadriiformes)
## Extracting the tree
tree <- charadriiformes$tree
## Extracting the data column that contains the clade assignments
data <- charadriiformes$data[, "clade"]
## Changing the levels names (the clade names) to colours
levels(data) <- c("orange", "blue", "darkgreen")
data <- as.character(data)
## Matching the data rownames to the tip order in the tree
data <- data[match(ladderize(tree)$tip.label, rownames(charadriiformes$data))]

## Matching the tip colours (labels) to their descending edges in the tree
## (and making the non-match edges grey)
clade_edges <- match.tip.edge(data, tree, replace.na = "grey")

## Plotting the results
plot(ladderize(tree), show.tip.label = FALSE, edge.color = clade_edges, main = "Charadriiformes")
legend("bottomleft", lty = c(1,1,1), col = c("blue", "darkgreen", "orange"), legend = c("plovers", "sandpipers", "gulls"))
axisPhylo()
```

Fgure 1: the charadriiformes consensus tree

```{r, echo = FALSE, eval = TRUE}
## Shapespace var
shape_var_prop <-  apply(shapespace[, selected_dim], 2, var)
shape_var_prop <- shape_var_prop/sum(shape_var_prop)*100

## Plotting fun
write.pc <- function(axis, var) {
      paste0("PC", axis, " (", round(var[axis], 2), "%)")
}
range.axis <- function(space, dims) {
      return(range(pretty(unlist(c(space[, dims])))))
}

## Creating a dispRity object from the charadriiformes model
plot(shapespace[, c(1,2)], xlab = write.pc(1, shape_var_prop), ylab = write.pc(2, shape_var_prop),
    xlim = range.axis(shapespace, selected_dim), ylim = range.axis(shapespace, selected_dim),
    main = "shapespace", pch = 19, col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
legend("bottomrigh", pch = 19, legend = c("gulls", "plovers", "sandpipers"), col= c("orange", "blue", "darkgreen"))
```

Figure 2: the charadriiformes shapespace

## The randomspace

To generate the `randomspace` we shuffle each beak to random species (i.e. reshuffling rows in the shapespace):

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12}
## Randomise each species
randomise.space <- function(space, dims) {
      rando <- space
      rando[, dims] <- space[sample(1:nrow(space), nrow(space)), dims]
      return(rando)
}

## Comparing D1 and D3 for the contrast
par(mfrow = c(1,2))
plot(shapespace[, c(1, 2)], xlab = write.pc(1, shape_var_prop), ylab = write.pc(2, shape_var_prop),
      xlim = range.axis(shapespace, selected_dim), ylim = range.axis(shapespace, selected_dim),
      main = "Original shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
plot(randomise.space(shapespace, selected_dim)[, c(1, 2)], xlab = write.pc(1, shape_var_prop), ylab = write.pc(2, shape_var_prop),
      xlim = range.axis(shapespace, selected_dim), ylim = range.axis(shapespace, selected_dim),
      main = "Randomised shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
```

Figure 3: difference between the shapespace and the randomspace

## The ballspace

To generate the `ballspace` we scaled the data on each dimension to have same variance:

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12}
## Get the scaling the variance of each dimension
ballspace <- shapespace
ballspace[, selected_dim] <- apply(shapespace[, selected_dim], 2, scale)

## Ballspace var
ball_var_prop <-  apply(ballspace[, selected_dim], 2, var)
ball_var_prop <- ball_var_prop/sum(ball_var_prop)*100

## Comparing D1 and D3 for the contrast
par(mfrow = c(1,2))
plot(shapespace[, c(1,3)], xlab = write.pc(1, shape_var_prop), ylab = write.pc(3, shape_var_prop),
      xlim = range.axis(shapespace, selected_dim), ylim = range.axis(shapespace, selected_dim),
      main = "Original shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
plot(ballspace[, c(1,3)], xlab = write.pc(1, ball_var_prop), ylab = write.pc(3, ball_var_prop),
      xlim = range.axis(ballspace, selected_dim), ylim = range.axis(ballspace, selected_dim),
      main = "Balled shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
```

Figure 4: difference between the shapespace and the ballspace

## The BMspace

To generate the `BMspace` we simulate 8 uncorrelated Brownian Motions on the consensus tree with the variance for each BM being the observed one for each dimension of the shapespace:

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12}
## Get the scaling the variance of each dimension
BMise.space <- function(space, tree, dims) {
      ## Get the space dimenions variance
      space_var_prop <-  apply(space[, dims], 2, var)
      space_var_prop <- space_var_prop/sum(space_var_prop)*100

      ## Simulate the space
      BMspace <- as.data.frame(sapply(space_var_prop, function(var, tree) phytools::fastBM(tree = tree, sig2 = var), tree))

      ## Match the species with the correct levels
      BMspace <- BMspace[match(space$animal, rownames(BMspace)), ]
      BMspace$level1 <- space$level1
      BMspace$animal <- space$animal
      return(BMspace)
}

## Comparing D1 and D3 for the contrast
par(mfrow = c(1,2))
plot(shapespace[, c(1, 2)], xlab = write.pc(1, shape_var_prop), ylab = write.pc(2, shape_var_prop),
      xlim = range.axis(shapespace, selected_dim), ylim = range.axis(shapespace, selected_dim),
      main = "Original shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
BMspace <- BMise.space(shapespace, tree, selected_dim)
BM_var_prop <-  apply(BMspace[, selected_dim], 2, var)
BM_var_prop <- BM_var_prop/sum(BM_var_prop)*100
plot(BMspace[, c(1, 2)], xlab = write.pc(1, BM_var_prop), ylab = write.pc(2, BM_var_prop),
      xlim = range.axis(BMspace, selected_dim), ylim = range.axis(BMspace, selected_dim),
      main = "BM simulated shapespace", pch = 19,
      col = c("orange", "blue", "darkgreen")[as.numeric(shapespace$level)])
```

Figure 5: difference between the shapespace and the BMspace

# Setting the mini-chains

## Original and randomised spaces (same parameters)

We used the same mini chains parameters for the BMspace and the randomspace since both are stochastic spaces (i.e. each space is a different itteration) making it to intensive computationally to estimate the starting parameters every iteration.
We therefore used the conservative parameters from the original space.
We used three chains of 20000 iterations for the parameterisation chains.

```{r, eval = FALSE, echo = FALSE}
## Set up the parametrising chains models
param_chains <- make.mini.chains(data         = shapespace,
                                 dimensions   = selected_dim,
                                 tree         = consensus_tree,
                                 trait.family = "gaussian",
                                 randoms      = c("global", "level1"),
                                 residuals    = "global",
                                 priors       = 0.02,
                                 verbose      = TRUE,
                                 parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

# Running the three MCMCglmm models
save(param_chains, file = "../Data/Processed/params_shapespace_charadriiformes.rda")
```

```{r, eval = FALSE, echo = FALSE}
## Run the parametrisation chains (local)
params <- run.mini.chains(param_chains, replicates = 3)

## Get the estimated parameters
est_params <- extract.parameters(params)
```

```{r, eval = FALSE, echo = FALSE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10
## Set up the mini-chains models
shapespace_charadriiformes <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 400)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                      nitt   = est_params$burnin + thin * samples,
                                      thin   = thin,
                                      burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_charadriiformes, file = "../Data/Cluster/shapespace_charadriiformes/shapespace_charadriiformes.mini.chains")
```

```{r, eval = FALSE, echo = FALSE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Setting up 20 random replicates
for(i in 1:25) {
      ## Set up the mini-chains models
      randospace_charadriiformes_ <- make.mini.chains(
                                      data         = randomise.space(shapespace, selected_dim),
                                      dimensions   = selected_dim,
                                      tree         = trees_list[sample(1:1000, 400)],
                                      trait.family = "gaussian",
                                      randoms      = c("global", "level1"),
                                      residuals    = "global",
                                      priors       = est_params$priors,
                                      verbose      = TRUE,
                                      parameters   = list(
                                            nitt   = est_params$burnin + thin * samples,
                                            thin   = thin,
                                            burnin = est_params$burnin))
      ## Save the mini.chains
      save(randospace_charadriiformes_, file = paste0("../Data/Cluster/shapespace_charadriiformes/randospace_charadriiformes_", i, ".mini.chains"))
}
```

## BM space

For the BM space, the parameters are estimated for each simulation individually to avoid corner case bugs (e.g. wrong Choleski decompositions!).
We used three chains of 20000 iterations for the parameterisation chains.

```{r, eval = FALSE, echo = FALSE}
## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_charadriiformes_family.rda")

## Select the data
shapespace <- shapespace_charadriiformes_family[[1]]$space
selected_dim <- shapespace_charadriiformes_family[[1]]$dimensions
consensus_tree <- shapespace_charadriiformes_family[[1]]$consensus_tree
trees_list <- shapespace_charadriiformes_family[[1]]$trees_list

for(i in 1:25) {
      ## Create the space
      BM_space <- BMise.space(shapespace, tree, selected_dim)

      ## Set up the parametrising chains models
      param_chains <- make.mini.chains(data         = BM_space,
                                       dimensions   = selected_dim,
                                       tree         = consensus_tree,
                                       trait.family = "gaussian",
                                       randoms      = c("global", "level1"),
                                       residuals    = "global",
                                       priors       = 0.02,
                                       verbose      = TRUE,
                                       parameters   = list(
                                             nitt   = 20000,
                                             thin   = 500,
                                             burnin = 0))

      ## Run the parametrisation chains (local)
      params <- run.mini.chains(param_chains, replicates = 3)

      ## Get the estimated parameters
      est_params <- extract.parameters(params)

      ## The thinning parameter
      thin <- 200
      ## The number of samples per chains
      samples <- 10

      ## Setting up 20 random replicates
      ## Set up the mini-chains models
      BMspace_charadriiformes_ <- make.mini.chains(
                                      data         = BM_space,
                                      dimensions   = selected_dim,
                                      tree         = trees_list[sample(1:1000, 400)],
                                      trait.family = "gaussian",
                                      randoms      = c("global", "level1"),
                                      residuals    = "global",
                                      priors       = est_params$priors,
                                      verbose      = TRUE,
                                      parameters   = list(
                                            nitt   = est_params$burnin + thin * samples,
                                            thin   = thin,
                                            burnin = est_params$burnin))
      ## Save the mini.chains
      save(BMspace_charadriiformes_, file = paste0("../Data/Cluster/shapespace_charadriiformes/BMspace_charadriiformes_", i, ".mini.chains"))
}
```

## Ball space

Since this space is not stochastic, we used the normal procedure for setting up the mini-chains.
We used three chains of 20000 iterations for the parameterisation chains.

```{r, eval = FALSE, echo = FALSE}
## Set up the parametrising chains models
param_chains <- make.mini.chains(data         = ballspace,
                                 dimensions   = selected_dim,
                                 tree         = consensus_tree,
                                 trait.family = "gaussian",
                                 randoms      = c("global", "level1"),
                                 residuals    = "global",
                                 priors       = 0.02,
                                 verbose      = TRUE,
                                 parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

# Running the three MCMCglmm models
save(param_chains, file = "../Data/Processed/params_ballspace_charadriiformes.rda")
```

```{r, eval = FALSE, echo = FALSE}
## Run the parametrisation chains (local)
params <- run.mini.chains(param_chains, replicates = 3)

## Get the estimated parameters
est_params <- extract.parameters(params)
```

```{r, eval = FALSE, echo = FALSE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10
## Set up the mini-chains models
ballspace_charadriiformes <- make.mini.chains(
                                data         = ballspace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 400)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                      nitt   = est_params$burnin + thin * samples,
                                      thin   = thin,
                                      burnin = est_params$burnin))
## Save the mini.chains
save(ballspace_charadriiformes, file = "../Data/Cluster/shapespace_charadriiformes/ballspace_charadriiformes.mini.chains")
```

## Running the mini-chains

The mini-chains were run for 100 replicates to have 1000 exploitable posterior VCV results for each model.
The following table summarises the modeel parameters, number of trees and spaces sampled (because the randomspace and the BMspace are stochastic we generated 25 of them) and number of explotable VCVs

Model       | parameters      | replicates | trees sampled | spaces sampled | posterior VCVs |
------------|-----------------|------------|---------------|----------------|----------------|
shapespace  | from shapespace | 100        | 100           | 1              | 1000 |
randomspace | from shapespace | 100        | 100           | 25             | 1000 |
BMspace     | from each BMspace | 100        | 100           | 25             | 1000 |
ballspace   | from ballspace  | 100        | 100           | 1              | 1000 |

## Combining and diagnosing the results

```{r, eval = TRUE, echo = FALSE}
## Set up the variables
models_path <- "../Data/Cluster/shapespace_charadriiformes/"
model_files <- list.files(models_path, pattern = ".rda")

## Original space
shapespace_models_names <- model_files[grep("shapespace", model_files)]
## Models placeholder
shapespace_models <- list()

## Load the models
for(i in 1:length(shapespace_models_names)) {
      load(paste0(models_path, shapespace_models_names[i], collapse = "/"))
      shapespace_models <- c(shapespace_models, shapespace_20)
}

## Combine and diagnose
shapespace_charadriiformes_posteriors <- combine.mini.chains(shapespace_models)
silent <- diagnose.mini.chains(shapespace_charadriiformes_posteriors)
save(shapespace_charadriiformes_posteriors, file = "../Data/Processed/shapespace_charadriiformes_posteriors.MCMCglmm")

```

Figure 6: Original shapespace mcmcmcglmmm diagnosis

```{r, eval = TRUE, echo = FALSE}
## Ball space
ballspace_models_names <- model_files[grep("ballspace", model_files)]
## Models placeholder
ballspace_models <- list()

## Load the models
for(i in 1:length(ballspace_models_names)) {
      load(paste0(models_path, ballspace_models_names[i], collapse = "/"))
      ballspace_models <- c(ballspace_models, ballspace_20)
}

## Combine and diagnose
ballspace_charadriiformes_posteriors <- combine.mini.chains(ballspace_models)
silent <- diagnose.mini.chains(ballspace_charadriiformes_posteriors)
save(ballspace_charadriiformes_posteriors, file = "../Data/Processed/ballspace_charadriiformes_posteriors.MCMCglmm")
```

Figure 7: Balled shapespace mcmcmcglmmm diagnosis

```{r, eval = TRUE, echo = FALSE}
## Random space
randospace_models_names <- model_files[grep("randospace", model_files)]
## Models placeholder
randospace_models <- list()

## Load the models
for(i in 1:length(randospace_models_names)) {
      load(paste0(models_path, randospace_models_names[i], collapse = "/"))
      randospace_models <- c(randospace_models, unlist(randospace_20, recursive = FALSE))
}

## Combine and diagnose
randospace_charadriiformes_posteriors <- combine.mini.chains(randospace_models)
silent <- diagnose.mini.chains(randospace_charadriiformes_posteriors)
save(randospace_charadriiformes_posteriors, file = "../Data/Processed/randospace_charadriiformes_posteriors.MCMCglmm")
```

Figure 8: Randomised shapespace mcmcmcglmmm diagnosis

```{r, eval = FALSE, echo = FALSE}
## Random space
BMspace_models_names <- model_files[grep("BMspace", model_files)]
## Models placeholder
BMspace_models <- list()

## Load the models
for(i in 1:length(BMspace_models_names)) {
      load(paste0(models_path, BMspace_models_names[i], collapse = "/"))
      BMspace_models <- c(BMspace_models, unlist(BMspace_20, recursive = FALSE))
}

## Combine and diagnose
BMspace_charadriiformes_posteriors <- combine.mini.chains(BMspace_models)
silent <- diagnose.mini.chains(BMspace_charadriiformes_posteriors)
save(BMspace_charadriiformes_posteriors, file = "../Data/Processed/BMspace_charadriiformes_posteriors.MCMCglmm")
```

<!-- Figure 9: BMised shapespace mcmcmcglmmm diagnosis -->

# Analysing the results

## Creating the dispRity objects

```{r, eval = TRUE, echo = FALSE}
## Get the shapespace
load(file = "../Data/Processed/params_shapespace_charadriiformes.rda")
shapespace <- param_chains$data
## Get the posteriors
load(file = "../Data/Processed/shapespace_charadriiformes_posteriors.MCMCglmm")

## Make the dispRity object
chara_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
                                     data          = shapespace,
                                     posteriors    = shapespace_charadriiformes_posteriors,
                                     group         = MCMCglmm.levels(shapespace_charadriiformes_posteriors)[1:4],
                                     rename.groups = c(levels(shapespace$level1), "phylogeny"))
covar.plot(chara_shapespace, ellipses = TRUE, major.axes = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"), legend = TRUE)
```

```{r, eval = TRUE, echo = FALSE}
## Get the shapespace
load(file = "../Data/Processed/params_ballspace_charadriiformes.rda")
shapespace <- param_chains$data
## Get the posteriors
load(file = "../Data/Processed/ballspace_charadriiformes_posteriors.MCMCglmm")

## Make the dispRity object
chara_ballspace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
                                    data          = shapespace,
                                    posteriors    = ballspace_charadriiformes_posteriors,
                                    group         = MCMCglmm.levels(ballspace_charadriiformes_posteriors)[1:4],
                                    rename.groups = c(levels(shapespace$level1), "phylogeny"))
covar.plot(chara_ballspace, ellipses = TRUE, major.axes = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"), legend = TRUE)
```

```{r, eval = TRUE, echo = FALSE}
## Get the shapespace
load(file = "../Data/Cluster/shapespace_charadriiformes/randospace_charadriiformes_1.mini.chains")
shapespace <- randospace_charadriiformes_$data
## Get the posteriors
load(file = "../Data/Processed/randospace_charadriiformes_posteriors.MCMCglmm")

## Make the dispRity object
chara_randospace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
                                     data          = shapespace,
                                     posteriors    = randospace_charadriiformes_posteriors,
                                     group         = MCMCglmm.levels(randospace_charadriiformes_posteriors)[1:4],
                                     rename.groups = c(levels(shapespace$level1), "phylogeny"))
covar.plot(chara_randospace, ellipses = TRUE, major.axes = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"), legend = TRUE)
```

```{r, eval = FALSE, echo = FALSE}
## Get the shapespace
load(file = "../Data/Cluster/shapespace_charadriiformes/BMspace_charadriiformes_1.mini.chains")
shapespace <- BMspace_charadriiformes_$data
## Get the posteriors
load(file = "../Data/Processed/BMspace_charadriiformes_posteriors.MCMCglmm")

## Make the dispRity object
chara_BMspace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
                                  data          = shapespace,
                                  posteriors    = BMspace_charadriiformes_posteriors,
                                  group         = MCMCglmm.levels(BMspace_charadriiformes_posteriors)[1:4],
                                  rename.groups = c(levels(shapespace$level1), "phylogeny"))
covar.plot(chara_BMspace, ellipses = TRUE, major.axes = TRUE, n = 100, col = c("orange", "blue", "darkgreen", "grey"), legend = TRUE)
```