### 2022-08-12 Analysis/05-Figure_with_phylogeny.Rmd
Added phylo figure repro

---
title: "Bird beak's elaboration and innovation"
author: "Gavin Thomas"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

## Plot the median elaboration/innovation on the phylogeny

```{r, echo = FALSE}
library(ggtree)
library(ggtreeExtra)
library(treeio)
library(treeplyr)
library(ggplot2)
library(phytools)
library(motmot)
library(viridis)
library(RColorBrewer)
library(colorspace)
library(ggnewscale)

### All birds plot - main text

# Load data for all phylo, superorder, and order comparisons
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_superorder.rda")
load("../Data/Processed/ploting_data_order.rda")

# Get clade info data (includes colours for clade labelling)
load("../Data/Processed/clade_info.rda")
load("../Data/Processed/tip_colours_orders.rda")

tree <- ploting_data$consensus_tree
tree <- ladderize(tree)

# Set up data frame - with some missing data will need to use full_join
beak_all <-  data.frame(species = names(ploting_data$median_elaborations), elab=ploting_data$median_elaborations, inov=ploting_data$median_innovations, phylogeny=ploting_data$distances_to_centre)
beak_superorder <- data.frame(species = names(ploting_data_superorder$median_elaborations), elabsuperorder=ploting_data_superorder$median_elaborations,  inovsuperorder=ploting_data_superorder$median_innovations)
beak_order <- data.frame(species = names(ploting_data_order$median_elaborations), elaborder=ploting_data_order$median_elaborations,  inovorder=ploting_data_order$median_innovations)

beak <- left_join(beak_all, beak_superorder, by="species")
beak <- left_join(beak, beak_order, by="species")

# Combine data and tree to plot estimated ancestral states for distance to centroid per species
beak_tree <- make.treedata(tree=tree, data= beak)
dcen_plot <- getVector(beak_tree, phylogeny)
fit <- phytools::fastAnc(beak_tree$phy, dcen_plot, vars=TRUE,CI=FALSE)

# Combine ancestral state estimates with tip values and names 
td <- data.frame(node = ggtree::nodeid(beak_tree$phy, names(dcen_plot)),
               dist_cent = dcen_plot)
nd <- data.frame(node = names(fit$ace), dist_cent = fit$ace)
d <- rbind(td, nd)
d$node <- as.numeric(d$node)

# Join data frame with tree
birdelabtree <- full_join(beak_tree$phy, d, by = "node")
birdelabtree@phylo <- ladderize(birdelabtree@phylo)

# The basic plotting call (1) including angle (5) 
innovelab_plot <- ggtree(birdelabtree, aes(color= dist_cent), layout="fan",
                      continuous = TRUE, size=0.5, ladderize=TRUE, open.angle=20) +
  scale_color_gradientn(colours=cividis(100))  +
  theme(legend.position="none", legend.box="horizontal")

# Set up data in long format with fudge to get variable names to plot correctly
heatmapdat <- beak[,c(1:3,5:8)]
colnames(heatmapdat) <- c("species", "Elaboration (all)", "Innovation (all)", 
                      "Elab (superorder)", "Innov (superorder)",
                      "Elab (order)", "Innov (order)")

heatmapdat <- reshape2::melt(heatmapdat, id.vars="species", variable.name = "type", value.name="score", factorsAsStrings=FALSE)

# Make data frame with node labels and node ids
n <- length(clade_info[[2]])
order_labels <- data.frame(order_label=rep(NA, n), node=rep(NA, n), colour_bar=rep(NA, n))

for (i in 1:n) {
  order_labels[i,1] <- names(clade_info[[2]][i])
  order_labels[i,2] <- MRCA(birdelabtree, clade_info[[2]][[i]][[1]])
}

# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
order_labels <-arrange(order_labels, order_label)
order_labels[,3] <- arrange(data.frame(colour=tip_colours_orders, orders=names(tip_colours_orders)), orders)[,1]

## Second part of plot - add heatmap for elaboration and innovation scores
innovelab_plot <- innovelab_plot +
  new_scale_fill() + 
  geom_fruit(data=heatmapdat, geom=geom_tile, 
             mapping=aes(y=species, x=type, fill=score),
             axis.params=list(
               axis="x",
               text.angle = -90,
               text.size = 4.5,
               line.size = 0,
               vjust = 0.5,
               hjust=0),
             offset = 0.02,size = 0.02) +
  scale_fill_continuous_divergingx(palette = 'Spectral', rev=TRUE,
                                   mid = 1, na.value="grey80") + 
  theme(legend.position=c(0.52,0.43), legend.key.size = unit(0.5, "cm"), 
        legend.key.width=unit(0.6, "cm")) + 
  labs(fill= "Elab & Innov") +
  labs(colour= "Distance")

## Third part of plot - add clade labels
innovelab_plot <- innovelab_plot +
  new_scale_colour() + 
  geom_cladelab(data=order_labels, 
                mapping=aes(node=node, 
                            label=order_label,
                            colour=order_label),
                textcolour="black",
                hjust=0,
                angle="auto",
                barsize=4,
                horizontal=TRUE, 
                fontsize=4,
                offset=30,
                offset.text=2,
                show.legend=FALSE
  ) +
  scale_colour_manual(
    values=order_labels$colour_bar, 
    guide="none"
    )

## Save as png and pdf
ggsave("../Manuscript/Figures/InnovElabTree_main_text.png", innovelab_plot_final, width=12, height=12)
ggsave("../Manuscript/Figures/InnovElabTree_main_text.pdf", innovelab_plot_final, width=12, height=12)

#####################################################################
#####################################################################

### Passerines supplementary plot

# Load data for all phylo, superorder, and order comparisons
load("../Data/Processed/ploting_data_passeriformes.rda")
load("../Data/Processed/ploting_data_passeriformes_suborder.rda")
load("../Data/Processed/ploting_data_passeriformes_family.rda")

# Get clade info data (includes colours for clade labelling)
load("../Data/Processed/clade_info_passeriformes.rda")
load("../Data/Processed/tip_colours_passeriformes_family.rda")

clades <- read.csv("../Data/Processed/BirdClades.csv")
passerines <- clades$TipLabel[which(clades$level2_order=="Passeriformes")]

tree_pass <- ploting_data_passeriformes$consensus_tree
tree_pass <- ladderize(tree_pass)

# Set up data frame - with some missing data will need to use full_join
beak_pass <-  data.frame(species = names(ploting_data_passeriformes$median_elaborations), elab=ploting_data_passeriformes$median_elaborations, inov=ploting_data_passeriformes$median_innovations, phylogeny=ploting_data_passeriformes$distances_to_centre)
beak_pass_suborder <- data.frame(species = names(ploting_data_passeriformes_suborder$median_elaborations), elabsuperorder=ploting_data_passeriformes_suborder$median_elaborations,  inovsuperorder=ploting_data_passeriformes_suborder$median_innovations)
beak_pass_fam <- data.frame(species = names(ploting_data_passeriformes_family$median_elaborations), elaborder=ploting_data_passeriformes_family$median_elaborations,  inovorder=ploting_data_passeriformes_family$median_innovations)

beak_pass <- left_join(beak_pass, beak_pass_suborder, by="species")
beak_pass <- left_join(beak_pass, beak_pass_fam, by="species")

# Ensure only passerines included
beak_pass <- subset(beak_pass, species %in% passerines)

# Combine data and tree to plot estimated ancestral states for distance to centroid per species
beak_tree_pass <- make.treedata(tree=tree_pass, data= beak_pass)
dcen_plot_pass <- getVector(beak_tree_pass, phylogeny)
fit_pass <- phytools::fastAnc(beak_tree_pass$phy, dcen_plot_pass, vars=TRUE,CI=FALSE)

# Combine ancestral state estimates with tip values and names 
td_pass <- data.frame(node = nodeid(beak_tree_pass$phy, names(dcen_plot_pass)),
                 dist_cent = dcen_plot_pass)
nd_pass <- data.frame(node = names(fit_pass$ace), dist_cent = fit_pass$ace)
d_pass <- rbind(td_pass, nd_pass)
d_pass$node <- as.numeric(d_pass$node)

# Join data frame with tree
birdelabtree_pass <- full_join(beak_tree_pass$phy, d_pass, by = "node")
birdelabtree_pass@phylo <- ladderize(birdelabtree_pass@phylo)

# The basic plotting call (1) including angle (5) 
innovelab_pass_plot <- ggtree(birdelabtree_pass, aes(color= dist_cent), layout="fan",
                         continuous = TRUE, size=0.5, ladderize=TRUE, open.angle=25) +
  scale_color_gradientn(colours=cividis(100))  +
  theme(legend.position="none", legend.box="horizontal")

# Set up data in long format with fudge to get variable names to plot correctly
heatmapdat_pass <- beak_pass[,c(1:3,5:8)]
colnames(heatmapdat_pass) <- c("species", "Elaboration (all)", "Innovation (all)", 
                          "Elab (suborder)", "Innov (suborder)",
                          "Elab (family)", "Innov (family)")

heatmapdat_pass <- reshape2::melt(heatmapdat_pass, id.vars="species", variable.name = "type", value.name="score", factorsAsStrings=FALSE)

# Make data frame with node labels and node ids
n_pass <- length(clade_info_passeriformes[[2]])
family_labels_pass <- data.frame(family_label=rep(NA, n_pass), node=rep(NA, n_pass), colour_bar=rep(NA, n_pass))

for (i in 1:n_pass) {
  family_labels_pass[i,1] <- names(clade_info_passeriformes[[2]][i])
  family_labels_pass[i,2] <- ggtree::MRCA(birdelabtree_pass, clade_info_passeriformes[[2]][[i]][[1]])
  family_labels_pass[i,3] <- clade_info_passeriformes[[2]][[i]][[2]]
}

# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
family_labels_pass <-arrange(family_labels_pass, family_label)
family_labels_pass[,3] <- arrange(data.frame(colour=tip_colours_passeriformes_family, family_label=names(tip_colours_passeriformes_family)), family_label)[,1]

## Second part of plot - add heatmap for elaboration and innovation scores
innovelab_pass_plot <- innovelab_pass_plot +
  new_scale_fill() + 
  geom_fruit(data=heatmapdat_pass, geom=geom_tile, 
             mapping=aes(y=species, x=type, fill=score),
             axis.params=list(
               axis="x",
               text.angle = -90,
               text.size = 4.5,
               line.size = 0,
               vjust = 0.5,
               hjust=0),
             offset = 0.02,size = 0.02) +
  scale_fill_continuous_divergingx(palette = 'Spectral', rev=TRUE,
                                   mid = 1, na.value="grey80") + 
  theme(legend.position=c(0.56,0.43), legend.key.size = unit(0.5, "cm"), 
        legend.key.width=unit(0.6, "cm")) + 
  labs(fill= "Elab & Innov") +
  labs(color= "Distance")

## Third part of plot - add clade labels
innovelab_pass_plot <- innovelab_pass_plot +
  new_scale_colour() + 
  geom_cladelab(data=family_labels_pass, 
                mapping=aes(node=node, 
                            label=family_label,
                            colour=family_label),
                textcolour="black",
                hjust=0,
                angle="auto",
                barsize=4,
                horizontal=TRUE, 
                fontsize=4,
                offset=18.5,
                offset.text=1.5,
                show.legend=FALSE
  ) +
  scale_colour_manual(
    values=family_labels_pass$colour_bar, 
    guide="none"
  )

## Save as png and pdf
ggsave("../Manuscript/Figures/InnovElabTree_passerine_supplement.png", innovelab_pass_plot, width=12, height=12)
ggsave("../Manuscript/Figures/InnovElabTree_passerine_supplement.pdf", innovelab_pass_plot, width=12, height=12)

```