### 2022-08-23 Analysis/01-Data_preparation.Rmd
Update 01-Data_preparation.Rmd

---
title: "Bird beak's elaboration and exploration: data preparation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

<!-- Compilation time: 30 minutes -->

All the custom code and custom functions called in this document can be downloaded and installed in `R`.

```{r packages, eval = TRUE}
## Loading packages
library(dispRity) #v>=1.7
library(xtable)
library(knitr)
source("../Functions/stats.utilities.R")
```

# Data preparation

Trees were manipulated in `R` [@Rcore] using the package `ape` [@ape].

 * The trait space

```{r}
## Reading the different trait spaces
shapespace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")
shapespace <- shapespace$PCscores
formspace <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_FORMSPACE_FULL.rds")
formspace <- formspace$PCscores
```

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
par(mfrow = c(2,1))
plot.dispRity(shapespace, type = "preview", pch = 21, main = "Shape space")
plot.dispRity(formspace, type = "preview", pch = 21, main = "Form space")
```

 * The tree(s)

```{r}
## Reading the consensus tree
consensus_tree <- read.nexus("../Data/Raw/Tree_data/9993taxa_1_tree_Global_MCC_CA.tre")

## Reading the distribution of trees
trees_list <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")
```

 * The clade data
We used the different clade levels (super orders, orders, sub-orders and families). For each level/group, we removed any groups with less than 20 species.

```{r}
## Reading the clades list
clades_list <- as.data.frame(read.csv("../Data/Raw/BirdClades.csv", row.names = 1, stringsAsFactors = TRUE))
## Remove the absent species
clades_list <- clades_list[rownames(clades_list) %in% rownames(shapespace), ]
## Sorting the list by rownames from the shapespace
clades_list <- clades_list[match(rownames(shapespace), rownames(clades_list)), -c(6, 7)]
## Renaming the levels
colnames(clades_list) <- c("superorder", "order", "suborder", "superfamily", "family")

## Cleaning the levels (removing any level with less than 15 elements)
update.level <- function(level, min = 15) {
    to_remove <- which(table(level) < min)
    remove_table <- data.frame("clade" = names(to_remove), "species" =  c(table(level)[to_remove]))
    level[which(level %in% c(names(to_remove)))] <-  ""
    return(list(level = as.factor(as.character(level)), removed = remove_table))
}

## Updated list
update_clade_list <- lapply(as.list(clades_list), update.level, min = 15)
clade_list <- data.frame(lapply(update_clade_list, `[[`, 1))
removed_table <- do.call(rbind, lapply(update_clade_list, function(x) {rownames(x$removed) <- NULL; return(x$removed)}))

## Saving the clade list
save(clade_list, file = "../Data/Processed/clade_list.rda")
```

```{r, echo = FALSE}
library(knitr)
kable(removed_table, caption = "List of removed clades by level and number of species in that clade in the dataset.", label = "Table X1")
```

### Selecting the number of axes

How many dimensions do we need to represent at least 95% of the variance in each group

```{r}
## Temporary change the empty levels to noclade
clades <- clades_list[, c(1)]
levels(clades)[1] <- "no clade"

## Make a dispRity object and get axes data
shape_axes <- select.axes(custom.subsets(shapespace, group = clades))
form_axes  <- select.axes(custom.subsets(formspace, group = clades))

pdf(file = "../Manuscript/Figures/axis_selection.pdf", height = 8, width = 12)
plot(shape_axes)
dev.off()
```

```{r, echo = FALSE}
kable(summary(shape_axes)[, c(1:((length(shape_axes$dimensions)+1)*2))], caption = "variance per axes for the Shape space", label = "Table X2")
to_export <- summary(shape_axes)[, c(1:((length(shape_axes$dimensions)+1)*2))]
rownames(to_export)[nrow(to_export)] <- "Entire space"
silent <- make.xtable(to_export, caption = "Variance per axis and per group in the shape space. The \"no clade\" group contains birds not attributed to any of the other groups (i.e. from a clade with less than 15 species).", label = "tab_variance_per_axis", path = "../Manuscript/Tables/", include.rownames = TRUE)
```

```{r, echo = FALSE}
kable(summary(form_axes)[, c(1:((length(form_axes$dimensions)+1)*2))], caption = "variance per axes for the Form space", label = "Table X3")
```

```{r, echo = FALSE, fig.height = 18, fig.width = 18}
plot(shape_axes)
```
Fig X1: variance per axes per superorder and order in the shapespace

```{r, echo = FALSE, fig.height = 18, fig.width = 18}
plot(form_axes)
```
Fig X1: variance per axes per superorder and order in the formspace

By analysing the distribution of the variance within each group, we will need `r length(shape_axes$dimensions)` dimensions for the shape space and `r length(form_axes$dimensions)` for the form space.

### Saving the data

Here we are going to group the spaces, trees and levels for subsequent levels (nested).
There are 5 levels: Super order (1), Order (2), Sub-order (3), Super families (4) and families (5).
The script above creates the model for the two first levels, the subsequent models create the mini-chains for the following overlapping pairs of models (order + sub-order; sub-order + super families; super families + families).

```{r}
source("../Functions/prep.data.R")
## Getting the data ready for all birds
shapespace_allbirds_lvl_superorder_order <- prep.data(
                                          level = 0, lvl.inc = c(1,2),
                                          clades    = clades_list,
                                          space     = shapespace,
                                          dim       = shape_axes$dimensions,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")

## Getting the data ready for each bird super order
shapespace_superorder_lvl_order_suborder <- prep.data(
                                          level = 1, lvl.inc = c(2,3),
                                          clades    = clades_list,
                                          space     = shapespace,
                                          dim       = shape_axes$dimensions,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_superorder_lvl_order_suborder, file = "../Data/Processed/shapespace_superorder_lvl_order_suborder.rda")

## Getting the data ready for each bird order
shapespace_order_lvl_suborder_superfam <- prep.data(
                                          level = 2, lvl.inc = c(3,4),
                                          clades    = clades_list,
                                          space     = shapespace,
                                          dim       = shape_axes$dimensions,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_order_lvl_suborder_superfam, file = "../Data/Processed/shapespace_order_lvl_suborder_superfam.rda")

## Getting the data ready for each bird sub order
shapespace_suborder_lvl_superfam_fam <- prep.data(
                                          level = 3, lvl.inc = c(4,5),
                                          clades    = clades_list,
                                          space     = shapespace,
                                          dim       = shape_axes$dimensions,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_suborder_lvl_superfam_fam, file = "../Data/Processed/shapespace_suborder_lvl_superfam_fam.rda")

## Getting the data ready for each bird super family
shapespace_superfam_lvl_fam <- prep.data(
                                          level = 4, lvl.inc = 5,
                                          clades    = clades_list,
                                          space     = shapespace,
                                          dim       = shape_axes$dimensions,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_superfam_lvl_fam, file = "../Data/Processed/shapespace_superfam_lvl_fam.rda")

## Get a list of all selected levels for the record (and number of elements per levels)
super_orders     <- shapespace_allbirds_lvl_superorder_order[[1]]$levels[[1]]
orders           <- shapespace_allbirds_lvl_superorder_order[[1]]$levels[[2]]
sub_orders       <- lapply(shapespace_order_lvl_suborder_superfam, function(x) x$levels[[1]])
names(sub_orders) <- NULL
sub_orders       <- unlist(sub_orders, recursive = FALSE)
super_families   <- lapply(shapespace_order_lvl_suborder_superfam, function(x) x$levels[[2]])
names(super_families) <- NULL
super_families   <- unlist(super_families, recursive = FALSE)
families         <- lapply(shapespace_superfam_lvl_fam, function(x) x$levels[[1]])
names(families) <- NULL
families         <- unlist(families, recursive = FALSE)

## Get the list of levels
levels_list <- list("super_orders"   = unlist(lapply(super_orders, length)),
                    "orders"         = unlist(lapply(orders, length)),
                    "sub_orders"     = unlist(lapply(sub_orders, length)),
                    "super_families" = unlist(lapply(super_families, length)),
                    "families"       = unlist(lapply(families, length)))
save(levels_list, file = "../Data/Processed/levels_list.rda")
```

## Passeriformes focus

```{r}
source("../Functions/prep.data.R")
## Get the traitspace for all passeriformes
load("../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
clades_list <- as.data.frame(read.csv("../Data/Raw/BirdClades.csv", row.names = 1, stringsAsFactors = TRUE))
shapespace_passeriformes <- shapespace_allbirds_lvl_superorder_order[[1]]$space[shapespace_allbirds_lvl_superorder_order[[1]]$space$level2 == "Passeriformes", ]

## Loading the Passeriformes clades
load("../Data/Processed/clade.info_passerines.rda")
## Clean the levels
shapespace_passeriformes$level1 <- rep("", nrow(shapespace_passeriformes))
shapespace_passeriformes$level2 <- rep("", nrow(shapespace_passeriformes))

## Fill in the level1
for(clade in 1:length(clade_info_passerines$passerine_big_clades)) {
  ## Get the clade name
  clade_name <- names(clade_info_passerines$passerine_big_clades)[clade]
  ## Attribute the species/clade
  shapespace_passeriformes[clade_info_passerines$passerine_big_clades[[clade]], "level1"] <- clade_name
  ## Remove species not present originally in the shapespace
  if(length(nas <- which(apply(shapespace_passeriformes[, 1:8], 1, function(row) any(is.na(row))))) > 0) {
      shapespace_passeriformes <- shapespace_passeriformes[-nas, ]
  }
}

## Fill in the level2
for(clade in 1:length(clade_info_passerines$passerine_small_clades)) {
  ## Get the clade name
  clade_name <- names(clade_info_passerines$passerine_small_clades)[clade]
  ## Attribute the species/clade
  shapespace_passeriformes[clade_info_passerines$passerine_small_clades[[clade]], "level2"] <- clade_name
  ## Remove species not present originally in the shapespace
  if(length(nas <- which(apply(shapespace_passeriformes[, 1:8], 1, function(row) any(is.na(row))))) > 0) {
      shapespace_passeriformes <- shapespace_passeriformes[-nas, ]
  }
}
shapespace_passeriformes[, "level1"] <- as.factor(shapespace_passeriformes[, "level1"])
shapespace_passeriformes[, "level2"] <- as.factor(shapespace_passeriformes[, "level2"])

## Do the prep data from here
shapespace_passeriformes_two_levels <- prep.data(
                                          level = 0, lvl.inc = c(1,2),
                                          clades    = shapespace_passeriformes[, c("level1", "level2")],
                                          space     = shapespace_passeriformes,
                                          dim       = 1:8,
                                          consensus = consensus_tree,
                                          trees     = trees_list,
                                          verbose   = TRUE)
save(shapespace_passeriformes_two_levels, file = "../Data/Processed/shapespace_passeriformes_two_levels.rda")
```

## References