### 2022-10-17 Analysis/11-variance_partitioning.Rmd
Fromaly added a variance test part

---
title: "Bird beak's elaboration and exploration"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script is a working script to run the chains on the SHARC Sheffield cluster.

```{r load_packages, eval = TRUE, message = FALSE}
set.seed(42)
## Loading beer
library(mcmcmcglmmm)
library(MCMCglmm)
library(dispRity)

## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")

## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
selected_dim <- shapespace_allbirds_lvl_superorder_order[[1]]$dimensions
consensus_tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
trees_list <- shapespace_allbirds_lvl_superorder_order[[1]]$trees_list
```

# All birds shapespace analyses

## Mini-chains parametrization

Making a 2D model with just the super order random terms

```{r, parametrise_the_model, eval = FALSE}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = shapespace,
                                   dimensions   = selected_dim,
                                   tree         = consensus_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "level1", "level2"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))
save(param_MCMCglmm, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
```

## Running the parametrisation chains (for the small model)

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k_chain1.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k_chain2.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k_chain3.rda")
end <- Sys.time()
end - start
```

## (for the big model)

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k_chain1.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k_chain2.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k_chain3.rda")
end <- Sys.time()
end - start
```

From these three chains, we first extracted the three burnin periods (defined as the generation when the chain reaches the median likelihood value times 1.25).
We then extracted the posterior estimates of the R-Structure, G-Structure and fixed effects (mean and covariance) (ignoring the runs from the previously estimated burnin periods) to serve as our priors for our mini-chains. 
For those values, we took the highest burnin as the overall burnin and the median of the three posteriors as the overall priors.

```{r, get_model_parameters, eval = FALSE}
## Load the computed models
param_chains <- list()
for(i in 1:3) {
    load(paste0("../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k_chain", i, ".rda"))
    param_chains[[i]] <- params_shapespace_2D_9levels_20k[[1]]
}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

### Running the mini-chains

```{r, set_the_mini_chains, eval = FALSE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Set up the mini-chains models
shapespace_allbirds_lvl_superorder_order_2D <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = 1:2,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_allbirds_lvl_superorder_order_2D, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order_2D.mini.chains")
```

The chains are then run on the cluster
TODO: details

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order_2D.mini.chains")
## Running the three MCMCglmm models
start <- Sys.time()
shapespace_allbirds_lvl_superorder_order_2D <- run.mini.chains(shapespace_allbirds_lvl_superorder_order_2D, replicates = 20)
save(shapespace_allbirds_lvl_superorder_order_2D, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k_chain1_20.rda")
end <- Sys.time()
end - start
```

### Combining the model data

```{r, eval = FALSE}
## Set up the variables
models_path <- "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/cluster/"
## Getting all the model files
model_files <- list.files(models_path, pattern = ".rda")

## Check if the model exists already
if(!("shapespace_allbirds_lvl_superorder_order.MCMCglmm" %in% list.files("../Data/Processed/"))) {
    ## Preparing the model files
    model_list <- list()
    class(model_list) <- c("beer", "mini.chains")
    ## Get the first model
    load(paste0(models_path, model_files[1], collapse = "/"))
    ## Move it to the storage
    system(paste("mv", paste0(models_path, model_files[1]), paste0(models_path, "done/")))
    print(model_files[1])
    ## Clean the bits that are not going to be used (but are huge)
    model[[1]][c("error.term", "family", "y.additional", "ginverse", "ZR")] <- NULL
    ## Update the model list
    model_list[[length(model_list)+1]] <- model[[1]]
    model <- NULL
    model_files <- model_files[-1]
} else {
    ## Updating the model
    load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
    model_list <- shapespace_allbirds_lvl_superorder_order
}

## Get the subsequent models and merge them as we go
while(length(model_files) != 0) {
    load(paste0(models_path, model_files[1], collapse = "/"))
    ## Move it to storage
    system(paste("mv", paste0(models_path, model_files[1]), paste0(models_path, "done/")))
    print(model_files[1])
    ## Clean the bits that are not going to be used (but are huge)
    model[[1]][c("error.term", "family", "y.additional", "ginverse", "ZR")] <- NULL
    ## Update the model list
    model_list[[length(model_list)+1]] <- model[[1]]
    ## Merge the models (overwrite the model list)
    model_list <- list(combine.mini.chains(model_list))
    ## DEBUG
    print("model list size:")
    print(format(object.size(model_list), units = "MB"))
    model <- NULL
    model_files <- model_files[-1]
}

## Combining the model files
shapespace_allbirds_lvl_superorder_order <- model_list

## Saving the results
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")

```

# Analysing the variance per terms

```{r}
## Get the variance for each level
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = TRUE)
## Rename the columns
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")
reorder_clades <- c(9:35,1:8, 36, 37)
# ## Pull by term
# variance_per_levels<- cbind("orders" = apply(variance_per_clades[, 9:35], 1, sum),
#                             "super-orders" = apply(variance_per_clades[, 1:8], 1, sum),
#                             "phylogeny" = variance_per_clades[, 36],
#                             "residuals" = variance_per_clades[, 37])
# variance_per_terms <- cbind("clades" = apply(variance_per_clades[, 1:35], 1, sum),
#                             "phylogeny" = variance_per_clades[, 36],
#                             "residuals" = variance_per_clades[, 37])
# ## Plot the results
# pdf(file = "../Manuscript/Figures/variance_per_levels_whole_tree.pdf", height = 8, width = 14)
# par(mar = c(5,8,2,2), mfrow = c(1, 3))
# # boxplot(variance_per_terms, horizontal = TRUE, las = 1, main = "Scaled variance by terms", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# # boxplot(variance_per_levels, horizontal = TRUE, las = 1, main = "Scaled variance by levels", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(variance_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled variance by clades", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(log(variance_per_clades[,reorder_clades])/log(sample_sizes), horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing sample size effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(log(variance_per_clades[,reorder_clades])/log(observed_variance)[names(variances)], horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing variance effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# dev.off()

# ## Getting the variances per clades
# variances <- apply(variance_per_clades[,reorder_clades], 2, mean)

# ## Getting the sample sizes per clades (the one for the residuals is the whole tree size)
# sample_sizes <- c("residuals" = size.subsets(order_shapespace)[28],
#                   size.subsets(superorder_shapespace),
#                   size.subsets(order_shapespace))[names(variances)]

# ## Measuring the variance for each clade and for the residuals
# ## the obs variance of the residuals is just the variance in the whole dataset (phylogeny)
# observed_variance <- c("residuals" = unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[28], 
#                        unlist(get.disparity(dispRity(superorder_shapespace, metric = sd))),
#                        unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[-28])^2

# plot(log(variances)/log(sample_sizes), log(sample_sizes), xlab = "log variance / log n", ylab = "log n")

```

## Standard error per terms

```{r}

## Get the number of observations per groups
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = FALSE)
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")

## Converting into standard errors
get.se <- function(var, n) {
    sqrt(var[[1]])/sqrt(n)
}
## Get the se per group
se_per_clades <- mapply(get.se, apply(variance_per_clades, 2, list), as.list(c(size.subsets(superorder_shapespace)[-9], size.subsets(order_shapespace), "residuals" = 8748)))
## Scale that
se_per_clades <- se_per_clades/rowSums(se_per_clades)
## Get the other levels
se_per_levels<- cbind("orders" = apply(se_per_clades[, 9:35], 1, sum),
                      "super-orders" = apply(se_per_clades[, 1:8], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])
se_per_terms <- cbind("clades" = apply(se_per_clades[, 1:35], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])

## Plot the results
pdf(file = "../Manuscript/Figures/se_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
boxplot(se_per_terms, horizontal = TRUE, las = 1, main = "Scaled standard error by terms", xlab = "Scaled se ")
boxplot(se_per_levels, horizontal = TRUE, las = 1, main = "Scaled standard error by levels", xlab = "Scaled se")
boxplot(se_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled standard error by clades", xlab = "Scaled variance")
dev.off()
```