### 2022-11-04 Analysis/09-Figure_key.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This section contains the code for reproducing the analyses. To get the details, refer to 03-elaboration_innovation_analyses.Rmd.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
source("../Functions/ploting.utilities.R")
source("../Functions/covar.utilities.R")
```

# All birds shapespace analyses

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Creating the subsets
superorder_shapespace <- MCMCglmm.subsets(n = 4000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)
```

# Extract only the Mirandornithes

```{r, eval = TRUE, echo = FALSE}
## Selecting the paleo data
space_paleo <- get.subsets(superorder_shapespace, subsets = c("phylogeny", "Paleognathae"))

## PANEL 1

## General plot
pdf("../Manuscript/Figures/figure_key_space_plot.pdf")
plot(shapespace[, c(1,2)], col = "lightgrey", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Plotting the projection cartoon

## Do the projection
matrix <- as.matrix(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)])
rownames(matrix) <- NULL
## Get the base vectors
covars <- sample.n(space_paleo$covar, 1000)
centres <- lapply(unlist(space_paleo$subsets, recursive = FALSE) ,function(group, data, fun) fun(data[c(group), ]), data = space_paleo$matrix[[1]], fun = colMeans)
covars <- mapply(recentre, covars, centres, MoreArgs = list(dimensions = c(1,2)), SIMPLIFY = FALSE)
covars <- lapply(covars, function(one_covar, scale) mapply(scale.VCV, one_covar, scale, SIMPLIFY = FALSE), scale = covars[[2]])
covars_cent_tend <- lapply(covars, VCV.cent.tend, mean)
base_vectors <- lapply(lapply(covars_cent_tend, get.one.axis, axis = 1, level = 0.95, dimensions = c(1,2)), list)

## Get projection lines
get.proj.line <- function(point, base) {
  point <- t(point)
  base <- t(base)
  proj_x <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[1];
  proj_y <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[2];
  return(rbind(cbind(proj_x, proj_y), point))
}

## Projection onto the phylo line
pdf("../Manuscript/Figures/figure_key_projection_on_phylo.pdf")
plot(NULL, xlab = "PC1", ylab = "PC2", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
# lines(base_vector)
for(i in 1:nrow(matrix)) {
      lines(get.proj.line(matrix[i, ], base_vectors$phylogeny[[1]]))
}
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Projection onto the group line
pdf("../Manuscript/Figures/figure_key_projection_on_group.pdf")
plot(NULL, xlab = "PC1", ylab = "PC2", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)

base <- base_vectors$Paleognathae[[1]]
base <- matrix(c(base)[c(1, 1, 4, 4)], 2, 2, byrow = FALSE) # or c(1, 4, 3, 2)
for(i in 1:nrow(matrix)) {
      lines(get.proj.line(matrix[i, ], test))
}

points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Phylo plot
## Get the tree
tree_paleo <- drop.tip(tree, tip = rownames(shapespace)[which(shapespace$level1 != "Paleognathae")])
## Ladderize
tree_paleo <- ladderize(tree_paleo)

## Get the data (in the right order)
load("../Data/Processed/ploting_data.rda")
elaborations <- ploting_data$median_elaborations[tree_paleo$tip.label]
innovations <- ploting_data$median_innovations[tree_paleo$tip.label]

## Transform them into a colour gradient
col.grad <- colorRamp(rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#66c2a5", "#3288bd")))
cols_elab <- col.grad(elaborations/max(elaborations))
cols_innov <- col.grad(innovations/max(innovations))

## Plot the tree
pdf("../Manuscript/Figures/figure_key_tree_example.pdf")
plot(tree_paleo, show.tip.label = FALSE, type = "fan")
tiplabels("", frame = "none", pch = 19, cex = 2, col = rgb(cols_elab, maxColorValue = 256), offset = -1)
tiplabels("", frame = "none", pch = 19, cex = 2, col = rgb(cols_innov, maxColorValue = 256), offset = 4)
dev.off()

```