### 2023-05-03 Analysis/13-mcmcmcglmmm.ppca.Rmd
Create 13-mcmcmcglmmm.ppca.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

# Comparing results of `mcmcmcglmmm` to phylo-pca results

## Comparing phylo pca to just pca

```{r}
library(phytools)
data(anoletree)
data(anole.data)

## Phylo PCA
phyl_pca <- phyl.pca(anoletree, anole.data)

## Normal PCA
norm_pca <- prcomp(anole.data)

## Comparing the loadings
scree_phyl <- apply(phyl_pca$S, 2, var)/sum(apply(phyl_pca$S, 2, var))
scree_norm <- apply(norm_pca$x, 2, var)/sum(apply(norm_pca$x, 2, var))

## Anolis example