### 2023-05-11 Analysis/04-Figure_with_ellipse__alternative.Rmd
Update 04-Figure_with_ellipse__alternative.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script reproduces the figure 2 in the main manuscript (the one with the ellipses) and takes around 5 minutes to compile.
Note that the published figure is slightly modified in Inkscape (added headers and separation lines).

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
source("../Functions/ploting.utilities.R")
source("../Functions/covar.utilities.R")
```

# All birds clade level projections plot

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Creating the subsets
superorder_shapespace <- MCMCglmm.subsets(n = 100,
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

order_shapespace <- MCMCglmm.subsets(n = 100,
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Loading the results
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
results <- shapespace_allbirds_lvl_superorder_order_results_list
rm(shapespace_allbirds_lvl_superorder_order_results_list)
```

```{r, eval = TRUE, echo = FALSE}
## Matching colours (fast utility)
match.col <- function(data, col) {
    return(c(col[match(names(size.subsets(data)), names(col))[-length(size.subsets(data))]], "grey"))
}
```

```{r}
pdf(file = "../Manuscript/Figures/figure_phylo_spiro.pdf", height = 11.693, width = 8.268)
## Plot splitting
x <- 82
super_orders <- rbind(
      #"Paleognathae" = "Tinamiformes",  
      matrix(c(rep(1+x,  1), rep(2+x,  1), rep(3+x,  1)), 1, 3),
      #"Galloanserae" = c("Galliformes", "Anseriformes"),
      matrix(c(rep(4+x,  1), rep(5+x,  1), rep(6+x,  1)), 1, 3),
      matrix(0, 1, 3),
      #"Mirandornithes" = c("Podicipediformes"),
      matrix(c(rep(7+x,  1), rep(8+x,  1), rep(9+x,  1)), 1, 3),
      #"Columbimorphae" = c("Columbiformes", "Pterocliformes"),
      matrix(c(rep(10+x, 1), rep(11+x, 1), rep(12+x, 1)), 1, 3),
      matrix(0, 1, 3),
      #"Otidimorphae" = c("Otidiformes", "Cuculiformes", "Musophagiformes"),
      matrix(0, 1, 3),
      matrix(c(rep(13+x, 1), rep(14+x, 1), rep(15+x, 1)), 1, 3),
      matrix(0, 1, 3),
      #"Gruiformes" (NOT A SUPER ORDER)
      matrix(0, 1, 3),
      #"Aequornithes" = c("Procellariiformes", "Sphenisciformes", "Ciconiiformes", "Pelecaniformes", "Suliformes"),
      matrix(0, 2, 3),
      matrix(c(rep(16+x, 1), rep(17+x, 1), rep(18+x, 1)), 1, 3),
      matrix(0, 2, 3),
      #"Strisores" = c("Caprimulgiformes", "Apodiformes"),
      matrix(c(rep(19+x, 1), rep(20+x, 1), rep(21+x, 1)), 1, 3),
      matrix(0, 1, 3),
      #"Charadriiformes" (NOT A SUPER ORDER)
      matrix(0, 1, 3),
      matrix(0, 4, 3),
      #"Telluraves" = c("Trogoniformes", "Bucerotiformes", "Coraciiformes", "Piciformes", "Strigiformes", "Accipitriformes", "Falconiformes", "Psittaciformes", "Passeriformes")
      matrix(c(rep(22+x, 1), rep(23+x, 1), rep(24+x, 1)), 1, 3),
      matrix(0, 4, 3))

layout_matrix <- cbind( 
                matrix(1, 27, 1), # first column
                matrix(seq(1:(27*3))+1, 27, 3, byrow = TRUE), # second/third column
                super_orders)

widths_vals <- c(5, 1, 1, 1, 1, 1, 1)
height_vals <- rep(1, 27)
layout_spiro <- layout(layout_matrix, widths = widths_vals, heights = height_vals, respect = TRUE)
# layout.show(layout_spiro)

## Plotting the tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

## Plotting the tree and getting the tip orders
tip_colours_orders <- plot.short.tree(tree, shapespace, "level2", tip.order = tip_order)
# save(tip_colours_orders, file = "../Data/Processed/tip_colours_orders.rda")

## Plotting each ellipse
for(i in 1:27) {
    wrap.plot.ellipses(i, tip_colours_orders, order_shapespace, results$order_results_between, ei.col = c("orange", "blue"), phylo.plot = TRUE)
}

#TODO:
 # - change the E/I thermos to scale with the amount of E/I <- maybe?
 # - change the E/I thermos to look less clutered (remove the box and labels), one tick = 5 units

load(file = "../Data/Processed/tip_colours_super_orders.rda")
for(i in 1:8) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_super_orders, superorder_shapespace, results$super_results_between, ei.col = c("orange", "blue"), phylo.plot = TRUE)
}
dev.off()

# ## Plotting the ellipses for each order
# for(i in 1:length(tip_order)) {
#     ## Plot each pannel
#     wrap.plot.ellipses(i, tip_colours_super_orders, superorder_shapespace, results$super_results_between, ei.col = c("orange", "blue"))
# }

```

### Ellipses plot super orders

```{r, eval = TRUE, echo = FALSE}
## Plot dimensions and name
box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses_super_order.pdf", height = box_size*4, width = box_size*8)

## New layout for super_order panel (separately)
lay_mat_super_2 <- rbind(
    matrix(c(1,1, rep(18, 3), 0, 2:3), ncol = 8, nrow = 1),
    matrix(c(1,1, rep(18, 3), 0, 4:5), ncol = 8, nrow = 1),
    matrix(c(6:13), ncol = 8, nrow = 1),
    matrix(c(14:17, rep(0, 4)), ncol = 8, nrow = 1))
widths_vals <- rep(c(1, 1/2), 4)
heights_vals <- rep(1, 4)
layout_super <- layout(lay_mat_super_2, widths = widths_vals, heights = heights_vals, respect = TRUE)
# layout.show(layout_super)

## PLOTTING THE SUPERORDERS

## Getting the tip order for the ladderised super order tree
tip_order <- c(
    "Paleognathae" = 1,
    "Galloanserae" = 2,
    "Otidimorphae" = 5,
    "Aequornithes" = 6,
    "Mirandornithes" = 3,
    "Columbimorphae" = 4,
    "Telluraves" = 8,
    "Strisores" = 7)

## Plotting the tree and getting the tip orders
tip_colours_super_orders <- plot.short.tree(tree, shapespace, "level1", tip.order = tip_order)
save(tip_colours_super_orders, file = "../Data/Processed/tip_colours_super_orders.rda")

## Plotting the ellipses for each order
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_super_orders, superorder_shapespace, results$super_results_between, ei.col = c("orange", "blue"))
}

## Adding the ellipses
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(superorder_shapespace, ellipses = mean,
           col = match.col(superorder_shapespace, tip_colours_super_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(1, length(tip_colours_super_orders)), 3),
           add = TRUE, n = 100)
dev.off()
```

### Ellipses plot orders

```{r, eval = TRUE, echo = FALSE}
## Plot dimensions and name
box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses_order.pdf", height = box_size*7, width = box_size*10)

## New layout for order panel (separately)
lay_mat_order_2 <- rbind(
        matrix(c(1,1, rep(56, 3), 0, 2:5), ncol = 10, nrow = 1), 
        matrix(c(1,1, rep(56, 3), 0, 6:9), ncol = 10, nrow = 1), 
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1),
        matrix(c(18:27), ncol = 10, nrow = 1),
        matrix(c(28:37), ncol = 10, nrow = 1), 
        matrix(c(38:47), ncol = 10, nrow = 1),
        matrix(c(48:55, rep(0, 2)), ncol = 10, nrow = 1))
widths_vals <- rep(c(1, 1/2), 5)
heights_vals <- rep(1, 7)
layout_order <- layout(lay_mat_order_2, widths = widths_vals, heights = heights_vals, respect = TRUE)
# layout.show(layout_order)

## PLOTTING THE SUPERORDERS

## Getting the tip order for the ladderised super order tree

## Getting the tip order for the ladderised super order tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

## Plotting the tree and getting the tip orders
tip_colours_orders <- plot.short.tree(tree, shapespace, "level2", tip.order = tip_order)
save(tip_colours_orders, file = "../Data/Processed/tip_colours_orders.rda")

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_orders, order_shapespace, results$order_results_between, ei.col = c("orange", "blue"))
}

## Adding the ellipses
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(order_shapespace, ellipses = mean,
           col = match.col(order_shapespace, tip_colours_orders),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(0.5, length(tip_colours_orders)), 3),
           add = TRUE, n = 100)
dev.off()

```

# Passeriformes clade level projections plot

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_passeriformes_two_levels.rda")
## Select the data
shapespace <- shapespace_passeriformes_two_levels[[1]]$space
tree <- shapespace_passeriformes_two_levels[[1]]$consensus_tree
## Clean garbage
rm(shapespace_passeriformes_two_levels)

## Load the covar
load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
shapespace_passeriformes <- shapespace_passeriformes[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_passeriformes) == "animal")
suborder_lvl <- grep("level1_", MCMCglmm.levels(shapespace_passeriformes))
family_lvl <- grep("level2_", MCMCglmm.levels(shapespace_passeriformes))

## Creating the subsets
suborder_shapespace <- MCMCglmm.subsets(
      data          = shapespace,
      posteriors    = shapespace_passeriformes,
      group         = MCMCglmm.levels(shapespace_passeriformes)[c(suborder_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

family_shapespace <- MCMCglmm.subsets(
      data          = shapespace,
      posteriors    = shapespace_passeriformes,
      group         = MCMCglmm.levels(shapespace_passeriformes)[c(family_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_passeriformes)

## Loading the results
load(file = "../Data/Processed/shapespace_passeriformes_suborder_family_results_list.rda")
results <- shapespace_passeriformes_suborder_family_results_list
rm(shapespace_passeriformes_suborder_family_results_list)
```

## Phylogeny window (detailed)

```{r}
## Adding the bar plot for the phylogeny component (to be added separately in Inkscape)
# pdf(file = "../Manuscript/Figures/phylo_loadings.pdf", height = 4, width = 4)
# plot(NULL, xlim = range(suborder_shapespace$matrix), ylim = range(suborder_shapespace$matrix))
# add.dims(suborder_shapespace, name.col = c(phylogeny = "grey"), n = 1000)
# dev.off()

box_size <- 1.5
pdf(file = "../Manuscript/Figures/ellipses_passeriformes.pdf", height = box_size*9, width = box_size*10)

## Global window
x <- 61
lay_mat_sub <- rbind(
        matrix(c(1,1, rep(x, 3), 0, 2:5), ncol = 10, nrow = 1), # line 7
        matrix(c(1,1, rep(x, 3), 0, 6:9), ncol = 10, nrow = 1), # line 8
        matrix(c(x+2,x+2, 10:13, rep(x+3, 4)), ncol = 10, nrow = 1)) # line 8)
y <- x - 13
lay_mat_family <- rbind(
        matrix(c(1,1, rep(y+1, 3), y+4, 2:5), ncol = 10, nrow = 1), 
        matrix(c(1,1, rep(y+1, 3), y+4, 6:9), ncol = 10, nrow = 1), 
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1),
        matrix(c(18:27), ncol = 10, nrow = 1),
        matrix(c(28:37), ncol = 10, nrow = 1), 
        matrix(c(38:47), ncol = 10, nrow = 1))
heights_vals <- rep(1, 8)
widths_vals <- rep(c(1, 1/3), 5)
layout_matrix <- layout(rbind(lay_mat_sub, (lay_mat_family+13)), widths = widths_vals, heights = heights_vals, respect = TRUE)
# layout.show(layout_matrix)

## PLOTTING THE SUBORDERS
## Getting the tip order for the ladderised super order tree
tip_order <- c(
    "Suboscines" = 1,
    "Meliphagoidea" = 2,
    "Corvides" = 3,
    "Sylviida" = 4,
    "Muscicapida" = 5,
    "Passerida" = 6)

## Plotting the tree and getting the tip orders
tip_colours_passeriformes_suborder <- plot.short.tree(tree, shapespace, "level1", tip.order = tip_order)
save(tip_colours_passeriformes_suborder, file = "../Data/Processed/tip_colours_passeriformes_suborder.rda")

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_passeriformes_suborder, suborder_shapespace, results$suborder_results_between)
}

## PLOTTING THE FAMILIES
## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Aegithaloidea"= 12,
    "Bombycilloidea"= 17,
    "Cisticolidae"= 14,
    "Corvoidea"= 7,
    "Emberizoidea"= 23,
    "Eurylaimides"= 1,
    "Fringillidae"= 10,
    "Furnariida"= 3,
    "Hirundinidae"= 11,
    "Locustelloidea"= 15,
    "Malaconotoidea"= 5,
    "Meliphagoidea"= 4,
    "Motacillidae"= 22,
    "Muscicapoidea"= 18,
    "Nectariniidae"= 19,
    "Orioloidea"= 6,
    "Paridae"= 9,
    "Passeridae"= 21,
    "Petroicidae"= 8,
    "PloceidaeEstrildidae"= 20,
    "Pycnonotidae"= 13,
    "Sylvioidea"= 16,
    "Tyrannida" = 2)

## Plotting the tree and getting the tip orders
tip_colours_passeriformes_family <- plot.short.tree(tree, shapespace, "level2", tip.order = tip_order)
save(tip_colours_passeriformes_family, file = "../Data/Processed/tip_colours_passeriformes_family.rda")

## Plotting the ellipses
for(i in 1:length(tip_order)) {
    ## Plot each pannel
    wrap.plot.ellipses(i, tip_colours_passeriformes_family, family_shapespace, results$family_results_between)
}

## Plotting the aligned and scale ellipses for the superorders
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(suborder_shapespace, ellipses = mean,
           col = match.col(suborder_shapespace, tip_colours_passeriformes_suborder),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(1, length(tip_colours_passeriformes_suborder)), 3),
           add = TRUE, n = 100)

## PLOTTING THE ORDERS
par(bty = "n")
lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = "n", ylab = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
covar.plot(family_shapespace, ellipses = mean,
           col = match.col(family_shapespace, tip_colours_passeriformes_family),
           legend = FALSE, points = FALSE, scale = "phylogeny",
           lines.lwd = c(rep(0.5, length(tip_colours_passeriformes_family)), 3),
           add = TRUE, n = 100)
dev.off()
```