### 2023-05-11 Analysis/05-Figure_with_phylogeny.Rmd
Update 05-Figure_with_phylogeny.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Gavin Thomas"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script reproduces the figure 1 in the main manuscript (the phylogeny) and takes around 15 minutes to compile.
Note that the published figure is slightly modified in Inkscape (boxplot added manually and changed the label position for passeriformes).

# Plot the median elaboration/innovation on the phylogeny

```{r, echo = FALSE}
library(ggtree)
library(ggtreeExtra)
library(treeio)
# install_github("uyedaj/treeplyr")
library(treeplyr)
library(dplyr)
library(ggplot2)
library(phytools)
library(motmot)
library(viridis)
library(RColorBrewer)
library(colorspace)
library(ggnewscale)

### All birds plot - main text

# Load data for all phylo, superorder, and order comparisons
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_superorder.rda")
load("../Data/Processed/ploting_data_order.rda")

# Get clade info data (includes colours for clade labelling)
load("../Data/Processed/clade_info.rda")
load("../Data/Processed/tip_colours_orders.rda")

tree <- ploting_data$consensus_tree
tree <- ladderize(tree)

# Set up data frame - with some missing data will need to use full_join
beak_all <-  data.frame(species = names(ploting_data$median_elaborations), elab=ploting_data$median_elaborations, inov=ploting_data$median_innovations, phylogeny=ploting_data$distances_to_centre)
beak_superorder <- data.frame(species = names(ploting_data_superorder$median_elaborations), elabsuperorder=ploting_data_superorder$median_elaborations,  inovsuperorder=ploting_data_superorder$median_innovations)
beak_order <- data.frame(species = names(ploting_data_order$median_elaborations), elaborder=ploting_data_order$median_elaborations,  inovorder=ploting_data_order$median_innovations)

beak <- left_join(beak_all, beak_superorder, by="species")
beak <- left_join(beak, beak_order, by="species")

# Combine data and tree to plot estimated ancestral states for distance to centroid per species
beak_tree <- make.treedata(tree=tree, data= beak)
dcen_plot <- getVector(beak_tree, phylogeny)
fit <- phytools::fastAnc(beak_tree$phy, dcen_plot, vars=TRUE,CI=FALSE)

# Combine ancestral state estimates with tip values and names 
td <- data.frame(node = ggtree::nodeid(beak_tree$phy, names(dcen_plot)),
               dist_cent = dcen_plot)
nd <- data.frame(node = names(fit$ace), dist_cent = fit$ace)
d <- rbind(td, nd)
d$node <- as.numeric(d$node)

# Join data frame with tree
birdelabtree <- full_join(beak_tree$phy, d, by = "node")
birdelabtree@phylo <- ladderize(birdelabtree@phylo)

# The basic plotting call (1) including angle (5) 
innovelab_plot <- ggtree(birdelabtree, aes(color= dist_cent), layout="fan",
                      continuous = TRUE, size=0.5, ladderize=TRUE, open.angle=20) +
  scale_color_gradientn(colours=cividis(100))  +
  theme(legend.position="none", legend.box="horizontal")

# Set up data in long format with fudge to get variable names to plot correctly
heatmapdat <- beak[,c(1:3,5:8)]
colnames(heatmapdat) <- c("species", "Elaboration (all)", "Innovation (all)", 
                      "Elab (superorder)", "Innov (superorder)",
                      "Elab (order)", "Innov (order)")
## Just orders
heatmapdat <- beak[,c(1,7,8)]
colnames(heatmapdat) <- c("species", "Elaboration", "Innovation")

heatmapdat <- reshape2::melt(heatmapdat, id.vars="species", variable.name = "type", value.name="score", factorsAsStrings=FALSE)

# Make data frame with node labels and node ids
n <- length(clade_info[[2]])
order_labels <- data.frame(order_label=rep(NA, n), node=rep(NA, n), colour_bar=rep(NA, n))

for (i in 1:n) {
  order_labels[i,1] <- names(clade_info[[2]][i])
  order_labels[i,2] <- MRCA(birdelabtree, clade_info[[2]][[i]][[1]])
}

# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
order_labels <-arrange(order_labels, order_label)
order_labels[,3] <- arrange(data.frame(colour=tip_colours_orders, orders=names(tip_colours_orders)), orders)[,1]

## Second part of plot - add heatmap for elaboration and innovation scores
innovelab_plot <- innovelab_plot +
  new_scale_fill() + 
  geom_fruit(data=heatmapdat, geom=geom_tile, 
             mapping=aes(y=species, x=type, fill=score),
             axis.params=list(
               axis="x",
               text.angle = -90,
               text.size = 4.5,
               line.size = 0,
               vjust = 0.5,
               hjust=0),
             offset = 0.02,size = 0.02) +
  scale_fill_continuous_divergingx(palette = 'Spectral', rev=TRUE,
                                   mid = 1, na.value="grey80") + 
  theme(legend.position=c(0.52,0.43), legend.key.size = unit(0.5, "cm"), 
        legend.key.width=unit(0.6, "cm")) + 
  labs(fill= "Elab & Innov") +
  labs(colour= "Distance")

## Third part of plot - add clade labels
innovelab_plot <- innovelab_plot +
  new_scale_colour() + 
  geom_cladelab(data=order_labels, 
                mapping=aes(node=node, 
                            label=order_label,
                            colour=order_label),
                textcolour="black",
                hjust=0,
                angle="auto",
                barsize=4,
                horizontal=TRUE, 
                fontsize=4,
                offset=30,
                offset.text=2,
                show.legend=FALSE
  ) +
  scale_colour_manual(
    values=order_labels$colour_bar, 
    guide="none"
    )

## Save as png and pdf
ggsave("../Manuscript/Figures/InnovElabTree_main_text.png", innovelab_plot, width=12, height=12)
ggsave("../Manuscript/Figures/InnovElabTree_main_text.pdf", innovelab_plot, width=12, height=12)

#####################################################################
#####################################################################

### Passerines supplementary plot

# Load data for all phylo, superorder, and order comparisons
load("../Data/Processed/ploting_data_passeriformes.rda")
load("../Data/Processed/ploting_data_passeriformes_suborder.rda")
load("../Data/Processed/ploting_data_passeriformes_family.rda")

# Get clade info data (includes colours for clade labelling)
load("../Data/Processed/clade_info_passeriformes.rda")
load("../Data/Processed/tip_colours_passeriformes_family.rda")

clades <- read.csv("../Data/Raw/BirdClades.csv")
passerines <- clades$TipLabel[which(clades$level2_order=="Passeriformes")]

tree_pass <- ploting_data_passeriformes$consensus_tree
tree_pass <- ladderize(tree_pass)

# Set up data frame - with some missing data will need to use full_join
beak_pass <-  data.frame(species = names(ploting_data_passeriformes$median_elaborations), elab=ploting_data_passeriformes$median_elaborations, inov=ploting_data_passeriformes$median_innovations, phylogeny=ploting_data_passeriformes$distances_to_centre)
beak_pass_suborder <- data.frame(species = names(ploting_data_passeriformes_suborder$median_elaborations), elabsuperorder=ploting_data_passeriformes_suborder$median_elaborations,  inovsuperorder=ploting_data_passeriformes_suborder$median_innovations)
beak_pass_fam <- data.frame(species = names(ploting_data_passeriformes_family$median_elaborations), elaborder=ploting_data_passeriformes_family$median_elaborations,  inovorder=ploting_data_passeriformes_family$median_innovations)

beak_pass <- left_join(beak_pass, beak_pass_suborder, by="species")
beak_pass <- left_join(beak_pass, beak_pass_fam, by="species")

# Ensure only passerines included
beak_pass <- subset(beak_pass, species %in% passerines)

# Combine data and tree to plot estimated ancestral states for distance to centroid per species
beak_tree_pass <- make.treedata(tree=tree_pass, data= beak_pass)
dcen_plot_pass <- getVector(beak_tree_pass, phylogeny)
fit_pass <- phytools::fastAnc(beak_tree_pass$phy, dcen_plot_pass, vars=TRUE,CI=FALSE)

# Combine ancestral state estimates with tip values and names 
td_pass <- data.frame(node = nodeid(beak_tree_pass$phy, names(dcen_plot_pass)),
                 dist_cent = dcen_plot_pass)
nd_pass <- data.frame(node = names(fit_pass$ace), dist_cent = fit_pass$ace)
d_pass <- rbind(td_pass, nd_pass)
d_pass$node <- as.numeric(d_pass$node)

# Join data frame with tree
birdelabtree_pass <- full_join(beak_tree_pass$phy, d_pass, by = "node")
birdelabtree_pass@phylo <- ladderize(birdelabtree_pass@phylo)

# The basic plotting call (1) including angle (5) 
innovelab_pass_plot <- ggtree(birdelabtree_pass, aes(color= dist_cent), layout="fan",
                         continuous = TRUE, size=0.5, ladderize=TRUE, open.angle=25) +
  scale_color_gradientn(colours=cividis(100))  +
  theme(legend.position="none", legend.box="horizontal")

# Set up data in long format with fudge to get variable names to plot correctly
heatmapdat_pass <- beak_pass[,c(1:3,5:8)]
colnames(heatmapdat_pass) <- c("species", "Elaboration (all)", "Innovation (all)", 
                          "Elab (suborder)", "Innov (suborder)",
                          "Elab (family)", "Innov (family)")

heatmapdat_pass <- reshape2::melt(heatmapdat_pass, id.vars="species", variable.name = "type", value.name="score", factorsAsStrings=FALSE)

# Make data frame with node labels and node ids
n_pass <- length(clade_info_passeriformes[[2]])
family_labels_pass <- data.frame(family_label=rep(NA, n_pass), node=rep(NA, n_pass), colour_bar=rep(NA, n_pass))

for (i in 1:n_pass) {
  family_labels_pass[i,1] <- names(clade_info_passeriformes[[2]][i])
  family_labels_pass[i,2] <- ggtree::MRCA(birdelabtree_pass, clade_info_passeriformes[[2]][[i]][[1]])
  family_labels_pass[i,3] <- clade_info_passeriformes[[2]][[i]][[2]]
}

# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
# Order data frame alphabetically according to the taxon labels
# Necessary because geom_cladelabel adds labels alphabetically but plots colours 
# in the order of the provided vector (alphatising ensures plotting order is correct)
family_labels_pass <-arrange(family_labels_pass, family_label)
family_labels_pass[,3] <- arrange(data.frame(colour=tip_colours_passeriformes_family, family_label=names(tip_colours_passeriformes_family)), family_label)[,1]

## Second part of plot - add heatmap for elaboration and innovation scores
innovelab_pass_plot <- innovelab_pass_plot +
  new_scale_fill() + 
  geom_fruit(data=heatmapdat_pass, geom=geom_tile, 
             mapping=aes(y=species, x=type, fill=score),
             axis.params=list(
               axis="x",
               text.angle = -90,
               text.size = 4.5,
               line.size = 0,
               vjust = 0.5,
               hjust=0),
             offset = 0.02,size = 0.02) +
  scale_fill_continuous_divergingx(palette = 'Spectral', rev=TRUE,
                                   mid = 1, na.value="grey80") + 
  theme(legend.position=c(0.56,0.43), legend.key.size = unit(0.5, "cm"), 
        legend.key.width=unit(0.6, "cm")) + 
  labs(fill= "Elab & Innov") +
  labs(color= "Distance")

## Third part of plot - add clade labels
innovelab_pass_plot <- innovelab_pass_plot +
  new_scale_colour() + 
  geom_cladelab(data=family_labels_pass, 
                mapping=aes(node=node, 
                            label=family_label,
                            colour=family_label),
                textcolour="black",
                hjust=0,
                angle="auto",
                barsize=4,
                horizontal=TRUE, 
                fontsize=4,
                offset=18.5,
                offset.text=1.5,
                show.legend=FALSE
  ) +
  scale_colour_manual(
    values=family_labels_pass$colour_bar, 
    guide="none"
  )

## Save as png and pdf
ggsave("../Manuscript/Figures/InnovElabTree_passerine_supplement.png", innovelab_pass_plot, width=12, height=12)
ggsave("../Manuscript/Figures/InnovElabTree_passerine_supplement.pdf", innovelab_pass_plot, width=12, height=12)
```

## Adding a panel showing the differences in elaboration and innovation across the tree

```{r}
source("../Functions/ploting.utilities.R")
colour.palette <- gg.color.hue
## Plotting the novelty for the phylo plot:
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_superorder.rda")
load("../Data/Processed/ploting_data_order.rda")

## Get the values
elaborations_tree  <- ploting_data$median_elaborations
innovations_tree   <- ploting_data$median_innovations
elaborations_super <- ploting_data_superorder$median_elaborations
innovations_super  <- ploting_data_superorder$median_innovations
elaborations_order <- ploting_data_order$median_elaborations
innovations_order  <- ploting_data_order$median_innovations

pdf("../Manuscript/Figures/median_novelty_distributions.pdf", height = 4, width = 8)
par(mfrow = c(1,2), bty = "n")
## Same but boxplot version
boxplot(cbind("Whole tree level"  = c(elaborations_tree, innovations_tree),
              "Super-order level" = c(elaborations_super, innovations_super),
              "Order level"       = c(elaborations_order, innovations_order)),
        xlab = "", ylab = "Elaboration/Innovation", col = colour.palette(5)[1:3])
boxplot(cbind("Elaboration" = c(elaborations_tree, elaborations_super, elaborations_order),
              "Innovation"  = c(innovations_tree, innovations_super, innovations_order)),
        xlab = "", ylab = "Elaboration/Innovation", col = colour.palette(5)[4:5])
dev.off()
```