### 2023-05-11 Manuscript/supplementaries.tex
Update supplementaries.tex

\documentclass[12pt,letterpaper]{article}

%Packages
% \usepackage{textcomp}
% \usepackage{latexsym}
% \usepackage{url}
% \usepackage{epsfig}
% \usepackage{graphicx}
% \usepackage{amssymb}
% \usepackage{amsmath}
% \usepackage{mathtools}
% \usepackage{bm}
% \usepackage{array}
% \usepackage[version=3]{mhchem}
% \usepackage{ifthen}
% \usepackage{caption}
% \usepackage{amsthm}
% \usepackage{amstext}
% \usepackage{enumerate}
% \usepackage[osf]{mathpazo}
% \usepackage{dcolumn}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[osf]{mathpazo}
\usepackage{fullpage}
% \usepackage{float}

\pagenumbering{arabic}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}
%Running head
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\medskip
\begin{center}

\noindent{\Large \bf Innovation and elaboration on the avian tree of life - supplementary materials}
\bigskip

\noindent {\normalsize \sc
Thomas Guillerme$^{1,*}$, 
Jen A. Bright$^{2}$,
Christopher R. Cooney${^1}$,
Emma C. Hughes${^1}$,
Zoe K. Varley${^3}$,
Natalie Cooper$^{3,+}$,
Andrew P. Beckerman$^{1,+}$,
and Gavin H. Thomas$^{1,4,+}$}\\
\noindent {\small \it 
$^1$School of Biosciences, University of Sheffield, Sheffield, S10 2TN, United Kingdom.\\
$^2$School of Natural Science, University of Hull, Hull, HU6 7RX, United Kingdom.\\
$^3$Natural History Museum, Cromwell Road, London, SW7 5BD, United Kingdom.\\
$^4$Bird Group, Department of Life Sciences, the Natural History Museum at Tring, Tring, United Kingdom.\\}

\end{center}
\bigskip
\noindent{*\bf Corresponding author:} \textit{guillert@tcd.ie}; \noindent{+\bf contributed equally.}\\
\vspace{1in}

%Line numbering
\modulolinenumbers[1]
\linenumbers

\noindent (Keywords: macroevolution, elaboration, innovation, variance-covariance matrices, line of least resistance, GLMM)\\

\vspace{1.5in}

\newpage

%---------------------------------------------
%
%       INTRODUCTION
%
%---------------------------------------------

\renewcommand{\thefigure}{S\arabic{figure}} 
 \renewcommand{\thetable}{S\arabic{table}}

\begin{landscape}
\section{Supplementary material 1: glossary}

\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{p{0.2\textwidth}p{0.3\textwidth}p{0.3\textwidth}p{0.1\textwidth}p{0.1\textwidth}}
  \hline
Term in this paper         & definition in this paper & mathematical definition & in macroevolution & in microevoltion\\
  \hline
major axis of variation    & the major evolutionary ``direction'' of traits co-evolution & first eigenvector of the variance-covariance matrix  & allometry & line of genetic least resistance \\
clade                      & a monophyletic group of \textbf{species} & a subset of \textbf{shapespace} & Birds & a group of species or a population\\
species                    & the lowest evolutionary units in our analysis & a row in the \textbf{shapespace} & \textit{Pauxi pauxi} & an individual\\
shapespace                 & all the possible morphological combination (observed or not) & an ordinate and orthonormal matrix & the beak shapespace & a gene space \\
novelty                    & morphological variation (either \textbf{innovation} or \textbf{elaboration}) & position of a \textbf{species} or a \textbf{clade} relative to the \textbf{major axis of variation} & & \\
elaboration                & morphological variation along the \textbf{major axis of variation} & the algebraic projection on the eigenvector & & \\
innovation                 & morphological variation away from the \textbf{major axis of variation} & the algebraic rejection from the eigenvector & & \\
variance-covariance matrix & the multidimensional relation between traits & the variation of the dimension within the \textbf{shapespace} taking into account the distance matrix between the observations (the \textbf{phylogeny} or the \textbf{clades}) & & \\
nested phylogenetic level  & the different \textbf{clades} used in this study (the whole phylogeny $>$ super-orders $>$ orders) & different nested \textbf{variance-covariance matrices} & & \\

\end{tabular}
\caption{Glossary of some specific terms in this manuscript}
\label{tab_glossary}
\end{table}
\end{landscape}

\section{Supplementary material 2: additional results}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/mini-chains_diagram.pdf}
\caption{mcmcmcglmmm: mini-chains MCMCglmm method diagram}
\label{Fig:mcmcmcglmm}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/axis_selection.pdf}
\caption{variance and cumulative variance for each 8 axis per super-order.}
\label{Fig:axes_variance}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/parameters_ESS_all_birds.pdf}
\caption{Histogram of the effective sample sizes (ESS) for each class of parameters in the MCMCglmm model for all birds for the 4000 posteriors. \textbf{solution} is the solution of the model; \textbf{Gcovariance} are the covariances for the random terms (the phylogeny and the clades); \textbf{Rcovariance} are the residuals error term.}
\label{Fig:model_ess_all_birds}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/parameters_ESS_passeriformes.pdf}
\caption{Histogram of the effective sample sizes (ESS) for each class of parameters in the MCMCglmm model for the passeriformes for the 1870 posteriors. \textbf{solution} is the solution of the model; \textbf{Gcovariance} are the covariances for the random terms (the phylogeny and the clades); \textbf{Rcovariance} are the residuals error term.}
\label{Fig:model_ess_passeriformes}
\end{figure}

\begin{figure}[htpb]
\centering
\centering
   \includegraphics[width=0.5\textwidth]{Figures/correlations.pdf}
\caption{Correlation plots of $elaboration_{species}$ and $innovation_{species}$ at different focal levels. \textbf{A}: correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on the whole phylogeny. The colours of the dots correspond to their density in the plot. Blue and yellow colours corresponds to respectively low and high density of species. The coloured quadrant in the back corresponds to $elaboration_{species}$ and $innovation_{species}$ scores higher than 1. The blue and yellow quadrants shows species with noticeable high $elaboration_{species}$ and $innovation_{species}$ respectively. The species in the green quadrant have noticeable high levels of both $elaboration_{species}$ and $innovation_{species}$. \textbf{B}: correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on the whole phylogeny (mega-level projection) but plotted by super-order and coloured by order (see list below). \textbf{C} correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on their super-orders' major axis of variation (macro-level projection). The colours correspond to each order. For the distribution of correlation scores, see Fig. \ref{Fig:fable_correlations}. The orders numbers are:
1 - others (orders with $<$ 15 species)
2 - Corvoidea;
3 - Malaconotoidea;
4 - Orioloidea;
5 - Meliphagoidea;
6 - Bombycilloidea;
7 - Muscicapoidea;
8 - Emberizoidea;
9 - Motacillidae;
10 - Nectariniidae ;
11 - Passeridae;
12 - PloceidaeEstrildidae;
13 - Eurylaimides;
14 - Furnariida;
15 - Tyrannida;
16 - Aegithaloidea;
17 - Cisticolidae;
18 - Fringillidae;
19 - Hirundinidae;
20 - Locustelloidea;
21 - Paridae;
22 - Pycnonotidae;
23 - Sylvioidea;}
\label{Fig:figure_correlations}
\end{figure}

\begin{figure}[htpb]
\centering
\centering
   \includegraphics[width=0.5\textwidth]{Figures/correlations_fable.pdf}
\caption{Posterior distributions of the correlations scores between $elaboration_{species}$ and $innovation_{species}$ as displayed for each super-orders in Fig. \ref{Fig:figure_correlations}. Species' projection onto the whole phylogeny's major axis of phenotypic variance-covariance (VCV) corresponds to the correlation scores for the panels A and B in Fig. \ref{Fig:figure_correlations} and species' projections onto their own super-order's major axis of phenotypic variance-covariance (VCV) corresponds to the correlation scores for the panel C in Fig. \ref{Fig:figure_correlations}}
\label{Fig:fable_correlations}
\end{figure}

\clearpage
\section{Supplementary material 3: Passeriformes results}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=0.9\textwidth]{Figures/InnovElabTree_passerine_supplement.pdf}
\caption{Passeriformes phylogeny coloured by beak shape innovation and elaboration level. See the caption of figure 1 in the main text for details.}
\label{fig_phylogeny_passeriformes}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/ellipses_passeriformes.pdf}
\caption{Elaboration and innovation at the group level for each passeriformes sub-orders and families. See the caption of figure 2 in the main text for details.}
\label{fig_ellipses_passeriformes}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/orthogonality_results_passeriformes.pdf}
\caption{Posterior results showing the orientation of phenotypic evolution (orthogonality) for each passeriformes sub-orders and families. See the caption of figure 3 in the main text for details.}
\label{fig_orthogonality_passeriformes}
\end{figure}

\newpage

\section{Supplementary materials 4: detailed projection and rejection operations}
\label{supp_projection}

The following section contains the detailed definition and procedure of how we measured elaboration and innovation

We can define the major axis from the variance-covariance matrices and then project and reject each element of interest in the space onto this axis.
The following steps are generalised to $n$ dimensions and detailed below (as well as the algorithm used in \texttt{dispRity} \cite{dispRity} to perform the transformations):

\subsection{Defining the major axis}

For $O_{n}$, the unit hypersphere matrix of $n$ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation}

In other words, $O_{n}$ is the matrix representing the hypersphere of $n$ dimensions and of radius $1$ that fits in the centre of the trait-space;

And $O'_{n}$ is the scaled matrix hypersphere to the 95\% confidence interval size using the $\chi^2$ distribution:

$$O'_{n} = O_{n} \sqrt{\chi^2(0.95)}$$

Then, for the variance-covariance matrix $VCV_{n}$ of $n$ dimensions obtained from the posterior distribution of the \texttt{mcmcmcglmmmm}:

\begin{equation}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation}

and the eigenvectors \textbf{v} and the eigenvalues $\lambda$ satisfying the following eigen decomposition:

$$VCV_{n} \textbf{v} = \lambda \textbf{v}$$

We can get $M_{n}$, the matrix containing all the edge coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors \textbf{v} and the product of the scaled 0.95 CI unit sphere $O'_{n}$ and the eigenvalues $\lambda$:

$$M_{n} = [(O'_{n}\sqrt{\lambda}) \times v]^{\text{T}}$$

Finally, we can centre the matrix $M_{n}$ on the estimated solution of each GLMM (\texttt{model\$Sol}, in \texttt{MCMCglmm} \cite{MCMCglmm}) corresponding to the estimation of the position of the variance-covariance matrix in the trait-space.
$M_{n}$ then contains all the major axes of the 0.95 hyper-ellipse fitting the variance-covariance matrix.
We can then define the first row of the matrix, $M_{1,n}$, as the major axis, the second row, $M_{2,n}$, as the second major axis (the minor axis in a 2D ellipse), etc.

The detailed procedure was adapted from \href{https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331}{Zheyuan Li's} post on Stack Overflow and implemented in \texttt{dispRity::axis.covar}.

\subsection{Measuring projection and rejection}

Once we have defined a major axis, we can project any elements in the trait-space onto that axis.
Specifically, for any elements in the trait-space, we can define it as the vector $\vec{a}$ with one set of coordinates in $n$ dimensions:

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x \\
    y \\
    \cdots \\
    n \\
    \end{bmatrix}
\end{equation}

And for any major axis that we can define as a vector $\vec{b}$ as a set of pairs of coordinates in $n$ dimensions:

\begin{equation}
    \vec{b} = 
    \begin{bmatrix}
    x_{1} & x_{2} \\
    y_{1} & y_{2} \\
    \cdots & \cdots \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

We can then calculate $\vec{a_{1}}$, the orthogonal projection of $\vec{a}$ onto $\vec{b}$ using:

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\vec{b}\|}
\end{equation}

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\sqrt{\vec{b} \cdot \vec{b}}\|}
\end{equation}

With $\|\vec{b}\| = \sqrt{\vec{b} \cdot \vec{b}}$ being the norm of $\vec{b}$.
And $\vec{a_{2}}$, the rejection of $\vec{a}$ onto $\vec{b}$:

\begin{equation}
    \vec{a_{2}} = \vec{a} - \vec{a_{1}}
\end{equation}

\subsubsection{Generalisation of projection onto any vector in a set space}

Using this, we can generalise the procedure so as to calculate the projection and rejection for any element within a trait-space $TS_{m,n}$:

\begin{equation}
    TS_{m,n} = 
    \begin{bmatrix}
    x_{1} & x_{2} & \cdots & x_{m} \\
    y_{1} & y_{2} & \cdots & y_{m} \\
    \vdots  & \vdots  & \ddots & \vdots \\
    n_{1} & n_{2} & \cdots & n_{m} \\
    \end{bmatrix}
\end{equation}

And any major axis defined as a vector $\vec{B}$:

\begin{equation}
    B = 
    \begin{bmatrix}
    x_{1} & x_{2}\\
    y_{1} & y_{2}\\
    \vdots  & \vdots  \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

By using the linear transformation $f_{\vec{B}}$ of the trait-space $TS$ moving $\vec{B}$ onto $TS$'s first axis unit vector $\vec{\hat{\imath}}$:

$$f_{\vec{B}}(TS) = \left( \frac{TS - [Bx_{1}, By_{1}, \cdots, Bn_{1}]^{\text{T}}}{\|\vec{B}\|} \right) \cdot R_{\vec{B}}$$

With $R_{\vec{B}}$ being the rotation matrix of the vector $\vec{B}$ onto $\vec{\hat{\imath}}$:

\begin{equation}
R_{\vec{B}} = I_{\vec{B}} - \vec{B}\vec{B}^\text{T} - \vec{\hat{\imath}}\vec{\hat{\imath}}^\text{T} + [\vec{B} \vec{\hat{\imath}}]
    \begin{bmatrix}
        cos(\theta) & -sin(\theta)\\
        sin(\theta) & cos(\theta)\\
    \end{bmatrix} [\vec{B} \vec{\hat{\imath}}]^\text{T}
\end{equation}

Where $\theta$ is:

\begin{equation}
    \theta = acos \left(\frac{\vec{B} \cdot \vec{\hat{\imath}}}{\|\vec{B}\| \cdot \|\vec{\hat{\imath}}\|} \right)
\end{equation}

Or $\theta = acos (B_x)$ since both $\|\vec{B}\|$ and $\|\vec{\hat{\imath}}\|$ are equal to 1 and $\|\vec{\hat{\imath}}\|$ is the unit vector on the first axis.

\subsection{Algorithm for calculating the projection/rejection of any element in a defined space}

In practice we followed \href{https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions}{this procedure} and applied a modification of \href{https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385}{this implementation} (see \cite{aguilera2004} for the formal generalisation of this algorithm in $n$ dimensions) using the following algorithm implemented in \texttt{dispRity::projections}:

\begin{enumerate}
 \item In the trait-space, define $\vec{B}$ as the base vector (typically $\vec{B}$ is defined as the pair of coordinates from the major axis described above).
 \item Centre the trait-space on the origin of $\vec{B}$ so that the first set of coordinates of $\vec{B}$ are 0.
 \item Scale the trait-space to the norm of $\vec{B}$ so that the norm of $\vec{B}$ is now 1.
 \item Rotate the trait-space using the rotation matrix $R_{\vec{B}}$ to satisfy the linear transformation $\vec{B} \rightarrow \vec{\hat{\imath}}$ (with $\vec{\hat{\imath}}$ being the first unit vector of the trait-space - typically the x axis unit vector). 
 \item Project/reject every element in the trait-space on $\vec{B}$ (that is now $\vec{\hat{\imath}}$). In practice, the first coordinate (x) of each element is now its projection onto $\vec{B}$.
\end{enumerate}

\bibliographystyle{naturemag}
\bibliography{references}

\end{document}