### 2023-05-16 Manuscript/supplementaries.tex
Update supplementaries.tex

\documentclass[12pt,letterpaper]{article}

%Packages
% \usepackage{textcomp}
% \usepackage{latexsym}
% \usepackage{url}
% \usepackage{epsfig}
% \usepackage{graphicx}
% \usepackage{amssymb}
% \usepackage{amsmath}
% \usepackage{mathtools}
% \usepackage{bm}
% \usepackage{array}
% \usepackage[version=3]{mhchem}
% \usepackage{ifthen}
% \usepackage{caption}
% \usepackage{amsthm}
% \usepackage{amstext}
% \usepackage{enumerate}
% \usepackage[osf]{mathpazo}
% \usepackage{dcolumn}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[osf]{mathpazo}
\usepackage{fullpage}
% \usepackage{float}

\pagenumbering{arabic}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}
%Running head
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\medskip
\begin{center}

\noindent{\Large \bf Innovation and elaboration on the avian tree of life - supplementary materials}
\bigskip

\noindent {\normalsize \sc
Thomas Guillerme$^{1,*}$, 
Jen A. Bright$^{2}$,
Christopher R. Cooney${^1}$,
Emma C. Hughes${^1}$,
Zoe K. Varley${^3}$,
Natalie Cooper$^{3,+}$,
Andrew P. Beckerman$^{1,+}$,
and Gavin H. Thomas$^{1,4,+}$}\\
\noindent {\small \it 
$^1$School of Biosciences, University of Sheffield, Sheffield, S10 2TN, United Kingdom.\\
$^2$School of Natural Science, University of Hull, Hull, HU6 7RX, United Kingdom.\\
$^3$Natural History Museum, Cromwell Road, London, SW7 5BD, United Kingdom.\\
$^4$Bird Group, Department of Life Sciences, the Natural History Museum at Tring, Tring, United Kingdom.\\}

\end{center}
\bigskip
\noindent{*\bf Corresponding author:} \textit{guillert@tcd.ie}; \noindent{+\bf contributed equally.}\\
\vspace{1in}

%Line numbering
\modulolinenumbers[1]
\linenumbers

\noindent (Keywords: macroevolution, elaboration, innovation, variance-covariance matrices, line of least resistance, GLMM)\\

\vspace{1.5in}

\newpage

%---------------------------------------------
%
%       INTRODUCTION
%
%---------------------------------------------

\renewcommand{\thefigure}{S\arabic{figure}} 
 \renewcommand{\thetable}{S\arabic{table}}

\begin{landscape}
\section{Supplementary material 1: glossary}

\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{p{0.2\textwidth}p{0.3\textwidth}p{0.3\textwidth}p{0.1\textwidth}p{0.1\textwidth}}
  \hline
Term in this paper         & definition in this paper & mathematical definition & in macroevolution & in microevoltion\\
  \hline
macroevolutionary scale    & analysis where the focal elements are the species & & \\
megaevolutionary scale     & analysis where the focal elements are clades (i.e. groups of species) & & \\
major axis of variation    & the major evolutionary ``direction'' of traits co-evolution & first eigenvector of the variance-covariance matrix  & allometry & line of genetic least resistance \\
clade                      & a monophyletic group of \textbf{species} & a subset of \textbf{shapespace} & Birds & a group of species or a population\\
species                    & the lowest evolutionary units in our analysis & a row in the \textbf{shapespace} & \textit{Pauxi pauxi} & an individual\\
shapespace                 & all the possible morphological combination (observed or not) & an ordinate and orthonormal matrix & the beak shapespace & a gene space \\
novelty                    & morphological variation (either \textbf{innovation} or \textbf{elaboration}) & position of a \textbf{species} or a \textbf{clade} relative to the \textbf{major axis of variation} & & \\
elaboration                & morphological variation along the \textbf{major axis of variation} & the algebraic projection on the eigenvector & & \\
innovation                 & morphological variation away from the \textbf{major axis of variation} & the algebraic rejection from the eigenvector & & \\
variance-covariance matrix & the multidimensional relation between traits & the variation of the dimension within the \textbf{shapespace} taking into account the distance matrix between the observations (the \textbf{phylogeny} or the \textbf{clades}) & & \\
nested phylogenetic level  & the different \textbf{clades} used in this study (the whole phylogeny $>$ super-orders $>$ orders) & different nested \textbf{variance-covariance matrices} & & \\

\end{tabular}
\caption{Glossary of some specific terms in this manuscript}
\label{tab_glossary}
\end{table}
\end{landscape}

\section{Supplementary material 2: additional results}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/mini-chains_diagram.pdf}
\caption{mcmcmcglmmm: mini-chains MCMCglmm method diagram}
\label{Fig:mcmcmcglmm}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/axis_selection.pdf}
\caption{Variance and cumulative variance for each 8 axis per super-order. Note that not all super orders have their own variance explained matching the variance explained by the entire ordination. For example, Strisores only need the first three dimensions to explain at least 95\% of their variance within the shapespace whilst Aequornithes need 8 dimensions to achieve at least the same amount of variance.}
\label{Fig:axes_variance}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/parameters_ESS_all_birds.pdf}
\caption{Histogram of the effective sample sizes (ESS) for each class of parameters in the MCMCglmm model for all birds for the 4000 posteriors. \textbf{solution} is the solution of the model; \textbf{Gcovariance} are the covariances for the random terms (the phylogeny and the clades); \textbf{Rcovariance} are the residuals error term.}
\label{Fig:model_ess_all_birds}
\end{figure}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/parameters_ESS_passeriformes.pdf}
\caption{Histogram of the effective sample sizes (ESS) for each class of parameters in the MCMCglmm model for the passeriformes for the 1870 posteriors. \textbf{solution} is the solution of the model; \textbf{Gcovariance} are the covariances for the random terms (the phylogeny and the clades); \textbf{Rcovariance} are the residuals error term.}
\label{Fig:model_ess_passeriformes}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/correlations.pdf}
\caption{Correlation plots of $elaboration_{species}$ and $innovation_{species}$ at different focal levels. \textbf{A}: correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on the whole phylogeny. The colours of the dots correspond to their density in the plot. Blue and yellow colours corresponds to respectively low and high density of species. The coloured quadrant in the back corresponds to $elaboration_{species}$ and $innovation_{species}$ scores higher than 1. The blue and yellow quadrants shows species with noticeable high $elaboration_{species}$ and $innovation_{species}$ respectively. The species in the green quadrant have noticeable high levels of both $elaboration_{species}$ and $innovation_{species}$. \textbf{B}: correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on the whole phylogeny (mega-level projection) but plotted by super-order and coloured by order (see list below). \textbf{C} correlation between $elaboration_{species}$ and $innovation_{species}$ for each species when projected on their super-orders' major axis of variation (macro-level projection). The colours correspond to each order. For the distribution of correlation scores, see Fig. \ref{Fig:fable_correlations}. The orders numbers are:
1 - others (orders with $<$ 15 species)
2 - Corvoidea;
3 - Malaconotoidea;
4 - Orioloidea;
5 - Meliphagoidea;
6 - Bombycilloidea;
7 - Muscicapoidea;
8 - Emberizoidea;
9 - Motacillidae;
10 - Nectariniidae ;
11 - Passeridae;
12 - PloceidaeEstrildidae;
13 - Eurylaimides;
14 - Furnariida;
15 - Tyrannida;
16 - Aegithaloidea;
17 - Cisticolidae;
18 - Fringillidae;
19 - Hirundinidae;
20 - Locustelloidea;
21 - Paridae;
22 - Pycnonotidae;
23 - Sylvioidea;}
\label{Fig:figure_correlations}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/correlations_fable.pdf}
\caption{Posterior distributions of the correlations scores between $elaboration_{species}$ and $innovation_{species}$ as displayed for each super-orders in Fig. \ref{Fig:figure_correlations}. Species' projection onto the whole phylogeny's major axis of phenotypic variance-covariance (VCV) corresponds to the correlation scores for the panels A and B in Fig. \ref{Fig:figure_correlations} and species' projections onto their own super-order's major axis of phenotypic variance-covariance (VCV) corresponds to the correlation scores for the panel C in Fig. \ref{Fig:figure_correlations}.}
\label{Fig:fable_correlations}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/Elaboration_and_innovation_clade_through_time.pdf}
\caption{Distribution of $elaboration_{clade}$ and $innovation_{clade}$ for each order and super-order through time.}
\label{Fig:figure_ei_clade_through_time}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/Elaboration_and_innovation_species_through_time.pdf}
\caption{Distribution of $elaboration_{species}$ and $innovation_{species}$ for each order and super-order through time.}
\label{Fig:figure_ei_species_through_time}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/50_random_axes_superorders.pdf}
\caption{Distribution of 50 random scaled major axes (scaled variance-covariance matrices eigen values) for each super order illustrating that individual major axes can span in many different directions with some clades have a higher spread of direction than others.}
\label{Fig:mikado}
\end{figure}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.5\textwidth]{Figures/random_skewers.pdf}
\caption{Random skewers test for the differences between the super orders (in purple) and the orders (in gold) variance-covariance matrices against the whole phylogeny ones. And, in light gold, between the orders and their super-orders variance covariance matrices. The different line types (dashed and full) represent respectively the 95\% and the 75\% CI distribution and the dots represent the median skewer correlation estimate. The stars denote the median p-value estimate (0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1).}
\label{Fig:random_skewers}
\end{figure}

\clearpage
\section{Supplementary material 3: Passeriformes results}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=0.9\textwidth]{Figures/InnovElabTree_passerine_supplement.pdf}
\caption{Passeriformes phylogeny coloured by beak shape innovation and elaboration level. See the caption of figure 1 in the main text for details.}
\label{fig_phylogeny_passeriformes}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/ellipses_passeriformes.pdf}
\caption{Elaboration and innovation at the group level for each passeriformes sub-orders and families. See the caption of figure 2 in the main text for details.}
\label{fig_ellipses_passeriformes}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/orthogonality_results_passeriformes.pdf}
\caption{Posterior results showing the orientation of phenotypic evolution (orthogonality) for each passeriformes sub-orders and families. See the caption of figure 3 in the main text for details.}
\label{fig_orthogonality_passeriformes}
\end{figure}

\newpage

\section{Supplementary materials 4: detailed projection and rejection operations}
\label{supp_projection}

The following section contains the detailed definition and procedure of how we measured elaboration and innovation

We can define the major axis from the variance-covariance matrices and then project and reject each element of interest in the space onto this axis.
The following steps are generalised to $n$ dimensions and detailed below (as well as the algorithm used in \texttt{dispRity} \cite{dispRity} to perform the transformations):

\subsection{Defining the major axis}

For $O_{n}$, the unit hypersphere matrix of $n$ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

\begin{equation}
O_{n} = 
    \begin{pmatrix}
        1 & 0 & \cdots & 0 \\
        0 & 1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & 1 \\
        -1 & 0 & \cdots & 0 \\
        0 & -1 & \cdots & 0 \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        0 & 0 & \cdots & -1 \\
    \end{pmatrix}
\end{equation}

In other words, $O_{n}$ is the matrix representing the hypersphere of $n$ dimensions and of radius $1$ that fits in the centre of the trait-space;

And $O'_{n}$ is the scaled matrix hypersphere to the 95\% confidence interval size using the $\chi^2$ distribution:

$$O'_{n} = O_{n} \sqrt{\chi^2(0.95)}$$

Then, for the variance-covariance matrix $VCV_{n}$ of $n$ dimensions obtained from the posterior distribution of the \texttt{mcmcmcglmmmm}:

\begin{equation}
VCV_{n} = 
    \begin{pmatrix}
        \sigma(a) & \sigma(a,b) & \cdots & \sigma(a,n) \\
        \sigma(a,b) & \sigma(b) & \cdots & \sigma(b,n) \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        \sigma(n,a) & \sigma(n,b) & \cdots & \sigma(n) \\
    \end{pmatrix}
\end{equation}

and the eigenvectors \textbf{v} and the eigenvalues $\lambda$ satisfying the following eigen decomposition:

$$VCV_{n} \textbf{v} = \lambda \textbf{v}$$

We can get $M_{n}$, the matrix containing all the edge coordinates of the 0.95 CI hypersphere from $VCV_{n}$ using the transposition of the cross product between the eigenvectors \textbf{v} and the product of the scaled 0.95 CI unit sphere $O'_{n}$ and the eigenvalues $\lambda$:

$$M_{n} = [(O'_{n}\sqrt{\lambda}) \times v]^{\text{T}}$$

Finally, we can centre the matrix $M_{n}$ on the estimated solution of each GLMM (\texttt{model\$Sol}, in \texttt{MCMCglmm} \cite{MCMCglmm}) corresponding to the estimation of the position of the variance-covariance matrix in the trait-space.
$M_{n}$ then contains all the major axes of the 0.95 hyper-ellipse fitting the variance-covariance matrix.
We can then define the first row of the matrix, $M_{1,n}$, as the major axis, the second row, $M_{2,n}$, as the second major axis (the minor axis in a 2D ellipse), etc.

The detailed procedure was adapted from \href{https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331}{Zheyuan Li's} post on Stack Overflow and implemented in \texttt{dispRity::axis.covar}.

The use of the scaled 95\% confidence interval hyper-sphere matrix $O'_{n}$ allows both (1) to standardise the resulting $M_{n}$ (the scaled version of $VCV_{n}$) so that the unit vector of the resulting space corresponds to the 95\% CI on every dimension (for any dimension $i$ in the matrix $M_{i,n}$, any vector with a norm smaller or greater than 1 is respectively inside or outside the 95\% CI boundary).
And (2) greatly speed up the calculations of projection and rejection in an algorithm (see below) since the hypersphere $O'_{n}$ can be calculated only once for any number of matrix and the projection of rejection values are directly columns in the output transformed matrix (after the projection procedure, see below).
This greatly reduces RAM and CPU usage and scaling for very large number of matrices.

\subsection{Measuring projection and rejection}

Once we have defined a major axis, we can project any elements in the trait-space onto that axis.
Specifically, for any elements in the trait-space, we can define it as the vector $\vec{a}$ with one set of coordinates in $n$ dimensions:

\begin{equation}
    \vec{a} = 
    \begin{bmatrix}
    x \\
    y \\
    \cdots \\
    n \\
    \end{bmatrix}
\end{equation}

And for any major axis that we can define as a vector $\vec{b}$ as a set of pairs of coordinates in $n$ dimensions:

\begin{equation}
    \vec{b} = 
    \begin{bmatrix}
    x_{1} & x_{2} \\
    y_{1} & y_{2} \\
    \cdots & \cdots \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

We can then calculate $\vec{a_{1}}$, the orthogonal projection of $\vec{a}$ onto $\vec{b}$ using:

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\vec{b}\|}
\end{equation}

\begin{equation}
    \vec{a_{1}} = \frac{\vec{a} \cdot \vec{b}}{\|\sqrt{\vec{b} \cdot \vec{b}}\|}
\end{equation}

With $\|\vec{b}\| = \sqrt{\vec{b} \cdot \vec{b}}$ being the norm of $\vec{b}$.
And $\vec{a_{2}}$, the rejection of $\vec{a}$ onto $\vec{b}$:

\begin{equation}
    \vec{a_{2}} = \vec{a} - \vec{a_{1}}
\end{equation}

\subsubsection{Generalisation of projection onto any vector in a set space}

Using this, we can generalise the procedure so as to calculate the projection and rejection for any element within a trait-space $TS_{m,n}$:

\begin{equation}
    TS_{m,n} = 
    \begin{bmatrix}
    x_{1} & x_{2} & \cdots & x_{m} \\
    y_{1} & y_{2} & \cdots & y_{m} \\
    \vdots  & \vdots  & \ddots & \vdots \\
    n_{1} & n_{2} & \cdots & n_{m} \\
    \end{bmatrix}
\end{equation}

And any major axis defined as a vector $\vec{B}$:

\begin{equation}
    B = 
    \begin{bmatrix}
    x_{1} & x_{2}\\
    y_{1} & y_{2}\\
    \vdots  & \vdots  \\
    n_{1} & n_{2} \\
    \end{bmatrix}
\end{equation}

By using the linear transformation $f_{\vec{B}}$ of the trait-space $TS$ moving $\vec{B}$ onto $TS$'s first axis unit vector $\vec{\hat{\imath}}$:

$$f_{\vec{B}}(TS) = \left( \frac{TS - [Bx_{1}, By_{1}, \cdots, Bn_{1}]^{\text{T}}}{\|\vec{B}\|} \right) \cdot R_{\vec{B}}$$

With $R_{\vec{B}}$ being the rotation matrix of the vector $\vec{B}$ onto $\vec{\hat{\imath}}$:

\begin{equation}
R_{\vec{B}} = I_{\vec{B}} - \vec{B}\vec{B}^\text{T} - \vec{\hat{\imath}}\vec{\hat{\imath}}^\text{T} + [\vec{B} \vec{\hat{\imath}}]
    \begin{bmatrix}
        cos(\theta) & -sin(\theta)\\
        sin(\theta) & cos(\theta)\\
    \end{bmatrix} [\vec{B} \vec{\hat{\imath}}]^\text{T}
\end{equation}

Where $\theta$ is:

\begin{equation}
    \theta = acos \left(\frac{\vec{B} \cdot \vec{\hat{\imath}}}{\|\vec{B}\| \cdot \|\vec{\hat{\imath}}\|} \right)
\end{equation}

Or $\theta = acos (B_x)$ since both $\|\vec{B}\|$ and $\|\vec{\hat{\imath}}\|$ are equal to 1 and $\|\vec{\hat{\imath}}\|$ is the unit vector on the first axis.

\subsection{Algorithm for calculating the projection/rejection of any element in a defined space}

In practice we followed \href{https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions}{this procedure} and applied a modification of \href{https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385}{this implementation} (see \cite{aguilera2004} for the formal generalisation of this algorithm in $n$ dimensions) using the following algorithm implemented in \texttt{dispRity::projections}:

\begin{enumerate}
 \item In the trait-space, define $\vec{B}$ as the base vector (typically $\vec{B}$ is defined as the pair of coordinates from the major axis described above).
 \item Centre the trait-space on the origin of $\vec{B}$ so that the first set of coordinates of $\vec{B}$ are 0.
 \item Scale the trait-space to the norm of $\vec{B}$ so that the norm of $\vec{B}$ is now 1.
 \item Rotate the trait-space using the rotation matrix $R_{\vec{B}}$ to satisfy the linear transformation $\vec{B} \rightarrow \vec{\hat{\imath}}$ (with $\vec{\hat{\imath}}$ being the first unit vector of the trait-space - typically the x axis unit vector). 
 \item Project/reject every element in the trait-space on $\vec{B}$ (that is now $\vec{\hat{\imath}}$). In practice, the first coordinate (x) of each element is now its projection onto $\vec{B}$.
\end{enumerate}

\subsection{Definition of species and clade innovation and elaboration}

From the definitions above, we can define $\text{elaboration}_{\text{species}}$ and $\text{innovation}_{\text{species}}$, the elaboration and innovation at the macroevolutionary (species level) and $\text{elaboration}_{\text{clade}}$ and $\text{innovation}_{\text{clade}}$, the elaboration and innovation at the megaevolutionary level (clade).

For any species with coordinates $\vec{species}$ in a $n$ dimensional space, we can estimate $VCV_{clade}$, the variance-covariance matrices representing the clade that contains $\vec{species}$.
We can then calculate the eigen vector and eigenvalues from these matrices as:

$\textbf{v}_{clade}$ defined as:

\begin{equation}
VCV_{clade} \textbf{v}_{clade} = \lambda_{clade} \textbf{v}_{clade}
\end{equation}

In practice, these are solved using a eigen decomposition in \texttt{R} \texttt{base::eigen}.

We can then define the followings for any species $x$:

\begin{equation}
\text{elaboration}_{\text{species}_{x}} = \|\frac{\vec{species}_{x} \cdot \textbf{v}_{clade}} {\sqrt{\textbf{v}_{clade} \cdot \textbf{v}_{clade}}}\|
\end{equation}

With $$\vec{species}_{x}$$ being the vector in $n$ dimensional space for species $x$. And $$\textbf{v}_{clade}$$ being the eigen vector of a specific clade (e.g. $$\textbf{v}_{clade_{c}}$$)

\begin{equation}
\text{innovation}_{\text{species}_{x}} = \| \vec{species}_{x} - \text{elaboration}_{\text{species}_{x}} \|
\end{equation}

And the followings for any clade $c$ (withing a parent clade $c-1$):

\begin{equation}
\text{elaboration}_{\text{clade}_{c}} = \| \frac{\textbf{v}_{clade_{c}} \cdot \textbf{v}_{clade_{c-1}}} {\sqrt{\textbf{v}_{clade_{c-1}} \cdot \textbf{v}_{clade_{c-1}}}}\|
\end{equation}

\begin{equation}
\text{innovation}_{\text{clade}_{c}} = \| \textbf{v}_{clade_{c}} - \text{elaboration}_{\text{clade}_{c}} \|
\end{equation}

\subsection{Relation between elaboration and innovation and metrics in micro/macro-evolution}

\subsubsection{Relative eigenvalues dispersion}

\cite{vanvalen1974,watanabe2022} propose to describe a variance-covariance matrix in terms of its eigen dispersion $V_{rel}$ as:

\begin{equation}
V_{rel} = \frac{\sum_{i}^{n}(\lambda_{i} - \sum{\frac{\lambda_{i}}{n}})^2}{n(n-1)\sum{\frac{\lambda_{i}}{n}}^2}
\end{equation}

Which is the dispersion of eigenvalues (squared difference between each eigen value and the average eigen value) where $\lambda_{i}$ are each individual eigen values and $n$ is the number of dimensions considered.

Using this definition and the definitions above of $\text{elaboration}_{\text{clade}_{c}}$ and $\text{innovation}_{\text{clade}_{c}}$ we can calculate the and plot the relation between elaboration and innovation and their relative eigen dispersion:

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/elaboration_innovation_dispersion.pdf}
\caption{Relation between relative eigen dispersion \cite{vanvalen1974,watanabe2022} and $\text{elaboration}_{\text{clade}_{c}}$ and $\text{innovation}_{\text{clade}_{c}}$ on the \texttt{charadriiformes} example dataset from the \texttt{dispRity} package.}
\label{Fig:eigen_dispersion}
\end{figure}

\subsubsection{Evolvability and respondability}

\cite{hansen2008measuring} propose to measure the evolvability and respondability of an individual as their ``response to selection in $n$ traits''.
Here, some equivalent can be drawn (albeit cautiously) between the response of an individual or a group (population) to selection pressure and a species.
Arguably, a species is the end product of this (and many other) selection pressures on a group of individuals.
If so, then, we can relate $\text{evolvability}$ and $\text{respondability}$ \cite{hansen2008measuring} for a $species_{x}$ to it's $\text{elaboration}_{\text{species}_{x}}$ and $\text{innovation}_{\text{species}_{x}}$ as follows:

\begin{equation}
\text{evolvability}_{\text{species}_{x}} = \text{elaboration}_{\text{species}_{x}}
\end{equation}

\noindent and, using Pythagorean theorem in a triangle rectangle:

\begin{equation}
\text{respondability}_{\text{species}_{x}}  = \sqrt{\text{elaboration}_{\text{species}_{x}}^2 + \text{innovation}_{\text{species}_{x}}^2}
\end{equation}

\begin{figure}[htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/respondability.pdf}
\caption{Relation between respondability \cite{hansen2008measuring} and $\text{elaboration}_{\text{species}}$ and $\text{innovation}_{\text{species}}$ on the \texttt{charadriiformes} example dataset from the \texttt{dispRity} package.}
\label{Fig:respondability}
\end{figure}

Conditional evolvability however will be more complex to determine since it is defined as the relation between evolvability and autonomy which is specific defined in \cite{hansen2008measuring} as the inverse of the \textbf{G} matrix.
In \cite{hansen2008measuring}, the inverse of \textbf{G} is defined as ``the fraction of the genetic variation that is independent of potentially constraining characters'' which in our case will be harder to define.
Although it would technically be possible to use the inverse of the \textbf{R} matrix used here ($VCV_{clade}$), it is unclear what it would represent biologically.

\bibliographystyle{naturemag}
\bibliography{references}

\end{document}