### 2023-06-01 Manuscript/main.tex
Updated introduction. Big time!

\documentclass[12pt,letterpaper]{article}

%Packages
% \usepackage{textcomp}
% \usepackage{latexsym}
% \usepackage{url}
% \usepackage{amssymb}
% \usepackage{amsmath}
% \usepackage{mathtools}
% \usepackage{bm}
% \usepackage{array}
% \usepackage[version=3]{mhchem}
% \usepackage{ifthen}
% \usepackage{amsthm}
% \usepackage{amstext}
% \usepackage{enumerate}
% \usepackage{dcolumn}

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[osf]{mathpazo}
\usepackage{fullpage}
\usepackage{float}
\usepackage{xr} %linking to supplementaries
\externaldocument{supplementaries}

\pagenumbering{arabic}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}
%Running head
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\bigskip
\begin{center}

\noindent{\Large \bf Innovation and elaboration on the avian tree of life}
\bigskip

\noindent {\normalsize \sc
Thomas Guillerme$^{1,*}$, 
Jen A. Bright$^{2}$,
Christopher R. Cooney${^1}$,
Emma C. Hughes${^1}$,
% Zoe K. Varley${^3}$,
Natalie Cooper$^{3,+}$,
Andrew P. Beckerman$^{1,+}$,
and Gavin H. Thomas$^{1,4,+}$}\\
\noindent {\small \it 
$^1$School of Biosciences, University of Sheffield, Sheffield, S10 2TN, United Kingdom.\\
$^2$School of Natural Science, University of Hull, Hull, HU6 7RX, United Kingdom.\\
$^3$Natural History Museum, Cromwell Road, London, SW7 5BD, United Kingdom.\\
$^4$Bird Group, Department of Life Sciences, the Natural History Museum at Tring, Tring, United Kingdom.\\}

\end{center}
\bigskip
\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\ 
\vspace{1in}

%Line numbering
\modulolinenumbers[1]
\linenumbers

%---------------------------------------------
%
%       ABSTRACT
%
%---------------------------------------------

\noindent (Keywords: macroevolution, elaboration, innovation, evolutionary rate matrix, line of least resistance, GLMM)\\

\section{Abstract}

Widely documented, megaevolutionary jumps in phenotypic diversity continue to perplex researchers because it remains unclear whether these dramatic changes can emerge from microevolutionary processes.
Here we tackle this question using new approaches for modeling multivariate traits %@L_multi_traits
 to evaluate the magnitude and distribution of elaboration and innovation in the evolution of bird beaks.
We find that elaboration, evolution along the major axis of phenotypic change, is common at both macro- and megaevolutionary scales whereas innovation, evolution away from the major axis of phenotypic change, is more prominent at megaevolutionary scales.
Indeed, the major axis of phenotypic change among species beak shapes at megaevolutionary scales is an emergent property of innovation across clades. Our analyses suggest that the reorientation of phenotypes via innovation is a ubiquitous route for divergence that can arise through gradual change alone, opening up new avenues for evolution to explore.

\section{Introduction}

Understanding patterns and constraints in the adaptive evolution of species traits is a major goal of evolutionary research. At macroevolutionary scales, renewed attention has been given to variation in the tempo and mode of trait evolution among lineages \cite{venditti2011multiple, khabbazian2016fast, castiglione2018, hansen2022, larouche2022mosaic, pagel2022general, navalon2022environmental, basu2023tempo}.
This has led to growing recognition of the importance of major jumps or discontinuities in evolutionary rates as drivers of diversification of species traits \cite{cooney2017mega, smaers2021evolution, goswami2022}.
For example, in both bird beaks and mammal skulls there are major jumps in the rate of trait evolution early in clade history as morphological diversity expands \cite{cooney2017mega, goswami2022} with lower magnitude jumps towards the present.
Major jumps in species trait values are consistent with the concept of megaevolution, describing major discontinuities in phenotypes observed from the fossil record, first proposed by Simpson.

Megaevolutionary dynamics appear to contrast with theory based on microevolutionary processes.
Over microevolutionary timescales, phenotypic evolution is expected to follow lines of least resistance \cite{schluter1996adaptive}.
The term ‘line of least resistance’ has been used both specifically to describe phenotypic evolution along the axis of greatest genetic variation \cite{schluter1996adaptive}, and more loosely to describe the direction of phenotypic evolution with greatest observable variation \cite{marroig2005size}.
Regardless of the definition used, the microevolutionary line of least resistance represents the direction with the most potential for evolution because it contains the most variation on which selection can act \cite{hansen2016, Rhoda2022, jablonski2022evolvability}.
This expectation is supported by numerous studies showing that phenotypic divergence is often biased along (i.e. aligned with) the line of least resistance \cite{marroig2005size, bolstad2014, walter2018, mcglothlin2018adaptive, tsuboi2018, Rhoda2022, mongle2022}.
These studies suggest that alignment of phenotypic divergence with the line of least resistance may be common over comparatively short (1-2 million year) timescales.
More generally, the alignment of macroevolutionary divergence with the microevolutionary line of least resistance is expected to decline over time \cite{mcglothlin2018adaptive}.
However, there is evidence, for example from Anolis lizards, that stability in the direction of phenotypic evolution can extend over timescales spanning tens of millions of years \cite{mcglothlin2018adaptive}.

The juxtaposition of microevolutionary predictions to macro- and mega-evolutionary observations is striking and raises the question of how phenotypic divergence accumulates across evolutionary scales.
In particular, how can megaevolutionary jumps emerge from gradual micro- or macro-evolutionary processes of phenotypic evolution implied by microevolutionary predictions? \cite{hansen2016,jablonski2022evolvability}
The original descriptions of microevolutionary lines of least resistance explicitly refer to genetic constraints.
However, at the macroevolutionary scale the major axis of phenotypic variation is an emergent property of broader set of genetic and developmental constraints interacting with selection on a moving adaptive landscape % GT: Masahito suggested a whole paragraph on adaptive landscapes and Simpon-Lande-Arnold ideas. I'm not sure that will help but hopefully the expanded intro overall works better. Possibly just needs another sentence or two here but it would break the flow. more citations of relevant papers here (the first 7 or so papers from Gustavo's paper would be relevant.
\cite{BurinWhales,jones2004evolution}. % GT: More references needed? I'm not sure where the idea originates.
At this scale the major axis of phenotypic variation among species \cite{marroig2005size, fasanelli2022allometry} can be estimated from a matrix of divergences in species mean phenotypic traits (e.g. the \textbf{D} matrix of \cite{BlowsHiggie2003, mcglothlin2018adaptive}.
These matrices do not take into account phylogenetic relationships among species.
In contrast, the \textbf{R} matrix of \cite{Houle2017} describes the rate of among-species divergence in species traits and explicitly incorporates phylogeny (also see the \textbf{B} matrix of \cite{Machado2020}) %@L_Rmatrix.
We hereafter refer to this \textbf{R} matrix as the \textit{evolutionary rate matrix}. %TG: TODO: change throughout (that's the VCV matrix now)
The major axis of the evolutionary rate matrix is a phylogenetic line of least resistance applicable at macro- and mega-evolutionary scales that is analogous to the genetic line of least resistance at microevolutionary scales.

By studying the phylogenetic line of least resistance and related properties of the evolutionary rate matrix \cite{caetano2017ratematrix, caetano2018estimating, slater2019hierarchy} at different scales we aim to gain new insights into the macro- and mega-evolutionary trajectories that lead to the diversity of beak shapes among extant bird species and clades.
While there is no strict definition that separates macro- and mega-evolution, we define the macroevolutionary scale as species divergence within clades and the megaevolutionary scale as divergence of clades within larger clades.
The diversity of bird beaks has been well-studied at broad scales, particularly in relation to dietary ecology \cite{olsen2017feeding, felice2019, bright2016shapes, navalon2019, pigot2020macroevolutionary}.
Previous studies have assessed how variation in rates of evolution of beak shape has generated this diversity \cite{cooney2017mega,chira2018correlates}, revealing a general pattern of phylogenetically clustered evolutionary rates interspersed with major shifts deep in avian evolutionary history and in isolated lineages.
Major shifts indicate lineages that occupy distinct areas of beak shape space relative to their close relatives.
For example, a major shift in beak shape arises in the early divergence of Strisores separating hummingbirds (Trochilidae) from swifts (Apodidae), perhaps marking changes in ontogenetic trajectories \cite{navalon2021} as hummingbirds diverged from broad billed ancestors (e.g. the Eocene stem Trochilidae fossils \textit{Parargornis messelensis} and \textit{Eocypselus rowei} \cite{mayr2003new, ksepka2013fossil}).
By assessing the multivariate directions of the evolution of beak shapes of species and clades we aim to deepen our understanding of how diversity in morphological form arises at macro- and mega-evolutionary scales in birds.
To do this we use data on bird beaks derived from 3D scans of museum specimens for $>$8700 species ($\sim$85\% of all extant bird species) and generate evolutionary rate matrices describing the directions of multivariate beak shape evolution.
The phylogenetic breadth of the data allows us to compare directions of beak evolution of subclades (e.g. orders) within the class Aves (the megaevolutionary scale) and of species within subclades (the macroevolutionary scale).%@L_macromega_definition2

Various approaches have been developed to measure properties of trait matrices \cite{hansen2008measuring, pavlicev2009measuring, watanabe2022}, broadly focused on describing the size, shape, and orientation of trait divergence \cite{phillips1999}.
Our specific focus is on divergence of species and clades along or away from the phylogenetic line of least resistance.
We adopt concepts developed in the specific context of signal evolution to describe these alternative routes to divergence \cite{endler2005animal}.
Specifically, Endler \textit{et al}. (2005) refer to evolution of new phenotypes along the same major axis of phenotypic variation as \textit{elaboration}.
This is akin to phenotypic divergence among species following lines of least resistance, and, by definition, is expected to be a major driver of phenotypic evolution.
In contrast, Endler \textit{et al}. (2005) refer to evolution away from the major axis of phenotypic variation, either in random directions or along limited axes, as \textit{innovation} \cite{endler2005animal}.
Relative to elaboration, innovation is more likely to generate phenotypic novelty, but, if directions of phenotypic evolution are constrained, would not be expected to be a primary driver of phenotypic diversity among closely related species.
We apply these terms to i) species divergence within clades, and ii) clade divergence within larger clades (Figure 1).
Both innovation and elaboration could in principle occur at the macro- or mega-evolutionary scale.
We expect that elaboration is more important at macroevolutionary scales (species diverging within clades) and innovation more important at megaevolutionary scale (divergence among clades).
For our bird beak data, macroevolutionary scales refer to the beak shapes of species elaborating or innovating relative to the phylogenetic line of least resistance of their taxonomic order, and megaevolutionary scales refer to avian orders elaborating or innovating relative to the phylogenetic line of least resistance for the entire class Aves.
For clarity throughout the manuscript, we refer to elaboration and innovation as the general concept described above but use the following terms to distinguish macroevolutionary scale (species) and megaevolutionary scale (clade) elaboration and innovation: $elaboration_{species}$, $innovation_{species}$, $elaboration_{clade}$ and $innovation_{clade}$.
We describe the algebraic formulation of elaboration and innovation for species and clades in detail in the Methods section.
Using these novel statistical tools, we estimate the relative contribution of elaboration and innovation to the remarkable global radiation and diversity of bird beaks.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/cheat_sheet.pdf}
\caption{.}
\label{Fig:cheatsheet}
\end{figure}
\bigskip

\noindent \textbf{Figure \ref{Fig:cheatsheet}}
Figure 1:
Relationships between key figures in the main text.
A) represents the two first principal components (PCs) of the eight dimensional trait space with all bird beak shapes represented as gray circles and the Tinamiformes as red circles.
The gray ellipse represents the overall phylogenetic major axis of beak variation in birds.
Species aligned or away from this axis are respectively elaborators or innovators relative to the class Aves.
The red ellipse and axis represents the phylogenetic major axis of beak variation in Tinamiformes only.
Species aligned or away from this axis are respectively elaborators and innovators relative to the Tinamiformes order.
B) In Figure 2, we measured elaboration and innovation at the macroevolutionary scale , i.e. the position of each species in terms of $elaboration_{species}$ and $innovation_{species}$ relative to their order's phylogenetic major axis of beak variation.
A species $elaboration_{species}$ score is the species projection onto their order's phylogenetic major axis of beak variation(i. e. their position on that axis) whereas their $innovation_{species}$ score is their rejection from it (i.e. their distance away from that axis).
The colours on the tips of the phylogeny correspond to the median $elaboration_{species}$ (orange gradient) and $innovation_{species}$ (blue gradient).
C) Figure 4, we measured elaboration and innovation at the megaevolutionary scale, i.e. $elaboration_{clade}$ and $innovation_{clade}$.
The ellipses are scaled and centered and represent the clade's phylogenetic major axis of beak variation relative to the overall phylogenetic major axis of beak variation with the relative length of the ellipse on each dimension represented on the barplots as variance on each dimension.
A clade's $elaboration_{clade}$ and $innovation_{clade}$ scores are measured by projecting the clade's phylogenetic major axis of beak variation onto the overall phylogenetic major axis of beak variation and measuring the projection (elaboration) and rejection (innovation) from this axis as described above.
The distribution elaboration and innovation scores are taken from the projection of the 4000 pairs of evolutionary rate matrices with each pair being the focal level, e.g. order,and the parent level, e.g. the whole bird phylogeny.

% \begin{figure}[!htbp]
% \centering
%    \includegraphics[width=1\textwidth]{Figures/ellipses.pdf}
% \caption{.}
% \label{Fig:ellipses}
% \end{figure}
% \bigskip

% \noindent \textbf{Figure \ref{Fig:ellipses}:}
% Panel A) shows  ellipses representing the scaled median posterior variance-covariance response from the pGLMM models for each superorder (coloured ellipses) compared to the all-birds phylogenetic component of the models (grey ellipses)
% We scaled the ellipses so the length of the major axis of the clade ellipses is the same length as that of the global ellipse (in eight dimensions).
% The first inset ellipse plot shows the positions of all superorder ellipses relative to the all-bird phylogenetic ellipse.
% Subsequent inset plots show the results for each superorder. Inset barplots show the proportion of variance associated with each of the eight PC axes in shape morphospace.
% The inset boxplots correspond to the elaboration (E) and innovation (I) scores for all 4000 posterior samples.
% The dots represent the median elaboration and innovation values while the thick and dashed lines represent the 50\% and 95\% confidence intervals respectively.
% Note that these scores were calculated on the unscaled ellipses resulting in different scales of elaboration and innovation scores for each plot. Panel B) shows  the results for each order.
% A companion plot showing further nested structure within the Passeriformes is shown in Fig. S\ref{fig_ellipses_passeriformes}.

% \bigskip

% \begin{figure}[!htbp]
% \centering
%    \includegraphics[width=0.9\textwidth]{Figures/orthogonality_results.pdf}
% \caption{.}
% \label{Fig:orthogonality}
% \end{figure}

% \bigskip

% \noindent \textbf{Figure \ref{Fig:orthogonality}:}
% Orthogonality of each clade's major axis compared to their parent or parent parent's clade.
% Orthogonality is represented on the horizontal axis scales from 0 (modulo of $0^\circ$) to 1 (modulo of $90^\circ$) with the background grey and dashed grey lines represent respectively an orthogonality of 0.5 (modulo of $45^\circ$) and 0.75 (modulo of $67.5^\circ$).
% Dots represent the clades median orthogonality and the lines their 95\% CI.
% No super order's major axis is parallel to the whole phylogeny (red lines and circles) and 12/25 orders are at least half orthogonal to their super-order (green lines and circles).
% These results are even clearer when comparing them to the whole phylogeny (green dashed lines and open circles) where only 3/25 are not at least half orthogonal.
% Furtheremore we also indicate here the number of species (n) and the standard deviation of their major axes orientation over the 4000 variance-covariance posteriors (sd; expressed in degrees).
% For each clade we also measured the posterior propbabilityof each clade's orientation being different from their parent's clade or the whole phylogeny relative to their sample size and sd.
% The stars represent the posterior probability of the clade's orientation being different from the comparison clade. (*** = pp $> 0.99$; ** = pp $>0.95$; * = pp $> 0.9$; . = pp $> 0.8$).
% Only two out of six super-orders; 6/25 orders relative to their super order and 9/25 orders relative to the phylogeny have a pp $> 0.99$.
% This is the result of a variation of sample size (n) and standard variation (sd) between clades. 
% A companion plot showing further nested structure within the Passeriformes is shown in Fig. S\ref{fig_orthogonality_passeriformes}.

% \bigskip

% However, this change in the line of least resistance's orientation between different clade levels varies between clades (Fig.\ref{Fig:ellipses}).
% While orthogonality provides a simple description of the degree of reorientation of phenotypic axes (measuring the two-dimensional plane of an angle), it does not fully capture the multi-dimensionality of the data.
% To further quantify the tension implied between theory on lines of least evolutionary resistance and observations of deep-time reorientation of trait covariance, we measured the elaboration and innovation of each different clades (\textit{cf.} the elaboration and innovation of a species - see below).
% As noted above, elaboration describes multivariate evolution that is primarily along the phylogenetically constrained phenotypic major axis whereas innovation describes multivariate evolution that is primarily away from the phylogenetically constrained phenotypic major axis \cite{endler2005animal}.
% Clades (or species) can be both elaborators and innovators.
% To measure innovation and elaboration we translated the major axes of each clade's (i.e. superorders and orders) phylogenetic variance-covariance matrices onto the global phylogenetic major axis so that they shared the same origin in the shapespace.
% We then used multilinear algebra to project a focal clade's major axis onto the phylogenetic major axis (Fig.S\ref{Fig:cheat_sheet}).
% We interpret the measured projection (distance onto the phylogenetic major axis) as the clade's elaboration score and the measured rejection (distance from the major axis) as the clade's innovation score.
% These scores were measured for each of 4000 posterior pairs of clade vs. phylogenetic variance-covariance matrix (Fig.\ref{Fig:ellipses} boxplots).
% We found that for most clades the phenotypic major axis is not aligned with the global phylogenetic major axis, (Fig. \ref{Fig:ellipses}).
% Remarkably, more than half of the assessed clades (4/8 superorders and 15/27 orders) displayed higher median innovation than median elaboration scores implying that, contrary to theory \cite{schluter1996adaptive,marroig2005size}, innovation is a more common generator of morphological diversity than elaboration.
% However, there is substantial variation in terms of elaboration and innovation across clades including relatively low elaboration and innovation (e.g. Cuculiformes; Fig. \ref{Fig:ellipses}), low elaboration and high innovation (e.g. Galliformes; Fig. \ref{Fig:ellipses}), high elaboration and low innovation (e.g. Coraciiformes; Fig. \ref{Fig:ellipses}) or both relatively high elaboration and innovation (e.g. Podicipediformes; Fig. \ref{Fig:ellipses}).
% These patterns hold across scales including within superorders and at finer scales within the order Passeriformes (Fig. S\ref{fig_ellipses_passeriformes}).
% This suggests that divergence away from line of least evolutionary resistance (i.e. the phylogenetic major axis) is a common feature of avian beak evolution across scales and further supports the suggestion of a remarkable flexibility in the direction of evolution in deep-time even in a developmentally constrained trait.

% This apparent lack of deep-time evolutionary constraint in bird beaks is supported by variation in both the dominant direction of evolution within clades and their relative dimensionality.
% Although the major axis of the global phylogenetic component aligns closely with the first dimension (70.96\% of the phylogenetic major axis is aligned on PC1; Fig. \ref{Fig:ellipses}) of the raw shapespace, no superorder, and only seven of the 27 orders (Procellariiformes, Pelecaniformes, Charadriiformes, Coraciiformes, Psittaciformes and Passeriformes), mainly align with PC1 in shapespace Fig. \ref{Fig:ellipses}-bar plots).
% These patterns are similar (albeit more contrasted) when looking at sub-orders and families within passeriformes half off the sub-orders (Meliphagoidea, Corvides and Passerida) and seven out of the 23 families (Eurylaimides, Meliphagoidea, Petroicidea, Fringilidae, Aegithaloidea, Pyconotiae, Nectariniidae) align with PC1 (Fig. S\ref{fig_ellipses_passeriformes}).
% Moreover, we found that within clades, phenotypic variation is either highly constrained (varying almost entirely along a single axis) or higher-dimensional than expected from the global phylogenetic partition.
% For example, Accipitriformes (hawks and allies) have major axes that mainly align with the second dimension of the shapespace suggesting uniformity in their beak shape components, whereas the Podicipediformes (grebes) are highly variable across all dimensions suggesting that all components of the shapespace are necessary to describe their beaks.
% These observations illustrate multiple routes to innovation and imply that such innovation could arise at any time throughout avian history.
% However, although we see repeated innovation at the clade level, providing a new hypothesis to explain the expansion of avian morphological space, these analyses do not allow us to distinguish whether reorientation arises through abrupt changes or more gradual realignment of species in trait space.

% Abrupt reorientation of trait space in higher taxa would be consistent with Simpsonian megaevolutionary dynamics \cite{simpson1953major}.
% This would lead to an expectation that species should elaborate relative to their parent clade (e.g. within orders) but show a stronger tendency towards innovation relative to the global phylogeny.
% To assess the relative contributions of elaboration and innovation in beak morphology among species we estimated the algebraic projection and rejection of their position in the shapespace onto (i) the global phylogenetic major phenotypic axis, (ii) the major phenotypic axis of their respective parent superorder, and (iii) the major phenotypic axis of their respective parent order.
% We did this by using the same multilinear algebraic method as described above but measuring elaboration and innovation for individual species (across the 4000 posterior samples) rather than for entire clades.
% We found that high levels of elaboration are more common across all birds than high levels of innovation (Fig. \ref{Fig:phylogeny}).
% However, we also found various combinations of elaboration and innovation across birds.
% For example, compared to the phylogenetic major axis, the Trochilidae (hummingbirds, with long narrow beaks) display high levels of elaboration and low innovation whereas the Bucerotidae (hornbills, with large and oddly shaped beaks) display high levels of innovation and low elaboration.
% Other clades such as the Apodidae (swifts, with large but gaping beaks) and the distantly related but convergent Hirundidae (swallows and martins) display high levels of both elaboration and innovation.
% Overall, we find no consistent evidence that innovation is either positively correlated with, or trades-off with, elaboration. %(Fig. S{#fig_correlations}). %TODO: maybe add to supp?
% The variability of, and apparent disconnect between, innovation and elaboration holds whether comparisons are made at the order, superorder, or class-wide phylogenetic level (Fig. \ref{Fig:phylogeny})%{#fig_correlations}),%TODO: maybe add to supp?
%  and in further nested analysis of the order Passeriformes (Fig. S\ref{fig_ellipses_passeriformes} and S\ref{fig_phylogeny_passeriformes}).
% The dominance of elaboration when measured at the species level is consistent with ideas of evolution along lines of least evolutionary resistance, yet contrasts with our observations of frequent innovation via reorientation of trait space within clades.

% \begin{figure}[!htbp]
% \centering
%     \includegraphics[width=0.9\textwidth]{Figures/InnovElabTree_main_text.pdf}
% \caption{.}
% \label{Fig:phylogeny}
% \end{figure}

% \bigskip

% \noindent \textbf{Figure \ref{Fig:phylogeny}:} The avian phylogeny (n=8748 species) coloured by species beak shape innovation and elaboration level.
% Blue and yellow colors highlight species with relatively low elaboration or low innovation, whereas orange to red colours indicate high innovation or elaboration.
% Innovations and elaboration are displayed as: a) comparisons of species to the phylogenetic major axis (inner two bands); b) comparisons of species to their superorder major axis (middle two bands); and c) comparisons of species to their order major axis (outer two bands).
% The branches are coloured by the species distance from the centroid of the shapespace and the estimation of their ancestors' position in that space.
% A companion plot showing further nested structure within the Passeriformes is shown in Fig. S\ref{fig_phylogeny_passeriformes}.

% \bigskip

% The disconnect between observations of reorientation of trait space at the clade level and a dominance of elaboration at the species level could be viewed as a paradox.
% We formalise this paradox by testing whether elaboration is more common at higher taxonomic levels than at the species level.
% We compared the area under the curve of the scaled density of elaboration and innovation at both the species level and the clade level.
% We then measured the difference between both areas and found that there was a significant difference in innovation between species and clades but no clear difference in elaboration (Fig. \ref{Fig:relative_EI}).
% This difference confirms our observation that innovation is more common at the clade level rather than at the species level and that although species evolve preferentially along a line of least evolutionary resistance, this line of least resistance changes frequently among clades.
% However, we suggest that rather than a paradox the reorientation of trait space can instead be described as a gradual process occuring in response to shifting selection pressures acting on species evolving in similar adaptive zones.
% This implies that gradual directional evolution \cite{pagel2022general} rather than megaevolutionary jumps \cite{cooney2017mega,venditti2011multiple} may be sufficient to explain diversity in avian beak morphology.

% \begin{figure}[!htbp]
% \centering
%    \includegraphics[width=0.9\textwidth]{Figures/relative_EI.pdf}
% \caption{.}
% \label{Fig:relative_EI}
% \end{figure}

% \bigskip

% \noindent \textbf{Figure \ref{Fig:relative_EI}:} Comparisons of relative elaboration and innovation among clades (illustrated in figure  \ref{Fig:ellipses}) and among species (figure \ref{Fig:phylogeny}).
% A - In blue, the elaboration for each species in each 4000 posterior samples compared to the elaboration for each 35 clades.
% The scores for each clade is scaled by the maximum score within each clade (corresponding to the scaled ellipses in figure  \ref{Fig:ellipses}) and the score for the species is scaled by the maximum species elaboration score.
% This allows us to compare both types (species vs. clades) despite their different sample sizes.
% The Bhattacharryya Coefficient represents the overlap of the area under the curve for species and clades (a BC of 0 represents no overlap).
% B - same as A but for the innovation score.
% We can see a clear difference in the amount of innovation across the posterior samples between species or clades.

% \bigskip

% Our results show that although at the species level most bird beaks are elaborating along a megaevolutionary line of least resistance, at the macroevolutionary level innovation way from the megaevolutionary line of least resistance is much more common.
% This nested structure of elaboration at a higher taxonomic level and innovation at a lower taxonomic level could thus explain the diversity of bird beaks we observe today.
% However, individual species-level variation in the past could also have led to a shift in a clade's line of least evolutionary resistance through different evolutionary routes on a dynamic adaptive landscape. Taken together, our results suggest that the signature of evolutionary orientations in deep-time, coupled with elaboration at the species-level, is consistent with recent suggestions that the apparently abrupt phenotypic shifts can be explained by normal Darwinian processes \cite{pagel2022general}.
% Our results help to bridge the gap between micro- and macro-evolution by proposing a multi-scale framework for  phenotypic evolution that allows species to evolve along lines of least evolutionary resistance while accommodating frequent shifts in the direction of evolutionary change.

% \section{Methods}

% \textbf{Beak shape data.}
% We used the dataset from \cite{cooney2017mega,chira2020signature,hughes2022global} which consists of four landmarks and three curves of 25 semi-landmarks each, taken from 3D scans of  the beaks of 8748 bird species.
% The landmark data were aligned with Procrustes superimposition using the R package \texttt{Morpho} \cite{Rcore,Morpho} and we removed the size component and controlled for symmetry (see \cite{cooney2017mega,chira2020signature,hughes2022global} for detailed descriptions of data collection and processing).
% We ordinated the superimposed landmarks using a principal components analysis (PCA) and selected the eight first dimensions of the resulting PCA.
% These eight dimensions contained at least 95\% of variance in each superorder, and 98.7\% of the variance in the whole dataset.
% We used a threshold of $>95$\% of variance explained for each superorder because although the overall distribution of the variance on each PC axis in the whole trait-space is decreasing (i.e.
% $\sigma$ PC1 $<$ $\sigma$ PC2 $<$ $\sigma$ PC3), this is not always true for each superorder.
% For example, in the Mirandornithes, PC2, then PC4, and then PC1 contained the most variance (Fig. \ref{Fig:ellipses} and S\ref{Fig:axes_variance}).
% This protocol for selecting the number of PC axes ensures that the resulting trait-space contains enough variance for each superorder (see Fig. S\ref{Fig:axes_variance} and Table \ref{tab_variance_per_axis}).
% This procedure resulted in an $8748 \times 8$ matrix that contains at least 95\% of the variance in all observed living bird beaks (hereafter, the shapespace).

% \textbf{Phylogenetic data.}
% We used a random subsample of 1000 trees from the avian tree distribution of \cite{jetz2012global} for the analyses and the one randomly selected tree for the figures (Fig. \ref{Fig:phylogeny}).
% We pruned these trees to contain only the 8748 species present in our beak shapespace.

% \textbf{Phylogenetic GLMM analysis.}
% We ran a phylogenetic Bayesian generalised linear mixed effects model (pGLMM) using the \texttt{MCMCglmm} R package \cite{MCMCglmm} with the shapespace data as a multidimensional response variable, the error in the dataset as residuals, and each clade's phylogeny as a nested random term, i.e., a model of:

% \begin{equation}
% \text{shapespace} \mathtt{\sim} \text{residuals(shapespace)} + \text{random(phylogeny)}
% \end{equation}

% This model allows us to estimate a phenetic variance-covariance matrix for each clade and the phylogeny as a whole, and one for the residuals in the shapespace itself \cite{robinson2013quantifying}.
% These variance-covariance matrices were then used in a base projection-rejection analysis to measure the elaboration and innovation scores (see below).
% We ran two separate nested models.
% One model using one random term for the whole phylogeny and 35 nested random terms for each superorder and order containing more than 15 species in our dataset (Fig. \ref{Fig:ellipses}).
% And one model on a subset of the dataset containing only the 5229 passeriformes species with 30 nested random terms.
% This resulted in two multidimensional GLMM.
% One with a $8748 \times 8$ response variables, one global residual term and 36 random terms.
% And one with a $5229 \times 8$ response variable, one global residual term and 30 random terms.

% To account for phylogenetic uncertainty we needed to run the model using different tree topologies from the tree distribution in \cite{jetz2012global}.
% Because of the very large size of our model and dataset, however, running the full model on multiple trees was computationally unfeasible (it would require 2 CPU years and 4TB of RAM for each - the whole tree and the passeriformes ones).
% Instead, we developed and implemented a ''mini chains'' MCMCglmm method using the R package \texttt{mcmcmglmmm} \cite{mcmcmcglmmm}.

% \textbf{mcmcmcglmm.}
% This method consists in running relatively short monte-carlo markov chains (MCMC) across a large number of different phylogenetic trees and concatenating the resulting short posteriors into one larger one that contains the phylogenetic uncertainty information.
% This was effectively done in four steps (see Fig. S\ref{Fig:mcmcmcglmm})
% 1) running three independent MCMCglmm chains (hereafter the parametrisation chains) for 50k generations with the traits data and the model described above on the consensus tree from the tree distribution, along with flat priors with a belief parameter of 2\% (effectively not a very low weight on the priors);
% 2) extracting the following parameters from the parameterisation chains:
% a) the conservative burnin average time (across the three chains) defined as the highest number of iterations required to first reach the median posterior likelihood value multiplied by 1.1; and
% b) the mini-chains priors defined as the median posterior model values from the parameterisation chains with a belief parameter of 5\%.
% We then ran 400 mini chains in parallel with the estimated burnin time and priors in order to just run 10 samples past the burnin.
% This effectively resulted in getting 10 exploitable posterior samples per tree.
% We used 400 trees because that was the number of models required to reach an effective sample size (ESS) of at least 200 for all the estimated parameters (see Fig. S\ref{Fig:model_ess_all_birds} and S\ref{Fig:model_ess_passeriformes}).
% Using this approach, we reduced the computational time to around 45 CPU hours and 9GB of RAM per mini-chain (a 400 fold improvement).
% The whole procedure is reproducible in the \href{https://raw.rawgit.net/TGuillerme/mcmcmcglmmm/main/inst/MCMCglmm_mini_chains.html}{\texttt{mcmcmcglmmm} vignette} \cite{mcmcmcglmmm}.

% \textbf{Projection/rejection analyses.}
% We used the distributions of the 4000 posterior variance-covariance matrices to run the projection/rejection analyses to measure the bird beaks' elaboration and innovation scores.
% In fact we can use multilinear algebra to interpret elaboration and innovation \textit{sensu} \cite{endler2005animal} by using the major axis of the variance-covariance matrices as the ''line of least resistance'', or, specifically here, the line of elaboration for each random terms (corresponding to the elaboration axes for each clade) and project either species or other major axis onto that clade.
% In brief we can interpret where a species or a clade falls on the major axis of reference as its elaboration score and how far a species is from that major axis as its innovation score.
% We calculated the elaboration and innovation scores using two main approaches:
% 1) by clades (i.e. to measure the elaboration/innovation between clades) where we calculated the projections of the major axes of each clade onto the global phylogenetic major axes (Fig. \ref{Fig:ellipses});
% and 2) by species (i.e. to measure the elaboration/innovation within clades) where we calculated the projections of each species onto a) the global phylogenetic major axes and b) the phylogenetic major axes of their respective clade (Fig. \ref{Fig:phylogeny}).
% To make the results easier to interpret, we centered and scaled them onto the centre of the major axis of interest and made the elaboration values absolute.
% This way, we can interpret an elaboration or innovation score within the 0-1 range to be not exceptional (i.e within the 95\% confidence interval of the variance-covariance matrix).
% The entire mathematical procedure is described in details in the supplementary materials (section \ref{supp_projection}).
% The whole procedure applied to our dataset is reproducible in \href{https://raw.rawgit.net/TGuillerme/dispRity/master/inst/vignettes/Projection_analysis.html}{this \texttt{dispRity} vignette} % TODO: update link to release when releasing
% and implemented in the \texttt{dispRity} package \cite{dispRity}.

% \textbf{Groups' orthogonality.}
% We measured the orthogonality of each random term in our models (clades) compared to their parent clade and their parent parent's.
% For example, the Columbimorphae random terms' variance-covariance matrices against the phylogeny ones, the Columbiformes ones against the Columbimorphae ones (parent) and the whole phylogeny ones (parent parents).
% To do so we measured the angles between the major axes of the focal clade and their parent clade for each posterior variance-covariance matrix as well as within each clade by randomly comparing major axes within a clade.
% We converted each angle measurement into an orthogonality metric where 0 corresponds to flat angles ($0^\circ$ or  $180^\circ$) and 1 to right angles ($90^\circ$ or $270^\circ$).
% We then measured the posterior probability of the angles within a clade being different than the angles between clades.
% These results are available in Fig. \ref{Fig:orthogonality}.

% \textbf{Reproducibility and repeatability.}
% The raw and processed data used in this analysis is available from \href{https://figshare.com/articles/dataset/Innovation_and_elaboration_on_the_avian_tree_of_life/20480355}{figshare} \cite{fighsaredata}.
% The code used to reproduce the figures and tables is available from \href{https://github.com/TGuillerme/elaboration_exploration_bird_beaks}{github} \cite{githubrepo}.

\textbf{Funding.}
This work was funded by UKRI-NERC Grant NE/T000139/1 and a Royal Society University Research Fellowship (URF R 180006 to GHT).

\bibliographystyle{naturemag}
\bibliography{references}

\end{document}