### 2023-06-19 Analysis/07-Orthogonality_and_statistics.Rmd
reinstall

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme and Gavin Thomas"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This section contains the code for reproducing the analyses. To get the details, refer to [`dispRity` vignette](link_to_projection_vignette).
This script take around 35 minutes to compile.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
```

# Measuring ellipse's statistics

For each group, we can calculate:

 * It's (dis)alignment relative to the the phylogeny
 * It's distance from the centroid of the phylogeny
 * It's angles sd
 * It's orientation (angle) relative to the whole phylogeny
 * Whether that orientation is significant

# Loading the data

```{r, eval = TRUE, echo = FALSE}
## Load the data for all birds
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace_allbirds <- shapespace_allbirds_lvl_superorder_order[[1]]$space
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_posteriors <- shapespace_allbirds_lvl_superorder_order[[1]]
rm(shapespace_allbirds_lvl_superorder_order)

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_posteriors) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_posteriors))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_posteriors))

## Create the subset for all birds
all_birds_shapespace <- MCMCglmm.subsets(
      data          = shapespace_allbirds,
      posteriors    = shapespace_allbirds_posteriors,
      group         = MCMCglmm.levels(shapespace_allbirds_posteriors)[c(super_lvl, order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace_allbirds$level1)[-1], levels(shapespace_allbirds$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_posteriors)

## Load the covar
load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
shapespace_passeriformes_posteriors <- shapespace_passeriformes[[1]]
rm(shapespace_passeriformes)
## Load the data for the passeriformes
load(file = "../Data/Processed/shapespace_passeriformes_two_levels.rda")
## Select the data
shapespace_passeriformes <- shapespace_passeriformes_two_levels[[1]]$space
## Clean garbage
rm(shapespace_passeriformes_two_levels)

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_passeriformes_posteriors) == "animal")
suborder_lvl <- grep("level1_", MCMCglmm.levels(shapespace_passeriformes_posteriors))
family_lvl <- grep("level2_", MCMCglmm.levels(shapespace_passeriformes_posteriors))

## Create the subset for all birds
passeriformes_shapespace <- MCMCglmm.subsets(
      data          = shapespace_passeriformes,
      posteriors    = shapespace_passeriformes_posteriors,
      group         = MCMCglmm.levels(shapespace_passeriformes_posteriors)[c(suborder_lvl, family_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace_passeriformes$level1)[-1], levels(shapespace_passeriformes$level2)[-1], "Passeriformes")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_passeriformes_posteriors)
```

```{r, eval = TRUE, echo = FALSE}
## Creating the nested levels subgroups
make.nested.subset <- function(group, data, group.data) {
    ## Get the subsets within the group
    level2_groups <- as.character(unique(group.data$level2[group.data$level1 == group]))
    if(any(empty <- level2_groups == "")) {
        level2_groups <- level2_groups[-which(empty)]
    }
    if(length(level2_groups) < 1) {
        return(NULL)
    }
    ## Create a subset
    return(get.subsets(data, subsets = c(level2_groups, group)))
}

## Get the subgroups for all_birds
all_birds_shapespace_subgroups <- lapply(as.list(levels(shapespace_allbirds$level1)[-1]), make.nested.subset, data = all_birds_shapespace, group.data = shapespace_allbirds)
names(all_birds_shapespace_subgroups) <- levels(shapespace_allbirds$level1)[-1]
## Get the subgroups for the passeriformes
passeriformes_shapespace_subgroups <- lapply(as.list(levels(shapespace_passeriformes$level1)[-1]), make.nested.subset, data = passeriformes_shapespace, group.data = shapespace_passeriformes)
names(passeriformes_shapespace_subgroups) <- levels(shapespace_passeriformes$level1)[-1]
```

### Groups compared to the phylogeny

```{r, eval = FALSE, echo = FALSE}
source("../Functions/stats.utilities.R")

## General measurement
all_birds_all_levels <- ellipse.stats(all_birds_shapespace)
passeriformes_all_levels <- ellipse.stats(passeriformes_shapespace)

## Nested measurements
all_birds_nested_levels <- lapply(all_birds_shapespace_subgroups, ellipse.stats)
passeriformes_nested_levels <- lapply(passeriformes_shapespace_subgroups, ellipse.stats)

## Big measurements list
ellipses_stats <- list("all_birds" = all_birds_all_levels)
ellipses_stats <- c(ellipses_stats, all_birds_nested_levels)
ellipses_stats$Passeriformes <- passeriformes_all_levels
ellipses_stats <- c(ellipses_stats, passeriformes_nested_levels)

save(ellipses_stats, file = "../Data/Processed/ellipses_stats.rda")
```

And here are the full results! Here's what's in each column:

 * **Group**: the name of the group (a random term of the model), the focal VCV
 * **Comparison**: to which group are the VCV compared (the group is either one or two levels higher)
 * **n**: number of species
 * **distance**: the distance between the focal and the comparison group's centers
 * **ellipse sd**: the standard deviation of the angles in the the focal VCV distribution
 * **disalignment**: the disalignment of the group compared to the focal group (0 = aligned), with the 95% CI
 * **orthogonality**: the orthogonality amount between both groups (0 = parallel, 1 = orthogonal), with the 95% CI
 * **Post. prob.**: the posterior probability of the orthogonality in the focal group being different from the comparison one (with the 95% CI estimated from the randomised posterior probabilities)

```{r, eval = TRUE, echo = FALSE}
load("../Data/Processed/ellipses_stats.rda")
source("../Functions/stats.utilities.R")
library(xtable)

## Sorting all results
all_results <- do.call(rbind, lapply(ellipses_stats[-12], table.stats)) ## Droping the Meliphagoidea (single group)
rownames(all_results) <- NULL

## csv results
write.csv(all_results, "../Data/Processed/ellipses.stats.csv")

## md results
knitr::kable(all_results)

## tex results
to_export <- all_results[, -c(4, 6:8)]
colnames(to_export)[c(6,9)] <- "2.5\\%"
colnames(to_export)[c(7,10)] <- "97.5\\%"
make.xtable(to_export, digit = 3, caption = "Posterior variance-covariance ellipses results for each clade compared to their parent clade or their parent parents clade (Comparison). n = the number of species per group. sd = the standard deviation of the ellipses orientation (across the posterior distribution). orthogonality = the degree of right angle for each group compared to their parent or parent's group (0 = parellel, 1 = orthogonal). Post. prob = the posterior probability of the orthogonality in the focal group being different from the comparison one (the 95\\% CI is from the randomised posterior probabilities).", label = "tab_ortho_results", longtable = FALSE, path = "../Manuscript/Tables/", include.rownames = FALSE)
```

Median orthogonality for all super-orders and orders to the phylogeny (paragraph 4)

```{r}
get.angles <- function(X) {
    unlist(get.disparity(X$angles, concatenate = FALSE))
}
## Median angle
median(get.angles(ellipses_stats$all_birds)) * 90
## Median angle CI
quantile(get.angles(ellipses_stats$all_birds), probs = c(0.025, 0.975)) * 90
```

Median orthogonality for all super-orders and orders to the phylogeny (paragraph 4)

```{r}
## Median angle
median(unlist(lapply(ellipses_stats[2:9], get.angles))) * 90
## Median angle CI
quantile(unlist(lapply(ellipses_stats[2:9], get.angles)), probs = c(0.025, 0.975)) * 90
```

<!-- length(which(all_results[which(all_results[,"Post. prob."] >= 0.95), ]$orthogonality > 0.5)) -->

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
source("../Functions/ploting.utilities.R")
## Plot the orthogonality into context
orthogonality <- all_results[, c(1:2, 9:14, 3, 5)]

## Group per level
super_orders <- orthogonality[c(1:8), ]
orders <- orthogonality[c(36:60), ]
orders_n2 <- orthogonality[c(9:35), ]
sub_orders <- orthogonality[c(61:66), ]
families <- orthogonality[c(90:110), ]
families_n2 <- orthogonality[c(67:89), ]

## Adding orders and families that don't have a higher level covered
if(any(no_match <- is.na(match(orders_n2$Group, orders$Group)))) {
    fill_in <- orders_n2[which(no_match), ]
    fill_in[,c(3:8)] <- NA
    orders <- rbind(orders, fill_in)
}
## Adding orders and families that don't have a higher level covered
if(any(no_match <- is.na(match(families_n2$Group, families$Group)))) {
    fill_in <- families_n2[which(no_match), ]
    fill_in[,c(3:8)] <- NA
    families <- rbind(families, fill_in)
}

pdf("../Manuscript/Figures/orthogonality_results.pdf", width = 9, height = 9)
plot.orthogonality(super_orders, orders_n2, orders, col = c("purple", "#FBB117"))
legend("topleft", legend = c("Super-orders (vs. phylogeny)", "Orders (vs. phylogeny)", "Orders (vs. super-orders)"), col = c("purple", "#FBB117", adjustcolor("#FBB117", alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8, bty = "n")
dev.off()
```

Distribution of the VCV ellipses' orthogonality for all the groups compared to the ellipse of their parent clade (or n-2). 
For the posterior probabilities: . >= 0.8; * >= 0.9; ** >= 0.95; *** >= 0.99.

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
## Same but for the passeriformes
pdf("../Manuscript/Figures/orthogonality_results_passeriformes.pdf", width = 9, height = 9)
plot.orthogonality(sub_orders, families_n2, families, col =c("purple", "#FBB117"))
legend("topleft", legend = c("Sub-orders (vs. passeriformes)", "Families (vs. sub-orders)", "Families (vs. passeriformes)"), col = c("purple", "#FBB117", adjustcolor("#FBB117", alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8, bty = "n")
dev.off()
```

## Testing whether clades innovate more than species

```{r, echo = FALSE, eval = TRUE}
source("../Functions/stats.utilities.R")
library(dispRity)
library(ggplot2)
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
results <- shapespace_allbirds_lvl_superorder_order_results_list
rm(shapespace_allbirds_lvl_superorder_order_results_list)

## Distribution of elaboration and innovation at the species level
elaboration_species <- get.disparity(get.subsets(results$super_results_phylo$position, subsets = "phylogeny"), concatenate = FALSE)[[1]]
innovation_species  <- get.disparity(get.subsets(results$super_results_phylo$distance, subsets = "phylogeny"), concatenate = FALSE)[[1]]
elaboration_species <- elaboration_species/max(elaboration_species)
innovation_species  <- innovation_species/max(innovation_species)

## Distribution of elaboration and innovation at the clade level
scale.group <- function(X) return(X/quantile(X, prob = 0.975)[1])
elaboration_super_orders <- lapply(get.disparity(results$super_results_between$position, concatenate = FALSE), scale.group)
innovation_super_orders  <- lapply(get.disparity(results$super_results_between$distance, concatenate = FALSE), scale.group)
elaboration_orders       <- lapply(get.disparity(results$order_results_between$position, concatenate = FALSE), scale.group)
innovation_orders        <- lapply(get.disparity(results$order_results_between$distance, concatenate = FALSE), scale.group)

## RAM cleaning
rm(results)

## The objects above are the ones that contain all the elaboration and innovation values by clade level
## These bits are for the Bhatt coeff
# elaboration_group <- rbind(do.call(rbind, elaboration_super_orders), do.call(rbind, elaboration_orders))
# innovation_group <- rbind(do.call(rbind, innovation_super_orders), do.call(rbind, innovation_orders))

## Measuring the area under the curves (for the Bhatt coeff)
# elaboration_species_auc <- apply(elaboration_species, 2, area.density)
# innovation_species_auc  <- apply(innovation_species, 2, area.density)
# elaboration_group_auc   <- apply(elaboration_group, 2, area.density)
# innovation_group_auc    <- apply(innovation_group, 2, area.density)

# library(tidyverse)
library(patchwork)

fake <- data.frame(elaboration = c(elaboration_species, c(unlist(elaboration_super_orders), unlist(elaboration_orders))), innovation = c(innovation_species, c(unlist(innovation_super_orders), unlist(innovation_orders))), Level = c(rep("Species", 34992000), rep("Clades", 140000)))

elab <- 
  ggplot(fake, aes(x = elaboration, colour=Level, fill = Level, ..scaled..)) +
  geom_density(alpha = 0.5) +
  theme_classic(base_size = 14) +
  scale_colour_manual(values = c(rgb(0,0,0,0), rgb(0,0,0,0))) +
  scale_fill_manual(values = c("#FBB117", "purple")) +
  xlab("Relative elaboration") +
  ylab("Density") +
  theme(legend.position = "none") +
  labs(tags="a")

innov <-
  ggplot(fake, aes(x = innovation, colour = Level, fill = Level, ..scaled..)) +
  geom_density(alpha = 0.5) +
  theme_classic(base_size = 14) +
  scale_colour_manual(values = c(rgb(0,0,0,0), rgb(0,0,0,0))) +
  scale_fill_manual(values = c("#FBB117", "purple")) +
  xlab("Relative innovation") + 
  theme(legend.position = c(0.8, 0.8)) +
  labs(tags="b")

remove_y <- 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p <- elab + innov + remove_y

ggsave("../Manuscript/Figures/relative_EI_gg_style_95CIscaled.pdf", plot=p, height=6, width=12)
```

## Testing whether the passeriformes ellipses are similar

Basically whether the level 2 passeriformes random terms VCVs in the all birds models are similar to the level 0 passeriformes random terms VCVs in the passeriformes only model.
We don't expect them to be identical due to randomness of the selected trees and random seeds but they should be at least very similar.

```{r}
## Get the VCVs from each model
model1 <- all_birds_shapespace$covar$Passeriformes
model2 <- passeriformes_shapespace$covar$Passeriformes
allbirds <- all_birds_shapespace$covar$phylogeny
passerida <- passeriformes_shapespace$covar$Passerida

## Testing the differences between passeriformes in both spaces
passeriformes_model1 <- covar.ellipse.test(model1, model2, n = 4000, measure = "orthogonality")
passeriformes_model2 <- covar.ellipse.test(model2, model1, n = 4000, measure = "orthogonality")

## Testing the differences between passeriformes in both spaces
passeriformes_model1_vs_allbirds <- covar.ellipse.test(model1, allbirds, n = 4000, measure = "orthogonality")
passeriformes_model2_vs_allbirds <- covar.ellipse.test(model2, allbirds, n = 4000, measure = "orthogonality")
passeriformes_model1_vs_passerida <- covar.ellipse.test(model1, passerida, n = 4000, measure = "orthogonality")
passeriformes_model2_vs_passerida <- covar.ellipse.test(model2, passerida, n = 4000, measure = "orthogonality")

## Merging the results
passeriformes_comparisons <- rbind(passeriformes_model1,
                                   passeriformes_model1_vs_allbirds,
                                   passeriformes_model1_vs_passerida,
                                   passeriformes_model2,
                                   passeriformes_model2_vs_allbirds,
                                   passeriformes_model2_vs_passerida)

## Visualising the differences
pdf(file = "../Manuscript/Figures/Passeriformes_differences.pdf", width = 9, height = 5)
par(mar = c(5, 12, 2, 2))
plot(NULL, xlim = c(0.5, 1), ylim = c(6,1), xlab = "Posterior probability of orthogonality being different\n(comparing variance of orthogonality within vs. between)", yaxt = "n", ylab = "")
axis(2, las = 2, at = c(1:6), labels = c("Model 2 (sd = 13.65)", "Whole phylogeny (sd = 3.4)", "Passerida (sd = 19.89)", "Model 1 (sd = 11.64)", "Whole phylogeny (sd = 3.4)", "
    Passerida (sd = 19.89)"))
legend("topleft", col = c("orange", "blue"), pch =  19, lty =  1, legend = c("Model 1 (whole tree; sd = 11.64) vs.", "Model 2 (passeriformes; sd = 13.65) vs."))

col <- c(rep("orange", 3), rep("blue", 3))
for(line in 1:6) {
    lines(matrix(c(passeriformes_comparisons[line, c(2,3)], rep(line, 2)), 2, 2), col = col[line])
    points(passeriformes_comparisons[line, 1], line, pch = 19, col = col[line])
}
dev.off()
```

## PGLS

```{r}
## Get the median estimates for elaboration and innovation
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_order.rda")
load("../Data/Processed/groups_median_EI.rda")
load("../Data/Processed/shapespace.rda")

## Initialise the data
model_data <- data.frame("orders" = shapespace$level2, "super_orders" = shapespace$level1)
rownames(model_data) <- rownames(shapespace) 
model_data <- cbind(model_data,
                    centroid            = ploting_data$distances_to_centre[rownames(shapespace)],
                    species_elaboration = ploting_data_order$median_elaborations[rownames(shapespace)],
                    species_innovation  = ploting_data_order$median_innovations[rownames(shapespace)],
                    orders_elaboration = NA, orders_innovation = NA)
## Attributing the median innov/elab value for each order
for(order in unique(model_data$orders)) {
    if(order != "") { # This is because a couple of species have no order
        group_num <- grep(order, names(groups_median_EI$orders$innovation))
        model_data[which(model_data$orders == order), "orders_elaboration"] <- groups_median_EI$orders$elaboration[group_num]
        model_data[which(model_data$orders == order), "orders_innovation"] <- groups_median_EI$orders$innovation[group_num]
    }
}

## Remove the tips left over from the phylogeny
trees_list <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")

run.models <- function(tree, data) {
    ## all predictors
    cat(".")
    all_preds <- phylolm::phylolm(centroid ~ species_elaboration + species_innovation + orders_elaboration + orders_innovation,
                         data = data, phy = tree, model = "lambda")
    ## macro predictors
    macro_preds <- phylolm::phylolm(centroid ~ species_elaboration + species_innovation,
                           data = data, phy = tree, model = "lambda")
    ## mega predictors
    mega_preds <- phylolm::phylolm(centroid ~ orders_elaboration + orders_innovation,
                          data = data, phy = tree, model = "lambda")
    return(list(all = all_preds, macro = macro_preds, mega = mega_preds))
}

## Run 100 models (9 sec per model - 15 minutes for 100 models)
pgls_models_out <- lapply(trees_list[sample(1:length(trees_list), size = 100)], run.models, data = model_data)
save(pgls_models_out, file = "../Data/Processed/pgls_models_out.rda")

## Summarise the models
summary_out <- lapply(pgls_models_out, lapply, summary)

## Extract the slopes, p-values and r2
extract.summary <- function(one_summary) {
    ## Get the fits
    fits <- unlist(lapply(one_summary, function(X) return(X$adj.r.squared)))
    names(fits) <- paste0(names(fits), ".r2")
    ## Get the coefficients and their p vals
    coeffs <- lapply(one_summary, function(X) return(X$coefficient))

    get.est <- function(one_coeffs) {
        vals <- c(one_coeffs[-1, c("Estimate", "p.value")])
        names(vals) <- c(paste0(rownames(one_coeffs[-1, ]), c(".slope")), paste0(rownames(one_coeffs[-1, ]), c(".pval")))
        return(vals)
    }
    return(c(fits, unlist(lapply(lapply(one_summary, function(X) return(X$coefficient)), get.est))))
}
all_estimates <- do.call(rbind, lapply(summary_out, extract.summary))

## Output table
summary_table <- data.frame(matrix(NA, ncol = 8, nrow = 8))
colnames(summary_table) <- c("model", "predictor", "median slope", "5% slope", "95% slope", "median p.val", "5% p.val", "95% p.val")
summary_table$model <- c(rep("all", 4), rep("macro", 2), rep("mega", 2))
summary_table$predictor <- rep(c("species elaboration", "species innovation", "orders elaboration", "orders innovation"), 2)
summary_table[, 3:5] <- t(apply(all_estimates[, c(4:7,12,13,16,17)], 2, quantile, probs = c(0.5, 0.05, 0.95)))
summary_table[, 6:8] <- t(apply(all_estimates[, -c(1:3,4:7,12,13,16,17)], 2, quantile, probs = c(0.5, 0.05, 0.95)))

```

## Random skewers

```{r}
## Random skewers function wrapper for dispRity
skewers.fun <- function(matrix, matrix2) {
    return(phytools::skewers(matrix, matrix2)$r)
}
skewers.fun.p <- function(matrix, matrix2) {
    return(phytools::skewers(matrix, matrix2)$p)
}
p.star <- function(x) {
    if(x > 0.1) {
        return("")
    } else {
        if(x > 0.05) {
            return(".")
        } else {
            if(x > 0.01) {
                return("*")
            } else {
                if(x > 0.001) {
                    return("**")
                } else {
                    return("***")
                }
            }
        }
    }
}

## Getting the same order as the orthogonality data
ortho_order <- c(
    "Columbimorphae",
    "Galloanserae",
    "Mirandornithes",
    "Telluraves",
    "Strisores",
    "Paleognathae",
    "Aequornithes",
    "Otidimorphae",
    "Galliformes",
    "Accipitriformes",
    "Strigiformes",
    "Musophagiformes",
    "Trogoniformes",
    "Columbimorphae",
    "Caprimulgiformes",
    "Bucerotiformes",
    "Falconiformes",
    "Pterocliformes",
    "Sphenisciformes",
    "Podicipediformes",
    "Ciconiiformes",
    "Tinamiformes",
    "Suliformes",
    "Anseriformes",
    "Apodiformes",
    "Otidiformes",
    "Cuculiformes",
    "Psittaciformes",
    "Gruiformes",
    "Pelecaniformes",
    "Procellariiformes",
    "Piciformes",
    "Passeriformes",
    "Coraciiformes",
    "Charadriiformes")

## Getting the group pairs
group_pairs <- sapply(name.subsets(all_birds_shapespace)[-36], function(x) c(x, "phylogeny"), simplify = FALSE)
group_pairs_orders <- list(
    "Galliformes" = c("Galliformes", "Galloanserae"),
    "Accipitriformes" = c("Accipitriformes", "Telluraves"),
    "Strigiformes" = c("Strigiformes", "Telluraves"),
    "Musophagiformes" = c("Musophagiformes", "Aequornithes"),
    "Trogoniformes" = c("Trogoniformes", "Telluraves"),
    "Columbiformes" = c("Columbiformes", "Columbimorphae"),
    "Caprimulgiformes" = c("Caprimulgiformes", "Strisores"),
    "Bucerotiformes" = c("Bucerotiformes", "Telluraves"),
    "Falconiformes" = c("Falconiformes", "Telluraves"),
    "Pterocliformes" = c("Pterocliformes", "Columbimorphae"),
    "Sphenisciformes" = c("Sphenisciformes", "Aequornithes"),
    "Podicipediformes" = c("Podicipediformes", "Mirandornithes"),
    "Ciconiiformes" = c("Ciconiiformes", "Aequornithes"),
    "Tinamiformes" = c("Tinamiformes", "Paleognathae"),
    "Suliformes" = c("Suliformes", "Aequornithes"),
    "Anseriformes" = c("Anseriformes", "Galloanserae"),
    "Apodiformes" = c("Apodiformes", "Strisores"),
    "Otidiformes" = c("Otidiformes", "Otidimorphae"),
    "Cuculiformes" = c("Cuculiformes", "Otidimorphae"),
    "Psittaciformes" = c("Psittaciformes", "Telluraves"),
    "Pelecaniformes" = c("Pelecaniformes", "Aequornithes"),
    "Procellariiformes" = c("Procellariiformes", "Aequornithes"),
    "Piciformes" = c("Piciformes", "Telluraves"),
    "Passeriformes" = c("Passeriformes", "Telluraves"),
    "Coraciiformes" = c("Coraciiformes", "Telluraves"))

## Calculating the random skewers for each group
skewers_r <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun), between.groups = group_pairs, verbose = TRUE)
skewers_p <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun.p), between.groups = group_pairs, verbose = TRUE)
skewers_r_order <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun), between.groups = group_pairs_orders, verbose = TRUE)
skewers_p_order <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun.p), between.groups = group_pairs_orders, verbose = TRUE)

## Extract disparity for fancy plotting
skewers_distrib <- lapply(get.disparity(skewers_r, concatenate = FALSE), quantile, probs = c(0.5, 0.025, 0.25, 0.75, 0.975))
skewers_median_p <- lapply(get.disparity(skewers_p, concatenate = FALSE), median)
plot_names <- unlist(strsplit(names(skewers_distrib), split = ":phylogeny"))
names(skewers_distrib) <- names(skewers_median_p) <- plot_names
skewers_distrib <- skewers_distrib[ortho_order]
skewers_median_p <- skewers_median_p[ortho_order]
plot_names <- names(skewers_median_p)

## Same but for the orders
skewers_distrib_order <- lapply(get.disparity(skewers_r_order, concatenate = FALSE), quantile, probs = c(0.5, 0.025, 0.25, 0.75, 0.975))
skewers_median_p_order <- lapply(get.disparity(skewers_p_order, concatenate = FALSE), median)
plot_names_order <- unlist(lapply(strsplit(names(skewers_distrib_order), split = ":"), `[[`, 1))
names(skewers_distrib_order) <- names(skewers_median_p_order) <- plot_names_order

## Set the colours to match the ones from the ortho figure
col <- c(rep("purple", 8), rep("#FBB117", 27))

pdf(height = 12, width = 7, file = "../Manuscript/Figures/random_skewers.pdf")
par(mar = c(5, 17, 4, 2))
plot(NULL, xlim = c(0, 1), ylim = c(length(skewers_distrib),1), xlab = "Skewers correlation", ylab = "", yaxt = "n")
abline(v = 0.5, lty = 2, col = "grey", lwd = 0.5)
axis(2, at = 1:length(skewers_distrib), labels = plot_names, las = 2)
## Skewers
for(i in length(skewers_distrib):1) {
    lines(x = skewers_distrib[i][[1]][c(2, 5)], y = rep(i, 2), lty = 2, lwd = 1, col = col[i])
    lines(x = skewers_distrib[i][[1]][c(3, 4)], y = rep(i, 2), lty = 1, lwd = 2, col = col[i])
    points(x = skewers_distrib[i][[1]][1], y = i, pch = 19, col = col[i])
    text(x = skewers_distrib[i][[1]][1]+0.03, y = i-0.2, label = p.star(skewers_median_p[[i]]), col = col[i])
    ## Add orders (if available)
    if(names(skewers_distrib)[i] %in% plot_names_order) {
        n <- which(names(skewers_distrib)[i] == plot_names_order)
        col_n <- adjustcolor(col[i], alpha = 1/2)
        ## Plotting the order-superorder level results
        lines(x = skewers_distrib_order[n][[1]][c(2, 5)], y = rep(i, 2)+0.2, lty = 3, lwd = 1, col = col_n)
        lines(x = skewers_distrib_order[n][[1]][c(3, 4)], y = rep(i, 2)+0.2, lty = 1, lwd = 2, col = col_n)
        points(x = skewers_distrib_order[n][[1]][1], y = i+0.2, pch = 21, col = col_n)
        text(x = skewers_distrib_order[n][[1]][1]+0.03, y = i, label = p.star(skewers_median_p_order[[n]]), col = col_n)
    }
}
dev.off()
```