### 2023-06-19 Analysis/09-Figure_key.Rmd

---
title: "Bird beak's elaboration and innovation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This section contains the code for reproducing the analyses. To get the details, refer to 03-elaboration_innovation_analyses.Rmd.

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
source("../Functions/ploting.utilities.R")
source("../Functions/covar.utilities.R")
```

# All birds shapespace analyses

```{r, eval = TRUE, echo = FALSE}
## Load the data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Creating the subsets
superorder_shapespace <- MCMCglmm.subsets(n = 1000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], "phylogeny")) #[-1] is because the first level is ""

## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)
```

# Extract only the Mirandornithes

```{r, eval = TRUE, echo = FALSE}
## Selecting the paleo data
space_paleo <- get.subsets(superorder_shapespace, subsets = c("phylogeny", "Paleognathae"))

## PANEL 1

## General plot
pdf("../Manuscript/Figures/figure_key_space_plot.pdf")
plot(shapespace[, c(1,2)], col = "lightgrey", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Plotting the projection cartoon

## Do the projection
matrix <- as.matrix(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)])
rownames(matrix) <- NULL
## Get the base vectors
covars <- sample.n(space_paleo$covar, 4000)
centres <- lapply(unlist(space_paleo$subsets, recursive = FALSE) ,function(group, data, fun) fun(data[c(group), ]), data = space_paleo$matrix[[1]], fun = colMeans)
covars <- mapply(recentre, covars, centres, MoreArgs = list(dimensions = c(1,2)), SIMPLIFY = FALSE)
covars <- lapply(covars, function(one_covar, scale) mapply(scale.VCV, one_covar, scale, SIMPLIFY = FALSE), scale = covars[[2]])
covars_cent_tend <- lapply(covars, VCV.cent.tend, mean)
base_vectors <- lapply(lapply(covars_cent_tend, get.one.axis, axis = 1, level = 0.95, dimensions = c(1,2)), list)

## Get projection lines
get.proj.line <- function(point, base, just.proj = FALSE) {
  point <- t(point)
  base <- t(base)
  proj_x <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[1];
  proj_y <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[2];
  if(just.proj) {
      return(cbind(proj_x, proj_y))
  } else {
      return(rbind(cbind(proj_x, proj_y), point))
  }
}
get.any.proj.lines <- function(points, base) {
      ## 1 - Get the coordinates of the points
      points_coords <- points[, ]
            # points(points_coords, col = "grey")
      ## 2 - Get the coordinates of the base
      base_coords <- base
            # lines(base_coords, col = "grey")

      ## 3 - translate the points and base onto the origin of the base
      translation_vector <- c(0, 0) - base_coords[2,]
      points_coords_translated <- t(t(points_coords) + translation_vector)
      base_coords_translated <- t(t(base_coords) + translation_vector)
            # lines(base_coords_translated)
            # points(points_coords_translated)

      ## 4 - project all the translated points onto the translated base
      lines_out_translated <- t(apply(points_coords_translated, 1, get.proj.line, base = base_coords_translated, just.proj = TRUE))
            # lines(rbind(lines_out_translated[1, ], points_coords_translated[1, ]))

      ## 5 - untranslate the points
      lines_out <- t(t(lines_out_translated) - translation_vector)
            # lines(rbind(lines_out[1, ], points_coords[1, ]), col = "red")
      return(lines_out)
}

## Projection onto the phylo line
pdf("../Manuscript/Figures/figure_key_projection_on_phylo.pdf")
plot(NULL, xlab = "PC1", ylab = "PC2", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
## Project all the points on the base line
proj_points <- get.any.proj.lines(matrix, base = base_vectors$phylogeny[[1]])
## Plot all these lines
for(i in 1:nrow(matrix)) {
    lines(rbind(matrix[i, ], proj_points[i, ]), col = "grey")
}
## Add the points
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
## Add the ellipses
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Projection onto the group line
pdf("../Manuscript/Figures/figure_key_projection_on_group.pdf")
plot(NULL, xlab = "PC1", ylab = "PC2", xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5))
## Project all the points on the base line
proj_points <- get.any.proj.lines(matrix, base = base_vectors$Paleognathae[[1]])
## Plot all these lines
for(i in 1:nrow(matrix)) {
    lines(rbind(matrix[i, ], proj_points[i, ]), col = gg.color.hue(1))
}
## Add the points
points(shapespace[which(shapespace$level1 == "Paleognathae"), c(1,2)], col = gg.color.hue(1), pch = 19)
## Add the ellipses
covar.plot(space_paleo,
      col = c("grey", gg.color.hue(1)),
      ellipses = mean,
      major.axes = mean,
      legend = FALSE,
      points = FALSE,
      cex = 0.5, apply.to.VCV = TRUE,
      lwd = 4,
      add = TRUE,
      scale = 2)
dev.off()

## Phylo plot
## Get the tree
tree_paleo <- drop.tip(tree, tip = rownames(shapespace)[which(shapespace$level1 != "Paleognathae")])
## Ladderize
tree_paleo <- ladderize(tree_paleo)

## Get the data (in the right order)
load("../Data/Processed/ploting_data_order.rda")
elaborations <- ploting_data_order$median_elaborations[tree_paleo$tip.label]
innovations <- ploting_data_order$median_innovations[tree_paleo$tip.label]

## Transform them into a colour gradient
## Change colours to grey -> blue innov. grey -> orange elab
col.grad.inno <- colorRamp(c("grey", "blue"))
col.grad.elab <- colorRamp(c("grey", "orange"))
# col.grad <- colorRamp(rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#66c2a5", "#3288bd")))
cols_elab <- col.grad.elab (elaborations/max(elaborations))
cols_inno <- col.grad.inno(innovations/max(innovations))

## Plot the tree
pdf("../Manuscript/Figures/figure_key_tree_example.pdf")
plot(tree_paleo, show.tip.label = FALSE, type = "fan")
tiplabels("", frame = "none", pch = 19, cex = 2, col = rgb(cols_elab, maxColorValue = 256), offset = -1)
tiplabels("", frame = "none", pch = 19, cex = 2, col = rgb(cols_inno, maxColorValue = 256), offset = 4)
dev.off()

# Focal species is Dromaius
```