### 2024-02-01 Analysis/update_procrustes_and_PCA.Rmd
Create update_procrustes_and_PCA.Rmd

---
title: "Updates a Procrustes and a PCA"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

Trying to update a Procrustes superimposition by adding new species to it (i.e. run the Procrustes and then add a new species following the same deformation).
And doing the same for a PCA (i.e. adding some new data to a previously calculated eigen transformation).

```{r}
library(Morpho)
require(landvR) # devtools::install_github("TGuillerme/landvR")
## Get some 3D data from Morpho
data(boneData)

## Run a procrustes and PCA
proc <- procSym(boneLM)
## Create a new specimens: the average, the median and the minimum, of all bones)
new_spec <- array(data = c(apply(boneLM, c(1,2), mean),
                           apply(boneLM, c(1,2), median),
                           unlist(landvR::select.procrustes(proc$rotated, selector = min))),
                           dim = c(10, 3, 3))
## Add them to the procrustes + PCA
new_proc <- align2procSym(proc, new_spec)
dimnames(new_proc)[[3]] <- c("mean", "median", "min")
## Transforming the procrustes arrays into matrices
proc_matrix <- landvR::array.to(proc$rotated, to = "matrix")
new_proc_matrix <- landvR::array.to(new_proc, to = "matrix")

## Run a PCA on the original data
pca <- prcomp(proc_matrix)
plot(pca$x[,c(1,2)])
## Note that this is equal to the pca from Morpho but rotated (*-1)
# plot(proc$PCscores[, c(1,2)]*-1)

## Run the PCA with the added procrustes
new_pca <- rbind(pca$x, predict(pca, newdata = new_proc_matrix))
points(new_pca[-c(81:83), c(1,2)], pch = 19, cex = 0.5)
points(new_pca[c(81:83), c(1,2)], pch = 19, col = "red")
legend("topright", pch = c(21, 19, 19), col = c("black", "black", "red"), pt.cex = c(1, 0.5, 1), legend = c("original PCA", "updated PCA", "new specimen"))
```