### 2025-04-14 Analysis/17-update_procrustes_and_PCA.Rmd
Update 17-update_procrustes_and_PCA.Rmd

---
title: "Updates a Procrustes and a PCA"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

In this script we will be looking at how to update both a Procrustes superimposition (GPA) and a PCA by projecting new specimens in it.
The goal is to use the GPA and PCA from a previous dataset (published with Guillerme et al 2023 - Science Advances) and to project new specimens onto it.

# The pipeline (testing)

Here is an example on how to do it using the `Morpho` package for the base GPA and PCA:

```{r}
library(Morpho)
library(landvR) # install from devtools::install_github("TGuillerme/landvR")

## Get some 3D data from Morpho
data(boneData)
original_landmarks <- boneLM

## Run the GPA for the original data
proc_orig <- Morpho::procSym(original_landmarks)

## Run the PCA for the original data
proc_matrix_orig <- landvR::array.to(proc_orig$rotated, to = "matrix")
pca_orig <- prcomp(proc_matrix_orig)

## Plot the results
plot(pca_orig$x[,c(1,2)])
```

And now let's generate some extra data out of the pipeline.
Here we're just creating a mean, median and minimum landmark set using the mean and median landmark coordinates values as well as one of a specific specimen (42).

```{r}
## Create a new specimens: the average, the median and the minimum, of all bones)
new_spec <- array(data = c(apply(original_landmarks, c(1,2), mean),
                           apply(original_landmarks, c(1,2), median),
                           original_landmarks[,,42]), dim = c(10, 3, 3))
dimnames(new_spec)[[3]] <- paste0("new_", c("mean", "median", "spec42"))

## Here is what the landmark positions look like in 3D
op <- par(mfrow = c(2,2))
plot(original_landmarks[, c(1,2), 1], main = "First specimen\n(reference from real dataset)")
plot(new_spec[, c(1,2), 1], main = "New specimen 1\n(average)")
plot(new_spec[, c(1,2), 2], main = "New specimen 2\n(median)")
plot(new_spec[, c(1,2), 3], main = "New specimen 3\n(spec 42)")
par(op)
```

We can then add these specimens to the GPA and PCA.

```{r}
## Updating the gpa and pca
updated_pca <- update.gpa.pca(proc_orig, new_spec)

## Plot all the results
plot(proc_orig$PCscores[,c(1,2)])
points(updated_pca[-c(81:83), c(1,2)], pch = 19, cex = 0.5)
points(updated_pca[c(81:83), c(1,2)], pch = 19, col = "red")
legend("topright", pch = c(21, 19, 19), col = c("black", "black", "red"), pt.cex = c(1, 0.5, 1), legend = c("original PCA", "updated PCA", "new specimen"))
```

Seems to work quiet well!
All the positions from previous specimens are respected.
And the two new median and average specimens plot in the centre of the traitspace (as expected) with the weird one plotting on some edge of the traitspace.

But does it work with the same specimens? I.e. does it place previous specimens in the correct position?

```{r}
## Old specimens
old_spec <- array(data = c(original_landmarks[,,1],
                           original_landmarks[,,2],
                           original_landmarks[,,3]), dim = c(10, 3, 3))
dimnames(old_spec)[[3]] <- paste0("old_", c("spec1", "spec2", "spec3"))

## Updating the gpa and pca
updated_pca <- update.gpa.pca(proc_orig, old_spec, center.pca = TRUE, scale.pca = FALSE)

## Plot all the results
plot(proc_orig$PCscores[,c(1,2)], col = "grey")
points(updated_pca[-c(81:83), c(1,2)], pch = 19, cex = 0.5, col = "grey")

points(proc_orig$PCscores[c(1,2,3),c(1,2)], col = "blue", pch = 21)
points(updated_pca[c(81:83),c(1,2)], col = "orange", pch = 19)

legend("topright", pch = c(21, 19, 21, 19), col = c("grey", "grey", "blue", "orange"), pt.cex = c(1, 0.5, 1, 0.5), legend = c("original PCA", "updated PCA", "old_spec (target)", "old_spec (results)"))
```

This is not ideal!
The reason why it doesn't recover them is because it slightly offsets the superimposition which cascades to big changes in the PCA.

For example, when using the same specimens:

```{r}
## Plotting the landmarks before superimposition
op <- par(mfrow = c(2,2))
plot(old_spec[, c(1,2), 1], main = "2D landmarks for spec 1", xlab = "", ylab = "")
plot(old_spec[, c(1,2), 2], main = "2D landmarks for spec 2", xlab = "", ylab = "")
plot(old_spec[, c(1,2), 3], main = "2D landmarks for spec 3", xlab = "", ylab = "")
par(op)

## Project the new specimens on the original GPA
new_gpa <- Morpho::align2procSym(proc_orig, old_spec, orp = TRUE)

## Slightly off?
op <- par(mfrow = c(2,2))
## Spec 1
plot.offset <- function(id, new_gpa, old_gpa) {
    main <- paste0("spec ", id, " reprojected onto itself\nTotal offset = ", sum(sqrt((new_gpa[, 1, id] - old_gpa[, 1, id])^2 + (new_gpa[, 2, id] - old_gpa[, 2, id])^2)))
    plot(new_gpa[, c(1,2), id], main = main, xlab = "", ylab = "")
    points(old_gpa[, c(1,2), id], pch = 19, cex = 0.5)
}

plot.offset(id = 1, new_gpa, proc_orig$rotated)
plot.offset(id = 2, new_gpa, proc_orig$rotated)
plot.offset(id = 3, new_gpa, proc_orig$rotated)

main <- paste0("spec 1 projected onto spec 2 (for ref.)\nTotal offset = ", sum(sqrt((proc_orig$rotated[, 1, 1] - proc_orig$rotated[, 1, 2])^2 + (proc_orig$rotated[, 2, 1] - proc_orig$rotated[, 2, 2])^2)))
plot(proc_orig$rotated[, c(1,2), 1], main = main, xlab = "", ylab = "")
points(proc_orig$rotated[, c(1,2), 2], pch = 19, cex = 0.5)

par(op)
```

This behaviour is actually highlighted in the notes for `align2procSym`: "this will never yield the same result as a pooled Procrustes analysis because the sample mean is iteratively updated and new data would change the mean."

> TL;DR: it's not working until we find a way to not update the mean.

# Mapping the chaffinches on the whole traitspace

```{r}
## Loading the entire bird GPA
library(Morpho)
whole_gpa <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")

## Reading in the chaffinches
land_pairs <- read.table("../Data/Raw/Beak_data/landpairs.txt")

## Read all the chaffinches
path <- "../Data/Raw/Beak_data/Chaffinch_data/"
individuals <- list.files(path)
all_individuals <- list()
while(length(individuals) != 0) {
    file <- paste0(path, individuals[1], "/", list.files(paste0(path, individuals[1]))) 
    if(grepl(".nts", file)) { 
        ## Does contain a file
        data <- readland.nts(paste0(path, individuals[1], "/", list.files(paste0(path, individuals[1]))))
        name <- paste0(strsplit(individuals[1], split = "_")[[1]][1:4], collapse = "_")
        all_individuals[[length(all_individuals) + 1]] <- data[, , 1]
        names(all_individuals)[length(all_individuals)] <- name
    } #else {
        # print(paste0(file, ": empty"))
    # }
    individuals <- individuals[-1]
}

## The following are missing:
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_africana_14_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_africana_19_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_coelebs_1_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_gengleri_1_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_gengleri_21_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_gengleri_5_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_moreletti_11_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_coelebs_palmae_13_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_teydea_polatzeki_1_intraspecific/: empty"
# [1] "../Data/Raw/Beak_data/Chaffinch_data/Fringilla_teydea_teydea_10_intraspecific/: empty"

## Converting into an array
individuals_array <- simplify2array(all_individuals)

## Project them! (*-1 is to flip it around)
updated_pca <- update.gpa.pca(whole_gpa, whole_gpa$rotated[,,c(2819, 2820)], scale.pca = FALSE)

## Some big shift!
plot(whole_gpa$PCscores[,c(1,2)], col = "grey")
points(updated_pca[-c((dim(whole_gpa$PCscores)[1]+1): dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "grey")
## Plot the chaffinches to see where they lay:
fringillas <- which(grepl("Fringilla", rownames(whole_gpa$PCscores)))
points(whole_gpa$PCscores[fringillas,c(1,2)], pch = 21, col = "blue")
points(updated_pca[c((dim(whole_gpa$PCscores)[1]+1):dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "orange")
legend("bottomright", pch = c(21, 19, 21, 19), col = c("grey", "grey", "blue", "orange"), pt.cex = c(1, 0.5, 1, 0.5), legend = c("original PCA", "updated PCA", "Fringilla (original)", "New specimens"))
```

```{r}
library(Morpho)

## Get some 3D data from Morpho
data(boneData)
original_landmarks <- boneLM

## Run the GPA for the original data
proc_orig <- Morpho::procSym(original_landmarks)

## Creating some new specimens (two old ones, for benchmarking the method)
new_spec <- array(c(original_landmarks[,,1],
                    original_landmarks[,,2]), dim = c(10, 3, 2))

## Project the new specimens on the original GPA
new_gpa <- Morpho::align2procSym(proc_orig, new_spec, orp = TRUE)

## Slightly off?
plot.offset <- function(id, new_gpa, old_gpa, main) {
    main <- paste0(main, "\nTotal offset = ", sum(sqrt((new_gpa[, 1, id] - old_gpa[, 1, id])^2 + (new_gpa[, 2, id] - old_gpa[, 2, id])^2)))
    plot(new_gpa[, c(1,2), id], main = main, xlab = "", ylab = "")
    points(old_gpa[, c(1,2), id], pch = 19, cex = 0.5)
}

op <- par(mfrow = c(2,1))
plot.offset(id = 1, new_gpa, proc_orig$rotated, main = "Spec 1 projected onto itself")
plot.offset(id = 2, new_gpa, proc_orig$rotated, main = "Spec 2 projected onto itself")
par(op)

data(boneData)
# run procSym on entire data set
proc <- procSym(boneLM)
# this is the training data
array1 <- boneLM[,,1:60]
proc1 <- procSym(array1)
newdata <- boneLM[,,1:2]
newalign <- align2procSym(x = proc1, newdata = newdata)

plot.offset(id = 1, newalign, proc1$rotated, main = "Spec 1 projected onto itself")

```

TODO: try doing the alignement of some species on the edges of the traitspace and see if it falls back on the edges or if it goes to the middle again.

TODO: check with the `morphospace` package?