### 2025-04-28 Analysis/17-update_procrustes_and_PCA.Rmd
Update 17-update_procrustes_and_PCA.Rmd

---
title: "Updates a Procrustes and a PCA"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

In this script we will be looking at how to update both a Procrustes superimposition (GPA) and a PCA by projecting new specimens in it.
The goal is to use the GPA and PCA from a previous dataset (published with Guillerme et al 2023 - Science Advances) and to project new specimens onto it.

# The pipeline (testing)

TODO: TRY CENTERING THE DATA BEFORE DOING THE PCA

Here is an example on how to do it using the `Morpho` package for the base GPA and PCA:

```{r}
library(Morpho)
library(landvR) # install from devtools::install_github("TGuillerme/landvR")

## Get some 3D data from Morpho
data(boneData)
original_landmarks <- boneLM

## Run the GPA for the original data
proc_orig <- Morpho::procSym(original_landmarks)

## Run the PCA for the original data
proc_matrix_orig <- landvR::array.to(proc_orig$rotated, to = "matrix")
pca_orig <- prcomp(proc_matrix_orig)

## Plot the results
plot(pca_orig$x[,c(1,2)])
```

And now let's generate some extra data out of the pipeline.
Here we're just creating a mean, median and minimum landmark set using the mean and median landmark coordinates values as well as one of a specific specimen (42).

```{r}
## Create a new specimens: the average, the median and the minimum, of all bones)
new_spec <- array(data = c(apply(original_landmarks, c(1,2), mean),
                           apply(original_landmarks, c(1,2), median),
                           original_landmarks[,,42]), dim = c(10, 3, 3))
dimnames(new_spec)[[3]] <- paste0("new_", c("mean", "median", "spec42"))

## Here is what the landmark positions look like in 3D
op <- par(mfrow = c(2,2))
plot(original_landmarks[, c(1,2), 1], main = "First specimen\n(reference from real dataset)")
plot(new_spec[, c(1,2), 1], main = "New specimen 1\n(average)")
plot(new_spec[, c(1,2), 2], main = "New specimen 2\n(median)")
plot(new_spec[, c(1,2), 3], main = "New specimen 3\n(spec 42)")
par(op)
```

We can then add these specimens to the GPA and PCA.

```{r}
## Updating the gpa and pca
updated_pca <- update.gpa.pca(proc_orig, new_spec)

## Plot all the results
plot(proc_orig$PCscores[,c(1,2)])
points(updated_pca[-c(81:83), c(1,2)], pch = 19, cex = 0.5)
points(updated_pca[c(81:83), c(1,2)], pch = 19, col = "red")
legend("topright", pch = c(21, 19, 19), col = c("black", "black", "red"), pt.cex = c(1, 0.5, 1), legend = c("original PCA", "updated PCA", "new specimen"))
```

Seems to work quiet well!
All the positions from previous specimens are respected.
And the two new median and average specimens plot in the centre of the traitspace (as expected) with the weird one plotting on some edge of the traitspace.

But does it work with the same specimens? I.e. does it place previous specimens in the correct position?

```{r}
## Old specimens
old_spec <- array(data = c(original_landmarks[,,1],
                           original_landmarks[,,2],
                           original_landmarks[,,3]), dim = c(10, 3, 3))
dimnames(old_spec)[[3]] <- paste0("old_", c("spec1", "spec2", "spec3"))

## Updating the gpa and pca
updated_pca <- update.gpa.pca(proc_orig, old_spec, center.pca = TRUE, scale.pca = FALSE)

## Plot all the results
plot(proc_orig$PCscores[,c(1,2)], col = "grey")
points(updated_pca[-c(81:83), c(1,2)], pch = 19, cex = 0.5, col = "grey")

points(proc_orig$PCscores[c(1,2,3),c(1,2)], col = "blue", pch = 21)
points(updated_pca[c(81:83),c(1,2)], col = "orange", pch = 19)

legend("topright", pch = c(21, 19, 21, 19), col = c("grey", "grey", "blue", "orange"), pt.cex = c(1, 0.5, 1, 0.5), legend = c("original PCA", "updated PCA", "old_spec (target)", "old_spec (results)"))
```

This is not ideal!
The reason why it doesn't recover them is because it slightly offsets the superimposition which cascades to big changes in the PCA.

For example, when using the same specimens:

```{r}
## Plotting the landmarks before superimposition
op <- par(mfrow = c(2,2))
plot(old_spec[, c(1,2), 1], main = "2D landmarks for spec 1", xlab = "", ylab = "")
plot(old_spec[, c(1,2), 2], main = "2D landmarks for spec 2", xlab = "", ylab = "")
plot(old_spec[, c(1,2), 3], main = "2D landmarks for spec 3", xlab = "", ylab = "")
par(op)

## Project the new specimens on the original GPA
new_gpa <- Morpho::align2procSym(proc_orig, old_spec, orp = TRUE)

## Slightly off?
op <- par(mfrow = c(2,2))
## Spec 1
plot.offset <- function(id, new_gpa, old_gpa) {
    main <- paste0("spec ", id, " reprojected onto itself\nTotal offset = ", sum(sqrt((new_gpa[, 1, id] - old_gpa[, 1, id])^2 + (new_gpa[, 2, id] - old_gpa[, 2, id])^2)))
    plot(new_gpa[, c(1,2), id], main = main, xlab = "", ylab = "")
    points(old_gpa[, c(1,2), id], pch = 19, cex = 0.5)
}

plot.offset(id = 1, new_gpa, proc_orig$rotated)
plot.offset(id = 2, new_gpa, proc_orig$rotated)
plot.offset(id = 3, new_gpa, proc_orig$rotated)

main <- paste0("spec 1 projected onto spec 2 (for ref.)\nTotal offset = ", sum(sqrt((proc_orig$rotated[, 1, 1] - proc_orig$rotated[, 1, 2])^2 + (proc_orig$rotated[, 2, 1] - proc_orig$rotated[, 2, 2])^2)))
plot(proc_orig$rotated[, c(1,2), 1], main = main, xlab = "", ylab = "")
points(proc_orig$rotated[, c(1,2), 2], pch = 19, cex = 0.5)

par(op)
```

Checking all over again with `princomp`

```{r}
## New test
library(Morpho)
library(landvR) # install from devtools::install_github("TGuillerme/landvR")

## Get some 3D data from Morpho
data(boneData)
original_landmarks <- boneLM

## Run the GPA for the original data
proc_orig <- Morpho::procSym(original_landmarks)

## New species (three to anchor and two new ones)
new_spec <- array(data = c(original_landmarks[,,1],
                           original_landmarks[,,2],
                           original_landmarks[,,3],
                           apply(original_landmarks, c(1,2), median),
                           apply(original_landmarks, c(1,2), min)),  dim = c(10, 3, 5))
dimnames(new_spec)[[3]] <- c(paste0("old_", c("spec1", "spec2", "spec3")), "median", "min")

## Run the PCA for the original data
pca_orig <- prcomp(landvR::array.to(proc_orig$rotated, to = "matrix"))
princomp_orig <- princomp(landvR::array.to(proc_orig$rotated, to = "matrix"))

## THIS PIPELINE SEEMS TO WORK:

## Test the reprojection using princomp
new_gpa <- Morpho::align2procSym(proc_orig, new_spec)
new_points <- stats::predict(princomp_orig, newdata = array.to(new_gpa, to = "matrix"))
plot(princomp_orig$scores[,c(1,2)], col = "grey")
points(new_points[,c(1,2)], col = "orange", pch = 19)
points(princomp_orig$scores[c(1,2,3),c(1,2)], col = "blue", pch = 21)

## Thest using prcomp
new_gpa <- Morpho::align2procSym(proc_orig, new_spec)
new_points <- stats::predict(pca_orig, newdata = array.to(new_gpa, to = "matrix"))
plot(pca_orig$x[,c(1,2)], col = "grey")
points(new_points[,c(1,2)], col = "orange", pch = 19)
points(pca_orig$x[c(1,2,3),c(1,2)], col = "blue", pch = 21)

## Test with pipeline
updated_pca <- update.gpa.pca(proc_orig, new_spec)*-1 # TODO: fix the -1
plot(proc_orig$PCscores[,c(1,2)])
points(updated_pca[,c(1,2)], col = "grey", pch = 19)
points(updated_pca[c(81:85),c(1,2)], col = "orange", pch = 19)
points(updated_pca[c(81:83),c(1,2)], col = "blue", pch = 21)

## Not really equal though!
expect_equal(updated_pca[-c(81:85),c(1,2)], proc_orig$PCscores[,c(1,2)])

```

This behaviour is actually highlighted in the notes for `align2procSym`: "this will never yield the same result as a pooled Procrustes analysis because the sample mean is iteratively updated and new data would change the mean."

> TL;DR: it's not working until we find a way to not update the mean.

# Mapping the chaffinches on the whole traitspace

```{r}
## Loading the entire bird GPA
library(Morpho)
whole_gpa <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")

## Reading in the chaffinches
land_pairs <- read.table("../Data/Raw/Beak_data/landpairs.txt")

## Read all the chaffinches
path <- "../Data/Raw/Beak_data/Chaffinch_data/"
individuals <- list.files(path)
all_individuals <- list()
while(length(individuals) != 0) {
    file <- paste0(path, individuals[1], "/", list.files(paste0(path, individuals[1]))) 
    if(grepl(".nts", file)) { 
        ## Does contain a file
        data <- readland.nts(paste0(path, individuals[1], "/", list.files(paste0(path, individuals[1]))))
        name <- paste0(strsplit(individuals[1], split = "_")[[1]][1:4], collapse = "_")
        all_individuals[[length(all_individuals) + 1]] <- data[, , 1]
        names(all_individuals)[length(all_individuals)] <- name
    } #else {
        # print(paste0(file, ": empty"))
    # }
    individuals <- individuals[-1]
}

## Converting into an array
new_individuals <- simplify2array(all_individuals)

## Project them! (*-1 is to flip it around)
updated_pca <- update.gpa.pca(whole_gpa, new_individuals)*-1

# No pipeline
# Get the CPA
original_data <- whole_gpa
new_landmarks <- new_individuals
original_pca <- prcomp(landvR::array.to(original_data$rotated, to = "matrix")) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
# original_pca <- prcompfast(original_data$tan) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
## Project the new specimens on the original GPA
new_data <- Morpho::align2procSym(original_data, new_landmarks)
## Project the new GPAed specimens onto the PCA
new_pca <- stats::predict(original_pca, newdata = array.to(new_data, to = "matrix"))
## Combine the new and old matrices for the output PCA
updated_pca <- rbind(original_pca$x, new_pca) *-1

## Some big shift!
plot(whole_gpa$PCscores[,c(1,2)], col = "grey")
points(updated_pca[-c((dim(whole_gpa$PCscores)[1]+1): dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "grey")
## Plot the chaffinches to see where they lay:
fringillas <- which(grepl("Fringilla", rownames(whole_gpa$PCscores)))
points(whole_gpa$PCscores[fringillas,c(1,2)], pch = 21, col = "blue")
points(updated_pca[c((dim(whole_gpa$PCscores)[1]+1):dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "orange")
legend("bottomright", pch = c(21, 19, 21, 19), col = c("grey", "grey", "blue", "orange"), pt.cex = c(1, 0.5, 1, 0.5), legend = c("original PCA", "updated PCA", "Fringilla (original)", "New specimens"))
```